---
title: "GoData: let's go clean"
author: "Thibaut Jombart, Sara Hollis, Christopher Jarvis"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: pygments
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 2
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---



<br>

<div class="report_meta">
  <span class="notice">**Notice**: this is a **stable, routine report**. 
  **Do not touch it unless it is broken.** To make a contribution, **carefully read 
  the [README](../../../../../README.html) file**.</span>
  
  **Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)
  
  **Code contributors:** Thibaut Jombart, Sara Hollis, Christopher Jarvis
  
  **Data contributors:** GoData team Beni
    
  **Reviewed by:** 
</div>





<!-- ====================================================== -->
<!-- ====================================================== -->
<!-- ====================================================== -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This is the universal cleaning script for raw data exported from godata.

### Data used

This report uses the latest exported .xlsx from godata.

### Method

The data preparation involves the following steps, detailed in the following
tabs:

* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Add new variables**: addition of new variables to the case and contact lists;
importantly adding hierarchy levels that are missing in initial case and contact lists

* **Filter the data**: keep only relevant confirmed and probable cases, possibly
  removing erroneous dates, for further analysis



<!-- ====================================================== -->
## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned MDC data stored as `x`

**Important**: we need to make sure the soucing of `R` scripts is done using the
current environment, using the argument `local = TRUE`. This is in particular
essential when `params` is used, as some functions and settings are dependent
upon it.

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names=TRUE)
for (file in scripts_files) source(file, local = TRUE)

ggthemr("grape")

## show parameters
params

```




<!-- ====================================================== -->
## Load data

We extract the completion date from the file names, for:

1. cases data

2. contacts data

3. relationships between cases and contacts

4. contact follow up data

And also add:

5. godata hierarchy


```{r load_data}

## step 1
cases_file
cases <- custom_import(cases_file) %>%
  as_tibble()


## step 2
contacts_file
contacts <- custom_import(contacts_file) %>%
  as_tibble()


## step 3
relationships_file
relationships <- custom_import(relationships_file) %>%
  as_tibble()

## step 4
followups_file
followups <- custom_import(followups_file) %>%
  as_tibble()

## step 5
hierarchy_file
hierarchy <- custom_import(hierarchy_file) %>%
  as_tibble()

```


<!-- ====================================================== -->
## Completion date

We extract the completion date from the name of the cases file.

```{r database_date}

## extract date of database based on cases file
cases_file_short <- gsub("^[^.]+/", "", cases_file)
database_date <- cases_file_short %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.





<!-- ====================================================== -->
## Data standardisation

Here we standardise data using `clean_data`, which set all data to lower case,
removes special characters, replaces non-ascii characters with their closest
ascii match.

```{r clean_data}

## standardise all data
cases <- clean_data(cases, guess_dates = FALSE)
contacts <- clean_data(contacts, guess_dates = FALSE)
relationships <- clean_data(relationships, guess_dates = FALSE)
followups <- clean_data(followups, guess_dates = FALSE)
hierarchy <- clean_data(hierarchy, guess_dates = FALSE)

```
  



<!-- ====================================================== -->
## Dates processing

Here we ensure that dates we will be using are indeed dates, and convert them
where necessary using `guess_dates`. We also make the following changes:

* `date_of_followup_end`: we erase the database version, which seems to have an
  issue, setting any date in the future compared to the database date to `NA`;
  this field is now calculated directly from `date_of_last_contact`

```{r clean_dates}

cases <- cases %>%
  mutate(date_of_reporting = guess_dates(date_of_reporting),
         date_of_onset = guess_dates(date_of_onset),
         date_of_outcome = guess_dates(date_of_outcome),
         date_of_infection = guess_dates(date_of_infection),
         date_of_becoming_case = guess_dates(date_of_becoming_case),
         date_of_burial = guess_dates(date_of_burial),
         date_of_case_notification = guess_dates(date_de_notification_du_cas))

contacts <- contacts %>%
  mutate(date_of_reporting = guess_dates(date_of_reporting),
         date_of_data_entry = guess_dates(created_at),
         date_of_becoming_case = guess_dates(date_of_becoming_case),
         date_of_onset = guess_dates(date_of_onset),
         date_of_last_contact = guess_dates(date_of_last_contact),
         date_of_followup_start = guess_dates(start_of_the_followup),
         date_of_followup_end = date_of_last_contact + 21)

relationships <- relationships %>%
  mutate(date_of_data_entry = guess_dates(created_at),
         date_of_last_contact = guess_dates(date_of_last_contact))

followups <- followups %>%
  mutate(date_of_data_entry = guess_dates(created_at),
         date_of_followup = guess_dates(date),
         date_of_onset = guess_dates(dateapparitionsignes),
         date_of_hospitalization = guess_dates(dateadmissioncte),
         date_of_outcome = guess_dates(datededeces))

```



<!-- ====================================================== -->
## Age variable standardization

Here we ensure that all age info is captured in one field `age_years` across all
datasets.

```{r clean_age}

cases <- cases %>%
  mutate(age_years = case_when(
                        is.na(age_years) ~ (age_months / 12),
                        TRUE ~ age_years))

contacts <- contacts %>%
  mutate(age_years = case_when(
                        is.na(age_years) ~ (age_months / 12),
                        TRUE ~ age_years))

```


<!-- ====================================================== -->
## Adding locations

Unfortunately at present as of 2019-11-02 the locations are in a graphical database and only the lowest level of location is stored in the contacts data. 

In order to merge on the location data at multiple levels we need to do multiple join and filter the data. 

Process

- get the id and location in a data frame. 
- join on an lowest area
- add that area to the data
- filter the data to keep
- filter the data for the next area
- repeat


```{r filter_by_locations}

hierarchy_join <- select(hierarchy, starts_with("name"))
x_join <- contacts %>% select(id, location_1)
  


### Name 6
# Join on name_6 area
n_6 <- left_join(x_join, hierarchy_join, by = c("location_1" = "name_6"))
# create name_6 variable as it doesn't stay after merge
n_6$name_6 <- NA
# Fill in name_6 for non-missing
n_6$name_6[!is.na(n_6$name_5)] <- n_6$location_1[!is.na(n_6$name_5)]

# Keep the name_6 for later
n_6_keep <- n_6 %>% 
  filter(!is.na(n_6$name_5)) %>% 
  select(id, location_1, name_0, name_1, name_2, name_3, name_4, name_5, name_6)
 
### Name 5
n_5 <- n_6 %>% 
  filter(is.na(n_6$name_5)) %>% 
  select(id, location_1)

n_5 <- n_5 %>% 
  left_join(hierarchy_join, by = c("location_1" = "name_5"))
  
n_5$name_5 <- NA
n_5$name_5[!is.na(n_5$name_4)] <- n_5$location_1[!is.na(n_5$name_4)]

n_5_keep <- n_5 %>% 
  filter(!is.na(n_5$name_4)) %>% 
  select(id, location_1, name_0, name_1, name_2, name_3, name_4, name_5, name_6)

### Name 4
n_4 <- n_5 %>% 
  filter(is.na(n_5$name_4)) %>% 
  select(id, location_1)

n_4 <- n_4 %>% 
  left_join(hierarchy_join, by = c("location_1" = "name_4"))
  
n_4$name_4 <- NA
n_4$name_4[!is.na(n_4$name_3)] <- n_4$location_1[!is.na(n_4$name_3)]

n_4_keep <- n_4 %>% 
  filter(!is.na(n_4$name_3)) %>% 
  select(id, location_1, name_0, name_1, name_2, name_3, name_4, name_5, name_6)
  
### Name 3
n_3 <- n_4 %>% 
  filter(is.na(n_4$name_3)) %>% 
  select(id, location_1)

n_3 <- n_3 %>% 
  left_join(hierarchy_join, by = c("location_1" = "name_3"))
  
n_3$name_3 <- NA
n_3$name_3[!is.na(n_3$name_2)] <- n_3$location_1[!is.na(n_3$name_2)]

n_3_keep <- n_3 %>% 
  filter(!is.na(n_3$name_2)) %>% 
  select(id, location_1, name_0, name_1, name_2, name_3, name_4, name_5, name_6)

### Name 2
n_2 <- n_3 %>% 
  filter(is.na(n_3$name_2)) %>% 
  select(id, location_1)

n_2 <- n_2 %>% 
  left_join(hierarchy_join, by = c("location_1" = "name_2"))
  
n_2$name_2 <- NA
n_2$name_2[!is.na(n_2$name_1)] <- n_2$location_1[!is.na(n_2$name_1)]
  
n_2_keep <- n_2 %>% 
  filter(!is.na(n_2$name_1)) %>% 
  select(id, location_1, name_0, name_1, name_2, name_3, name_4, name_5, name_6)


### Name 1
n_1 <- n_2 %>% 
  filter(is.na(n_2$name_1)) %>% 
  select(id, location_1)

n_1 <- n_1 %>% 
  left_join(hierarchy_join, by = c("location_1" = "name_1"))
  

n_1$name_1 <-  NA
n_1$name_1[!is.na(n_1$name_0)] <- n_1$location_1[!is.na(n_1$name_0)]

n_1_keep <- n_1 %>% 
  filter(!is.na(n_1$name_0)) %>% 
  select(id, location_1, name_0, name_1, name_2, name_3, name_4, name_5, name_6)

### Name 0

n_0_keep <- n_1 %>% 
  filter(is.na(n_1$name_0)) %>% 
  select(id, location_1, name_0, name_1, name_2, name_3, name_4, name_5, name_6)

n_all <- rbind(n_6_keep,
               n_5_keep,
               n_4_keep,
               n_3_keep,
               n_2_keep,
               n_1_keep,
               n_0_keep)


```

Now we can add on the locations to the main contacts data. 

```{r merge_location}

contacts <- left_join(contacts, n_all)


head(contacts)

```





<!-- ====================================================== -->
## Adding new variables

In the following we add variables to the datasets:

* `cases`
    + `top_zones`: the 5 health zones having reported most cases over the last
      42 days
    * `epiweek_report_label`: the epiweek for the date of reporting
    * `epiweek_report` the first day of the epiweek for the date of reporting
    * `age_class`
    * `province`
    * `zone_de_sante`
    * `aire_de_sante`
    * `quartier`
    * `cellule`
    * `reco`
    
* `contacts`
    + `top_zones`: the 5 health zones having reported most cases over the last
      42 days
    * `age_class`
    * `province`
    * `zone_de_sante`
    * `aire_de_sante`
    * `quartier`
    * `cellule`
    * `reco`

* `followups`
    + `uid`: unique identifier for matching in `contacts$id` (long form)

```{r new_variables}

recent_date <- database_date - 42
recent_cases <- cases$date_of_reporting > recent_date
 
cases %>%
  mutate(top_zones = top_values(location_1, 5, subset = recent_cases),
       epiweek_report_label = aweek::date2week(date_of_reporting,
                                               week_start = "Monday",
                                               floor_day = TRUE),
       epiweek_report = aweek::week2date(epiweek_report_label,
                                         week_start = "Monday")) %>%
  mutate(
    age = as.numeric(age_years),
    age_class = factor(
        case_when(
            age <= 5 ~ "<=5",
            age <= 10 ~ "6-10",
            age <= 17 ~ "11-17",
            age <= 25 ~ "18-25",
            age <= 35 ~ "26-35",
            age <= 45 ~ "36-45",
            age <= 55 ~ "46-55",
            age <= 65 ~ "56-65",
            is.finite(age) ~ "66+",
            TRUE ~ "unknown"
        ), levels = c(
               "<=5",
               "6-10",
               "11-17",
               "18-25",
               "26-35",
               "36-45",
               "46-55",
               "56-65",
               "66+",
               "unknown"
           )),
    age_class_plot = factor(
        age_class,
        levels = rev(levels(age_class))))
  
```







<!-- ===================================================== -->
<!-- ===================================================== -->
<!-- ===================================================== -->
# Export data {.tabset .tabset-fade .tabset-pills}

<!-- ===================================================== -->
## Outline
In this part we export the clean datasets to `rds` files, and create a
`current_clean_data.R` script in `/scripts/` which will create object pointing
to the right files.




<!-- ===================================================== -->
## R objects

We export some of the clean database, placed in `produced_rds/` as well as in
`data/clean/`:

```{r export_rds}

## check if a directory exists and if not then creates it
if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}


## export cases file
cases_rds_file_name <- sprintf(
    "%sclean_%s.rds",
    undated_file_name(cases_file),
    format(database_date, "%Y-%m-%d"))

cases_rds_file_name
rio::export(cases,
            file.path("produced_rds", cases_rds_file_name))


## export relationships file
relationships_rds_file_name <- sprintf(
    "%sclean_%s.rds",
    undated_file_name(relationships_file),
    format(database_date, "%Y-%m-%d"))

relationships_rds_file_name
rio::export(relationships,
            file.path("produced_rds", relationships_rds_file_name))


## export contacts file
contacts_rds_file_name <- sprintf(
    "%sclean_%s.rds",
    undated_file_name(contacts_file),
    format(database_date, "%Y-%m-%d"))

contacts_rds_file_name
rio::export(contacts,
            file.path("produced_rds", contacts_rds_file_name))


## export followups file
followups_rds_file_name <- sprintf(
    "%sclean_%s.rds",
    undated_file_name(followups_file),
    format(database_date, "%Y-%m-%d"))

followups_rds_file_name
rio::export(followups,
            file.path("produced_rds", followups_rds_file_name))


```

We copy these files to the `data/clean` folder:

```{r copy_rds}

## loop over files to copy
files <- c(cases_rds_file_name,
           relationships_rds_file_name,
           contacts_rds_file_name,
           followups_rds_file_name)

for (e in files) {
  destination <- here("data",
                      "clean",
                      e)
  file.copy(from = file.path("produced_rds", e),
            to = destination,
            overwrite = TRUE)
}

```



<!-- ===================================================== -->
## Update the `current_clean_data.R` script

```{r update_script}

## path to the output file 
script_destination <- here::here("scripts",
                                 "current_clean_data.R")

## comments to say when this was updated
txt <- paste("## This file is generated automatically by `aaa_clean_linelist`",
             "## Do not edit it by hand!\n",
             sep = "\n")
cat(txt, file = script_destination, append = FALSE)

txt <- sprintf("\n## This file was last updated on the: %s\n",
               Sys.time())
cat(txt, file = script_destination, append = TRUE)


## add to script: cases file
txt <- sprintf('\ncurrent_clean_cases <- here::here("data",
                                  "clean",
                                  "%s")',
               cases_rds_file_name)
cat(txt, file = script_destination, append = TRUE)


## add to script: relationships file
txt <- sprintf('\ncurrent_clean_relationships <- here::here("data",
                                                "clean",
                                                "%s")',
               relationships_rds_file_name)
cat(txt, file = script_destination, append = TRUE)


## add to script: contacts file
txt <- sprintf('\ncurrent_clean_contacts <- here::here("data",
                                                "clean",
                                                "%s")',
               contacts_rds_file_name)
cat(txt, file = script_destination, append = TRUE)


## add to script: followups file
txt <- sprintf('\ncurrent_clean_followups <- here::here("data",
                                                "clean",
                                                "%s")',
               followups_rds_file_name)
cat(txt, file = script_destination, append = TRUE)


```





<!-- ===================================================== -->
<!-- ===================================================== -->
<!-- ===================================================== -->
# System information {.tabset .tabset-fade .tabset-pills}

The following information documents the system on which the document was
compiled.



<!-- ===================================================== -->
## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```



<!-- ===================================================== -->
## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```



<!-- ===================================================== -->
## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```



<!-- ============================================ -->
## Compilation parameters

The following parameters (`params`) have been used during compilation:

```{r params}
params
```


