---
title: "GoData: let's go clean"
author: "Thibaut Jombart and Sara Hollis"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: pygments
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 2
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---



<br>

<div class="report_meta">
  <span class="notice">**Notice**: this is a **stable, routine report**. 
  **Do not touch it unless it is broken.** To make a contribution, **carefully read 
  the [README](../../../../../README.html) file**.</span>
  
  **Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)
  
  **Code contributors:** Thibaut Jombart, Sara Hollis
  
  **Data contributors:** GoData team Beni
    
  **Reviewed by:** 
</div>





<!-- ====================================================== -->
<!-- ====================================================== -->
<!-- ====================================================== -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This is the universal cleaning script for raw data exported from godata.

### Data used

This report uses the latest exported .xlsx from godata.

### Method

The data preparation involves the following steps, detailed in the following
tabs:

* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Add new variables**: addition of new variables to the Master linelist
  database, like top affected health zones.

* **Filter the data**: keep only relevant confirmed and probable cases, possibly
  removing erroneous dates, for further analysis



<!-- ====================================================== -->
## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned MDC data stored as `x`

**Important**: we need to make sure the soucing of `R` scripts is done using the
current environment, using the argument `local = TRUE`. This is in particular
essential when `params` is used, as some functions and settings are dependent
upon it.

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names=TRUE)
for (file in scripts_files) source(file, local = TRUE)

ggthemr("grape")

## show parameters
params

```




<!-- ====================================================== -->
## Load data

We extract the completion date from the file names, for:

1. cases data

2. contacts data

3. relationships between cases and contacts

4. contact follow up data


```{r load_data}

## step 1
cases_file
cases <- custom_import(cases_file) %>%
  as_tibble()


## step 2
contacts_file
contacts <- custom_import(contacts_file) %>%
  as_tibble()


## step 3
relationships_file
relationships <- custom_import(relationships_file) %>%
  as_tibble()

## step 4
followups_file
followups <- custom_import(followups_file) %>%
  as_tibble()

```


<!-- ====================================================== -->
## Completion date

We extract the completion date from the name of the cases file.

```{r database_date}

## extract date of database based on cases file
cases_file_short <- gsub("^[^.]+/", "", cases_file)
database_date <- cases_file_short %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.





<!-- ====================================================== -->
## Data standardisation

Here we standardise data using `clean_data`, which set all data to lower case,
removes special characters, replaces non-ascii characters with their closest
ascii match.

```{r clean_data}

## standardise all data
cases <- clean_data(cases, guess_dates = FALSE)
contacts <- clean_data(contacts, guess_dates = FALSE)
relationships <- clean_data(relationships, guess_dates = FALSE)
followups <- clean_data(followups, guess_dates = FALSE)

```
  



<!-- ====================================================== -->
## Dates processing

Here we ensure that dates we will be using are indeed dates, and convert them
where necessary using `guess_dates`.

```{r clean_dates}

cases <- cases %>%
  mutate(date_of_reporting = guess_dates(date_of_reporting),
         date_of_onset = guess_dates(date_of_onset),
         date_of_outcome = guess_dates(date_of_outcome),
         date_of_infection = guess_dates(date_of_infection),
         date_of_becoming_case = guess_dates(date_of_becoming_case),
         date_of_burial = guess_dates(date_of_burial),
         date_of_case_notification = guess_dates(date_de_notification_du_cas))

contacts <- contacts %>%
  mutate(date_of_reporting = guess_dates(date_of_reporting),
         date_of_data_entry = guess_dates(created_at),
         date_of_becoming_case = guess_dates(date_of_becoming_case),
         date_of_onset = guess_dates(date_of_onset),
         date_of_followup_start = guess_dates(start_of_the_followup),
         date_of_followup_end = guess_dates(end_of_the_followup),
         date_of_last_contact = guess_dates(date_of_last_contact))

relationships <- relationships %>%
  mutate(date_of_data_entry = guess_dates(created_at),
         date_of_last_contact = guess_dates(date_of_last_contact))

followups <- followups %>%
  mutate(date_of_data_entry = guess_dates(created_at),
         date_of_followup = guess_dates(date),
         date_of_onset = guess_dates(dateapparitionsignes),
         date_of_hospitalization = guess_dates(dateadmissioncte),
         date_of_outcome = guess_dates(datededeces))
         date_of_burial = guess_dates(date_of_burial),
         date_of_case_notification = guess_dates(date_de_notification_du_cas))

```



<!-- ====================================================== -->
## Age variable standardization

Here we ensure that all age info is captured in one field `age_years` across all
datasets.

```{r clean_age}

cases <- cases %>%
  mutate(age_years = case_when(
                        is.na(age_years) ~ (age_months / 12),
                        TRUE ~ age_years))

contacts <- contacts %>%
  mutate(age_years = case_when(
                        is.na(age_years) ~ (age_months / 12),
                        TRUE ~ age_years))

```



<!-- ====================================================== -->
## Adding new variables

In the following we add variables to the datasets:

* `cases`
    + `top_zones`: the 5 health zones having reported most cases over the last
      42 days
    * `epiweek_report_label`: the epiweek for the date of reporting
    * `epiweek_report` the first day of the epiweek for the date of reporting
    * `age_class`

```{r new_variables}

recent_date <- database_date - 42
recent_cases <- cases$date_of_reporting >= recent_date
 
cases <- cases %>%
  mutate(top_zones = top_values(location_1, 5, subset = recent_cases),
         epiweek_report_label = aweek::date2week(date_of_reporting,
                                                 week_start = "Monday", 
                                                 floor_day = TRUE),
         epiweek_report = aweek::week2date(epiweek_report_label,
                                           week_start = "Monday")) %>%
  mutate(
      age = as.numeric(age_years),
      age_class = factor(
          case_when(
              age <= 5 ~ "<=5",
              age <= 10 ~ "6-10",
              age <= 17 ~ "11-17",
              age <= 25 ~ "18-25",
              age <= 35 ~ "26-35",
              age <= 45 ~ "36-45",
              age <= 55 ~ "46-55",
              age <= 65 ~ "56-65",
              is.finite(age) ~ "66+",
              TRUE ~ "unknown"
          ), levels = c(
                 "<=5",
                 "6-10",
                 "11-17",
                 "18-25",
                 "26-35",
                 "36-45",
                 "46-55",
                 "56-65",
                 "66+",
                 "unknown"
             )),
      age_class_plot = factor(
          age_class,
          levels = rev(levels(age_class))))

```





<!-- ===================================================== -->
<!-- ===================================================== -->
<!-- ===================================================== -->
# Export data {.tabset .tabset-fade .tabset-pills}

<!-- ===================================================== -->
## Outline
In this part we export the clean datasets to `rds` files, and create a
`current_clean_data.R` script in `/scripts/` which will create object pointing
to the right files.




<!-- ===================================================== -->
## R objects

We export some of the clean database, placed in `produced_rds/` as well as in
`data/clean/`:

```{r export_rds}

## check if a directory exists and if not then creates it
if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}


## export cases file
cases_rds_file_name <- sprintf(
    "%sclean_%s.rds",
    undated_file_name(cases_file),
    format(database_date, "%Y-%m-%d"))

cases_rds_file_name
rio::export(cases,
            file.path("produced_rds", cases_rds_file_name))


## export relationships file
relationships_rds_file_name <- sprintf(
    "%sclean_%s.rds",
    undated_file_name(relationships_file),
    format(database_date, "%Y-%m-%d"))

relationships_rds_file_name
rio::export(relationships,
            file.path("produced_rds", relationships_rds_file_name))


## export contacts file
contacts_rds_file_name <- sprintf(
    "%sclean_%s.rds",
    undated_file_name(contacts_file),
    format(database_date, "%Y-%m-%d"))

contacts_rds_file_name
rio::export(contacts,
            file.path("produced_rds", contacts_rds_file_name))


## export followups file
followups_rds_file_name <- sprintf(
    "%sclean_%s.rds",
    undated_file_name(followups_file),
    format(database_date, "%Y-%m-%d"))

followups_rds_file_name
rio::export(followups,
            file.path("produced_rds", followups_rds_file_name))


```

We copy these files to the `data/clean` folder:

```{r copy_rds}

## loop over files to copy
files <- c(cases_rds_file_name,
           relationships_rds_file_name,
           contacts_rds_file_name,
           followups_rds_file_name)

for (e in files) {
  destination <- here("data",
                      "clean",
                      e)
  file.copy(from = file.path("produced_rds", e),
            to = destination,
            overwrite = TRUE)
}

```



<!-- ===================================================== -->
## Update the `current_clean_data.R` script

```{r update_script}

## path to the output file 
script_destination <- here::here("scripts",
                                 "current_clean_data.R")

## comments to say when this was updated
txt <- paste("## This file is generated automatically by `aaa_clean_linelist`",
             "## Do not edit it by hand!\n",
             sep = "\n")
cat(txt, file = script_destination, append = FALSE)

txt <- sprintf("\n## This file was last updated on the: %s\n",
               Sys.time())
cat(txt, file = script_destination, append = TRUE)


## add to script: cases file
txt <- sprintf('\ncurrent_clean_cases <- here::here("data",
                                  "clean",
                                  "%s")',
               cases_rds_file_name)
cat(txt, file = script_destination, append = TRUE)


## add to script: relationships file
txt <- sprintf('\ncurrent_clean_relationships <- here::here("data",
                                                "clean",
                                                "%s")',
               relationships_rds_file_name)
cat(txt, file = script_destination, append = TRUE)


## add to script: contacts file
txt <- sprintf('\ncurrent_clean_contacts <- here::here("data",
                                                "clean",
                                                "%s")',
               contacts_rds_file_name)
cat(txt, file = script_destination, append = TRUE)


## add to script: followups file
txt <- sprintf('\ncurrent_clean_followups <- here::here("data",
                                                "clean",
                                                "%s")',
               followups_rds_file_name)
cat(txt, file = script_destination, append = TRUE)


```





<!-- ===================================================== -->
<!-- ===================================================== -->
<!-- ===================================================== -->
# System information {.tabset .tabset-fade .tabset-pills}

The following information documents the system on which the document was
compiled.



<!-- ===================================================== -->
## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```



<!-- ===================================================== -->
## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```



<!-- ===================================================== -->
## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```



<!-- ============================================ -->
## Compilation parameters

The following parameters (`params`) have been used during compilation:

```{r params}
params
```


