---
title: "Epicurves"
author: "Thibaut Jombart and Sara Hollis"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: pygments
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 2
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---



<br>

<div class="report_meta">
  <span class="notice">**Notice**: this is a **stable, routine report**. 
  **Do not touch it unless it is broken.** To make a contribution, **carefully read 
  the [README](../../../../../README.html) file**.</span>
  
  **Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)
  
  **Code contributors:** Sara Hollis, Thibaut Jombart, Amy Gimma
  
  **Data contributors:** Yannick Tutu, Richy Ngombo
  
  **Version:** 1.0.0
  
  **Reviewed by:** 
</div>





<!-- ====================================================== -->
<!-- ====================================================== -->
<!-- ====================================================== -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This report produces basic epicurves by dates of onset or reporting, with
various stratification, from the cleaned Master linelist database.

### Data used

This report uses the latest cleaned Master linelist data.

### Method

The data preparation involves the following steps, detailed in the following
tabs:

* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Add new variables**: addition of new variables to the Master linelist database, like top
  affected health zones.

* **Filter the data**: keep only relevant confirmed and probable cases, possibly
  removing erroneous dates, for further analysis



<!-- ====================================================== -->
## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned MDC data stored as `x`

**Important**: we need to make sure the soucing of `R` scripts is done using the
current environment, using the argument `local = TRUE`. This is in particular
essential when `params` is used, as some functions and settings are dependent
upon it.

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names=TRUE)
for (file in scripts_files) source(file, local = TRUE)

ggthemr("grape")

## show parameters
params

```




<!-- ====================================================== -->
## Load data

We load the current clean cases, whose path is defined in `scripts/current_clean_data.R`

```{r load_data}

## step 1
current_clean_cases
x <- rio::import(current_clean_cases) %>%
  as_tibble()


```


<!-- ====================================================== -->
## Completion date

We extract the completion date from the name of the cases file.

```{r database_date}

## extract date of database based on cases file
cases_file_short <- gsub("^[^.]+/", "", current_clean_cases)
database_date <- cases_file_short %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.


  



<!-- ====================================================== -->
## Filter data

In this section we subset the data so that:

* `cases` contains only confirmed and probable cases

```{r filter}

cases <- x %>%
  filter(classification %in% c("confirmed", "probable"))

```





<!-- ====================================================== -->
## Customised color scales

We define custom colors to be used in graphs.

```{r colors}

scale_classification <- scale_fill_manual(
    "",
    values = c(confirmed = "#C83D52",
               probable = "#D78E99"),
    labels = c(confirmed = "Confirmé",
               probable = "Probable"))

```






<!-- ===================================================== -->
<!-- ===================================================== -->
<!-- ===================================================== -->
# Epidemic curves {.tabset .tabset-fade .tabset-pills}

<!-- ===================================================== -->
## Outline

This section analyses the `cases` database. We derive overall epicurves and
stratified by:

* confirmed/probable (`classification` variable)

* health zone (`location_1` variable)




<!-- ===================================================== -->
## Overall cases

```{r epicurve_overall}

i_report_classification <- incidence(cases$date_of_reporting, 7, groups = cases$classification)

plot(i_report_classification, border = "white") +
  theme_bw() +
  rotate_x_text(45) +
  scale_classification +
  large_txt +
  theme(legend.title = element_blank(),
        legend.position = c(0, 1),
        legend.justification = c(-0.05, 1.1),
        legend.background = element_rect(fill = "transparent", 
                                         color = "white")) +
  labs(title = paste0("Courbe épidemique: cas confirmés et probables",
                     "\npar date de début de symptômes à la date de", " ", 
                     format(database_date, "%d/%m/%Y")),
       y = "Nombre de nouveaux cas par semaine") +
  scale_months

```

We provide the corresponding numbers in a table:


```{r table_epicurve_overall}

table_epicurve_overall <- i_report_classification %>%
  as.data.frame() %>%
  adorn_totals(c("row", "col"))

table_epicurve_overall %>%
  show_table()

```




<!-- ===================================================== -->
## Stratified by health zone

```{r epicurve_health_zone}

cases %>%
  filter(top_zones != "other") %>% 
  ggplot(aes(x = epiweek_report, fill = classification)) +
  geom_bar(alpha = .7) +
  theme_bw() +
  scale_classification +
  rotate_x_text(45) +
  large_txt +
  facet_grid(top_zones ~ .) +
  theme(legend.position = "bottom") +
   labs(title = paste0("Courbe épidemique: cas confirmés et probables",
                      "\npar semaine de notification, à la date de", " ", 
                      format(database_date, "%d/%m/%Y")),
        y = "Nombre de nouveaux cas par semaine",
        x = "") +
  custom_vert_facet +
  scale_months

```

We provide the corresponding numbers in a table:

```{r table_epicurve_health_zone}

table_epicurve_health_zone <- cases %>%
  count(epiweek_report, top_zones, classification) %>%
  spread(classification, n, fill = 0) %>%
  adorn_totals(c("row", "col"))

table_epicurve_health_zone %>%
  show_table()

```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Export tables {.tabset .tabset-fade .tabset-pills}

## Outline

Provide an outline of exported files.


## Excel files

```{r xlsx_exports}

if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

to_export <- c("table_epicurve_overall",
               "table_epicurve_health_zone")

for (e in to_export) {
  rio::export(get(e),
              file.path("produced_xlsx",
                        paste0(e, ".xlsx")))
}

```

Click on the following links to open the files (only works if the files above
have been generated and are in the same folder as this document):


```{r links, results = "asis"}

for (e in to_export) {
  txt <- sprintf("- [%s.xlsx](%s.xlsx)",
                 e,
                 file.path("produced_xlsx",
                           e))
  cat(txt, sep = "\n")
}

```







<!-- ===================================================== -->
<!-- ===================================================== -->
<!-- ===================================================== -->
# System information {.tabset .tabset-fade .tabset-pills}

The following information documents the system on which the document was
compiled.

## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```


### Change log:

#### Version 2.1.0

* added new variable names
* set color scale for zone de sante across document
* fixed spelling mistakes
* changed table and figure names as appropriate ("last 21 days" to "recent onset")
* added figures and tables for top recent onset ZS, daily for the last 3 months
* exported new table

### To do: 
* check use of epiweek_onset_label (lolwa?)
* rename variables (last_21_days)
* describe 21 (or other timeframe) days more clearly in text
* add table and fig for "top ZS recent report" data
* decide on "recent" timeframe
* decide on plot timeline in days for daily plots (63 days?)
* review with Thibaut
* decide maintainer
