---
title: "Contact followup"
author: "Thibaut Jombart and Sara Hollis"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: pygments
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 2
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---



<br>

<div class="report_meta">
  <span class="notice">**Notice**: this is a **stable, routine report**. 
  **Do not touch it unless it is broken.** To make a contribution, **carefully read 
  the [README](../../../../../README.html) file**.</span>
  
  **Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)
  
  **Code contributors:** Sara Hollis, Thibaut Jombart, Amy Gimma
  
  **Data contributors:** Yannick Tutu, Richy Ngombo
  
  **Version:** 1.0.0
  
  **Reviewed by:** 
</div>





<!-- ====================================================== -->
<!-- ====================================================== -->
<!-- ====================================================== -->
# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This report produces basic epicurves by dates of onset or reporting, with
various stratification, from the cleaned Master linelist database.

### Data used

This report uses the latest cleaned Master linelist data.

### Method

The data preparation involves the following steps, detailed in the following
tabs:

* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Add new variables**: addition of new variables to the Master linelist database, like top
  affected health zones.

* **Filter the data**: keep only relevant confirmed and probable cases, possibly
  removing erroneous dates, for further analysis



<!-- ====================================================== -->
## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned MDC data stored as `x`

**Important**: we need to make sure the soucing of `R` scripts is done using the
current environment, using the argument `local = TRUE`. This is in particular
essential when `params` is used, as some functions and settings are dependent
upon it.

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names=TRUE)
for (file in scripts_files) source(file, local = TRUE)

ggthemr("grape")

## show parameters
params

```




<!-- ====================================================== -->
## Load data

We load the current *contacts* and *followup* data, whose path is defined in
`scripts/current_clean_data.R`

```{r load_data}

current_clean_contacts
contacts <- rio::import(current_clean_contacts) %>%
  as_tibble()

current_clean_followups
followups <- rio::import(current_clean_followups) %>%
  as_tibble()


```



<!-- ====================================================== -->
## Completion date

We extract the completion date from the name of the cases file.

```{r database_date}

## extract date of database based on cases file
file_short <- gsub("^[^.]+/", "", current_clean_contacts)
database_date <- file_short %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.




<!-- ====================================================== -->
## Quality controls

We perform the following checks on the data:

* check that contacts in `followups` are also documented in `contacts`; those
  who are not are output to a table
  
```{r quality_controls}

table_missing_contacts <- followups %>%
  filter(!(lng_follow_up_field_label_contact_id %in% contacts$id))

table_missing_contacts %>%
  show_table()

```




## Adding variables

We add to `followup` the following variables:

* `seen`: a logical indicating if patient was seen
* `date_seen`: the date of the followup, if seen; `NA` otherwise

```{r add_variables}

followups <- followups %>%
  mutate(seen = status %in% c("seen_no_signs", "seen_with_signs"),
         date_seen = if_else(seen, date_of_followup, as.Date(NA)))

```


  
<!-- ====================================================== -->
## Filtering

We subset the rows of `followups` so that:

1. `status` is either: *not_performed*, *not_seen*, *seen_no_signs*,
*seen_with_signs*

2. contacts are identified in the `contacts` database 


```{r filter}

## step 1 and 2
status_to_keep <- c("not_performed",
                    "not_seen",
                    "seen_no_signs",
                    "seen_with_signs")
followups <- followups %>%
  filter(status %in% status_to_keep,
         lng_follow_up_field_label_contact_id %in% contacts$id)

```




<!-- ====================================================== -->
## Join *followups* and *contacts*

We add to the `followups` data some information drawn from the `contacts` data,
including:

* the date at which followup should start (`date_of_followup_start`)
* the date at which followup should end (`date_of_followup_end`)

```{r join}

info_for_merging <- contacts %>%
  select(id, date_of_last_contact, date_of_followup_start) %>%
  mutate(date_of_followup_end = date_of_followup_start + 20)

followups <- followups %>%
  right_join(info_for_merging,
             by = c("uid" = "id"))

```




<!-- ====================================================== -->
## Customised color scales

We define custom colors to be used in graphs.

```{r colors}

scale_status <- scale_fill_manual(
    "",
    values = c(not_performed = "#938083",
               not_seen = "#C23F54",
               seen_no_signs = "#539998",
               seen_with_signs = "#7E000B"),
    labels = c(not_performed = "Non visité",
               not_seen = "Non vu",
               seen_no_signs = "Vu, pas de symptômes",
               seen_with_signs = "Vu, symptômes"))

scale_classification <- scale_fill_manual(
    "",
    values = c(inconnu = "#8c8c8c",
               vu = "#79d2a6",
               non_vu = "#9494b8",
               perdu_de_vue = "#ff9966",
               jamais_vu_court = "#ff6666",
               jamais_vu_long = "#b30000",
               pas_encore_vu = "#5E9281"),
    labels = c(inconnu = "Inconnu",
               vu = "Vu",
               non_vu = "Non vu",
               perdu_de_vue = "Perdu de vue",
               jamais_vu_court = "Jamais vu (court)",
               jamais_vu_long = "Jamais vu (long)",
               pas_encore_vu = "Pas encore vu"))

```






<!-- ===================================================== -->
<!-- ===================================================== -->
<!-- ===================================================== -->
# Follow-up completion {.tabset .tabset-fade .tabset-pills}

<!-- ===================================================== -->
## Outline

* numbers of contacts by followup status over time

* timeliness of contact listing


## Contact followup over time

```{r followup_time}

ggplot(followups, aes(x = date_of_followup, fill = status)) +
  geom_bar() +
  theme_bw() +
  large_txt +
  scale_status +
  scale_weeks +
  rotate_x_text(45) +
  labs(title = "Suivi des contacts par jour",
       x = "",
       y = "Nombre de contacts suivis") +
  theme(legend.position = "bottom")

table_followup_time <- followups %>%
  count(date_of_followup, status) %>%
  spread(status, n, fill = 0) %>%
  adorn_totals(c("row", "col"))

table_followup_time %>%
  show_table()

```



<!-- ======================================================= -->
## Contact classification: today

In this part we get contact classification for the current day of the database,
i.e. **`r format(database_date, format = "%d %B %Y")`**.


```{r classification_current}

classification_current <- classify_contacts(contacts, followups, on = database_date)
classification_current %>%
  count(classification)

classification_current %>%
  show_table()

```




<!-- ======================================================= -->
## Contact classification over time

Case classification includes the following categories, all with respect to a
given date *d*:

* **`vu`**: seen on day *d*
* **`non_vu`**: last seen 1 or 2 days before *d*
* **`perdu_de_vue`**: last seen 3+ days before *d*
* **`pas_encore_vu`**: will be seen in the future, but have not been seen yet
* **`jamais_vu_court`**: never seen, for 1-7 days since last exposure
* **`jamais_vu_long`**: never seen, for 8-21 days since last exposure

We derive contact classification for the 42 days leading up to the current
database date, i.e. from the  **`r format(database_date - 41, format = "%d %B %Y")`** to the  **`r format(database_date, format = "%d %B %Y")`**..

```{r classification_time}

starting_date <- database_date - 41
dates <- seq(starting_date, database_date, by = 1L)

classification_all_dates <- lapply(dates,
                                   function(date)
                                     classify_contacts(
                                     contacts,
                                     followups,
                                     date))

classification_time <- Reduce(rbind, classification_all_dates)
classification_time <- classification_time %>%
  mutate(date =
           rep(dates, sapply(classification_all_dates, nrow)))
                               
    
ggplot(classification_time, aes(x = date, fill = classification)) +
  geom_bar() +
  theme_bw() +
  scale_classification +
  large_txt +
  scale_weeks +
  theme(legend.position = "bottom") +
  rotate_x_text(45) +
  labs(title = "Classification des contacts au cours du temps",
       x = "",
       y = "Nombre de contacts par jour")

table_classification_time <- classification_time %>%
  count(date, classification) %>%
  spread(classification, n, fill = 0) %>%
  adorn_totals(c("row", "col"))

table_classification_time%>%
  show_table()

```




<!-- ======================================================= -->
## Delays to follow-up

```{r delays_to_followup}

info_seen_dates <- select(classification_time,
                          id,
                          first_seen = first_seen_global,
                          last_seen = last_seen_global)

contacts <- left_join(contacts, info_seen_dates)

contacts <- contacts %>%
  mutate(delay_to_followup = as.integer(first_seen - date_of_last_contact - 1),
         followup_duration = as.integer(last_seen - first_seen + 1))

contacts %>%
  filter(date_of_last_contact >= starting_date) %>% 
ggplot(aes(x = delay_to_followup)) +
  geom_bar() +
  theme_bw() +
  large_txt +
  labs(title = "Delais au suivi des contacts (21 derniers jours)",
       y = "Nombre de contacts",
       x = "Nombre de jours avant le premier suivi")

## summary of the delay to follow-up
delay_followup <- contacts$delay_to_followup %>%
  na.omit() %>%
  as.integer()
delay_followup %>%
  summary()

```

The **average** delay to follow-up is **`r round(mean(delay_followup), 1)`
days**, with an IQR of **`r round(mean(delay_followup), 1)` days**.






<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Export tables {.tabset .tabset-fade .tabset-pills}

## Outline

Provide an outline of exported files.


## Excel files

```{r xlsx_exports, eval = FALSE}

if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

to_export <- c("table_",
               "table_")

for (e in to_export) {
  rio::export(get(e),
              file.path("produced_xlsx",
                        paste0(e, ".xlsx")))
}

```

Click on the following links to open the files (only works if the files above
have been generated and are in the same folder as this document):


```{r links, results = "asis", eval = FALSE}

for (e in to_export) {
  txt <- sprintf("- [%s.xlsx](%s.xlsx)",
                 e,
                 file.path("produced_xlsx",
                           e))
  cat(txt, sep = "\n")
}

```







<!-- ===================================================== -->
<!-- ===================================================== -->
<!-- ===================================================== -->
# System information {.tabset .tabset-fade .tabset-pills}

The following information documents the system on which the document was
compiled.



<!-- ============================================ -->
## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```



<!-- ============================================ -->
## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```



<!-- ============================================ -->
## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```




<!-- ============================================ -->
## Compilation parameters

The following parameters (`params`) have been used during compilation:

```{r params}
params
```


