---
title: "Analyses of transmission chains"
author: "Thibaut, Maddy, Sam, Luc, Sonia, Elias, Xavier, the Butembo team, the analytic cell Goma"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 4
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE)
```



<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This report focusses on transmission chains of the outbreak, derived from the
narratives. This work was pioneered by the Butembo team, who assembled specific
linelists with transmission information. To simplify this workflow for other
teams (with less data management resources), we streamline this process as
follows:

1. information on the cases are taken from the clean linelist data (the *master
   linelist*), thus not requiring additional work from the field teams
   
2. information on transmissions is stored in a separate file following the same
   model as the master linelist, i.e. a google spreadsheet which can be edited
   by different data managers; this file only contains information on the
   *transmissions*, not on the cases, e.g.: from whom, to whom, which kind of
   transmission context etc.
   
The two sources of data will be merged using *epicontacts*.



### Data used

What is the source of your data? Does it need running some previous reports like
preliminary data cleaning? If so, list these reports here.

### Method

The data preparation involves the following steps, detailed in the following tabs:


* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Clean data**: this section contains *ad hoc* data cleaning, i.e. which is
  not used in other reports (otherwise cleaning should be done in a dedicated
  report); this section is also used to create new variables used in the
  analyses




## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned VHF data stored as `x`

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$",
                     full.names = TRUE)
for (file in scripts_files) source(file, local = TRUE)

```



## Load data

Two types of data are used in this report:

1. the clean *master linelist*, called `linelist` in this report
2. the raw *master transmission list*, called `chains` in this report

```{r load_data}

## load the master linelist ('MLL')
current_linelist
linelist <- rio::import(current_linelist) %>%
    as_tibble()
linelist

## load the chains
current_chains
chains_raw <- custom_import(current_chains) %>%
    as_tibble()
chains_raw

## extract completion dates
linelist_file_name <- gsub("^[^.]+/", "", current_linelist)
linelist_date <- linelist_file_name %>%
  guess_dates()
linelist_date

chains_file_name <- gsub("^[^.]+/", "", current_chains)
chains_date <- chains_file_name %>%
  guess_dates()
chains_date

```

The **completion dates** of the databases are:

* master linelist: **`r format(linelist_date, format = "%A %d %b %Y")`**
* master transmission: **`r format(chains_date, format = "%A %d %b %Y")`**





## Clean data

As we use a clean master linelist, cleaning only applies here to the *master
transmission list* stored in the object called `chains_raw`. Once cleaned,
transmission data will be stored as `chains`.

We use *linelist*'s function `clean_data()` to:

- remove all special characters from the data
- set all characters to lower case
- replace all accentuated and diacritic characters with their closest ascii
  match in the latin alphabet
- replace all separators with a single `_`
- replace all mis-spelling using a overall dictionary (see the file
  `cleaning_rules.xlsx` in `/dictionary/`, or type `cleaning_rules` in this R
  session, after compiling the document
- (optionally) detect date formats and convert data to dates, including cases
  where format varies within a column
  

This cleaning is achieved with:

```{r data_cleaning}

chains <- chains_raw %>%
  clean_data(guess_dates = FALSE) %>%
  as_tibble()

chains

```




## Cleaning using dictionary

Here we correct typos etc using our dictionary. We then tabulate the categories
of key variables to make sure nothing has been left out. New items to clean
should be added to the dictionary in `dictionary/cleaning_rules.xlsx`.

```{r clean_dic}

## clean using dictionary
chains <- chains %>%
  clean_variable_spelling(wordlists = cleaning_rules)

## check variables
table(chains$type_contamination, useNA = "ifany")
table(chains$lien_cas_source, useNA = "ifany")

```







<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Making an *epicontacts* object {.tabset .tabset-fade .tabset-pills}

## Outline 

In this section we build an *epicontacts* object and provide some simple
descriptive statistics and graphics.


## Use the devel version!

The following parts need the current development version of *epicontacts*. You
will need `devtools` to install it directly from github, by typing::

```{r install_epicontacts, eval = FALSE}

devtools::install_github("reconhub/epicontacts")

```

Note that you need to do this only once! As an alternative, use a recent
deployer: it contains the right version of the package.




## Building the epicontacts

We build the *epicontacts* object using:

* the *master linelist* stored in `linelist`, keeping only selected variables
  useful for further analyses, and entries for which we are missing VHF
  identifiers (`vhf_code`), as well as duplicates
  
* the *master transmission list* data stored in `chains`

```{r epicontacts}

select_linelist <- linelist %>%
  rename(mll_id = id,
         vhf_id = vhf_code) %>% 
  select("vhf_id",
         "mll_id",
         "epicasedef",
         "date_onset",
         "date_report",
         "date_death",
         "date_admission",
         "date_exposure_start",
         "date_exposure_end",
         "zone_de_sante",
         "aire_de_sante",
         "age",
         "age_class",
         "gender",
         "hcw",
         "firstname",
         "surname",
         "current_status",
         "report_status",
         "outcome",
         "hosp1_name") %>%
  filter(!is.na(vhf_id),
         !duplicated(vhf_id))

x <- make_epicontacts(linelist = select_linelist,
                      contacts = chains,
                      directed = TRUE,
                      id = "vhf_id",
                      from = "source_case_id",
                      to = "case_id")
x

```



## Summary

This provides an overview of the content of the *epicontacts* object. Pay
attention, in particular, to cases that are in the transmission tree (called
"contacts" in *epicontacts* terminology) but not in the linelist. These could be
recovered by using a larger database for linelist, such as VHF or the new MLL
(Master Line List) currated in Goma.

```{r summary}
summary(x)
```



## Visualisations

This is a general visusalisation of the whole dataset:

```{r plot_chains}

plot_x <- plot(x,
               node_color = "zone_de_sante",
               node_shape = "gender",
               shapes = c(female = "female",
                          male = "male",
                          unknown = "question-circle"),
               legend_max = 100,
               edge_color = "type_contamination"
               )
plot_x

```




## Active chains on last 21 days

We define as active chains, on a given limit date: 

1. cases that were isolated after the limit date, set up to be 21 days prior to
   the completion date of the database (i.e. here, **`r format(chains_date - 21,
   format = "%A %d %b %Y")`**)
   
2. all the cases of the transmission chains linked to 1)

Graphs will be shown using the date of isolation (`date_admission`) if no value
is missing; otherwise the default plotting of transmissions will be used.

```{r plot_chains_time_window}

## identify transmission chains ("clusters" or "connected components" in graph
## terminology)
x <- get_clusters(x, member_col = "cluster_id", size_col = "chain_size")

## force character type
x$linelist$cluster_id <- paste("chain", x$linelist$cluster_id, sep = "_")


## find source case of the transmission chains; these are cases who have no edge
## coming in, and at least one coming out

in_degree <- data.frame(in_degree = get_degree(x, "in")) %>%
  rownames_to_column(var = "id")
out_degree <- data.frame(out_degree = get_degree(x, "out")) %>%
  rownames_to_column(var = "id")

x$linelist <- x$linelist %>%
  left_join(in_degree) %>%
  left_join(out_degree) %>%
  mutate(is_root = (in_degree == 0) & (out_degree > 0),
         root = if_else(is_root, id, NULL))

## find roots of each transmission chain
roots <- split(x$linelist$root, x$linelist$cluster_id) %>%
  lapply(function(e) e[!is.na(e)]) %>%
  stack() %>%
  setNames(c("root", "chain_id")) %>%
  as_tibble() %>% 
  select("chain_id", "root")

roots

## add this information to the linelist as `chain_id`
x$linelist <- x$linelist %>%
  mutate(chain_id = clean_spelling(cluster_id, wordlist = roots))


## set limit date
date_limit <- chains_date - 42

## find transmission chains whose latest case is after the limit date
chains_to_keep <- x$linelist %>%
  group_by(chain_id) %>%
  summarise(most_recent = max(date_admission, na.rm = TRUE)) %>%
  filter(most_recent > date_limit) %>%
  pull(chain_id)

x_active <- subset(x,
                   node_attribute = list(chain_id = chains_to_keep)) %>%
  thin("contacts")

## check if some date_admission are missing
missing_date <- get_pairwise(x_active, "date_admission") %>% is.na()

if (!any(missing_date)) {
  x_axis_info <- "date_admission"
} else {
  x_axis_info <- NULL
}

## make the plot
plot_x_active <- plot(x_active,
                      node_color = "chain_id",
                      node_shape = "gender",
                      shapes = c(female = "female",
                                 male = "male",
                                 unknown = "question-circle"),
                      legend_max = 100,
                      edge_color = "type_contamination",
                      x_axis = x_axis_info,
                      edge_width = 5
                      )
plot_x_active


```








<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Descriptive analyses {.tabset .tabset-fade .tabset-pills}

## Outline

In the following we provide some simple descriptions of the transmission
network. We focus on spatial patterns in transmissions, comparing the locations
of the sources cases and their secondary cases, and the relative frequencies of
these transmissions.




## By sex

```{r contacts_sex}

pairs_sex <- get_pairwise(x,
                          "gender",
                          function(from, to)
                            as.data.frame(table(from, to)))

ggplot(pairs_sex,
       aes(y = from, x = to, size = Freq)) +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 20),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(45) + 
  labs(title = "Transmissions between sex",
       y = "Source case",
       x = "Infectee")

```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_sex_circi}

## Chord diagram
circle_plot(pairs_sex, label_space = 2)

```




## By age class

```{r contacts_age}

pairs_age <- get_pairwise(x,
                          "age_class",
                          function(from, to)
                            as.data.frame(table(from, to)))

ggplot(pairs_age,
       aes(y = from, x = to, size = Freq)) +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 20),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(45) + 
  labs(title = "Transmissions between sex",
       y = "Source case",
       x = "Infectee")

```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_age_circi}

## Chord diagram
circle_plot(pairs_age, label_space = 2)

```
 
## By health zones

We plot transmissions by health zones:

```{r contacts_zone_de_sante}

pairs_zone_de_sante <- get_pairwise(x,
                                    "zone_de_sante",
                                    function(from, to)
                                      as.data.frame(table(from, to)))

ggplot(pairs_zone_de_sante,
       aes(y = from, x = to, size = Freq)) +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 20),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(45) + 
  labs(title = "Transmissions across health zones",
       y = "Source case",
       x = "Infectee")

```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_zone_de_sante_circi}

## Chord diagram
circle_plot(pairs_zone_de_sante, label_space = 2)

```

We compute the proportion of transmission happening across different health
zones:

```{r move_zones}

move_zone <- get_pairwise(x,
                          "zone_de_sante",
                          function(from, to)
                            as.logical(na.omit(from != to)))
move_zone

## mean and 95% CI
mean(move_zone)
prop_ci(k = sum(move_zone), n = length(move_zone))

```




## By health areas

We plot transmissions by health areas:

```{r contacts_aire_de_sante, fig.width = 12}

pairs_aire_de_sante <- get_pairwise(x,
                                    "aire_de_sante",
                                    function(from, to)
                                      as.data.frame(table(from, to)))

ggplot(pairs_aire_de_sante,
       aes(y = from, x = to, size = Freq)) +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 10),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(90) + 
  labs(title = "Transmissions across health areas",
       y = "Source case",
       x = "Infectee")


```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_aire_de_sante_circi}

## Chord diagram
circle_plot(pairs_aire_de_sante, label_space = 2)

```

We compute the proportion of transmission happening across different health
areas:

```{r move_areas}

move_area <- get_pairwise(x,
                          "aire_de_sante",
                          function(from, to)
                            as.logical(na.omit(from != to)))
move_area

## mean and 95% CI
mean(move_area)
prop_ci(k = sum(move_area), n = length(move_area))

```




## Sources and recipients of transmissions

In this section we characterise health areas in terms of the number of
transmissions given ("*sources*"), or received ("*recipients*").

```{r sources_recipients, fig.width = 10}

## count cases sourced by area
df_sources <- pairs_aire_de_sante %>%
  filter(as.character(from) != as.character(to), Freq > 0) %>%
  group_by(from) %>%
  summarise(n_sourced = sum(Freq)) %>%
  rename(aire_de_sante = from)

## count cases received by area
df_receivers <- pairs_aire_de_sante %>%
  filter(as.character(from) != as.character(to), Freq > 0) %>%
  group_by(to) %>%
  summarise(n_received = sum(Freq)) %>%
  rename(aire_de_sante = to)

## put the information together
df_sources_recipients <- full_join(df_sources, df_receivers) %>%
  filter(aire_de_sante != "inconnu") %>% 
  replace_na(list(n_sourced = 0, n_received = 0)) %>%
  arrange(desc(n_sourced), desc(n_received))

## add info on zones
zone_info <- linelist %>%
  group_by(zone_de_sante, aire_de_sante) %>%
  count() %>%
  select(1:2)
df_sources_recipients <- right_join(zone_info, df_sources_recipients)

## display the table
df_sources_recipients %>%
  show_table()

## make the plot
ggplot(df_sources_recipients,
       aes(x = n_received,
           y = n_sourced,
           color = aire_de_sante)) +
  geom_point() +
  geom_label_repel(aes(label = aire_de_sante), size = 3) +
  scale_x_continuous(breaks = int_breaks) +
  scale_y_continuous(breaks = int_breaks) +
  large_txt +
  facet_wrap(~ zone_de_sante) +
  scale_color_discrete(guide = FALSE) +
  theme_bw() +
  labs(title = "Sources and recipients of cases",
       x = "Number of cases received",
       y = "Number of cases sourced")

```









<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Analysis of transmissibility {.tabset .tabset-fade .tabset-pills}

## Reproduction number

We can assess the effective reproduction number ($R$) by looking at the numbers
of secondary cases by index case:

```{r R}

## in-degree
x %>% get_degree() %>% table()

## out-degree = reproduction number
x %>% get_degree("out") %>% table()

R_df <- x %>%
  get_degree("out") %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  rename("id" = "rowname", "R" = ".")

head(R_df)

```

We create a new `data.frame` with R values and case info from the linelist;
unlike the linelist, where each line is a case, here each line is a source case:

```{r R_df}

R_df <- left_join(R_df, x$linelist, by = "id") %>%
  as_tibble()

R_df %>%
  show_table()

```

Because of delays from primary to secondary onset, and from onset to
notification, we need to ignore the latest source cases to avoid
under-estimating the $R$. We use a conservative limit of 45 days:

```{r R_plots}

old_cases <- R_df %>%
  filter(date_onset < (chains_date - 45))

ggplot(old_cases, aes(x = R)) +
  geom_bar() +
  large_txt +
  labs(title = "Number of secondary cases by index case",
       x = "Observed number of secondary cases (R)",
       y = "Frequency")

```




## Estimating the distribution of R

In this part we estimate the underlying distribution of $R$, accounting for
superspreading using a Negative Binomial distibution. The two parameters that
will be estimated are:

* the reproduction number
* the dispersion parameter, which increases when super-spreading events increase
  in frequency.

```{r estimate_R}

## likelihood function for a Neg Bin distribution of R
## (m: the mean; disp: dispersion parameter)

likelihood <- function(mean, disp) {
  ## exp(sum(dpois(old_cases$R, m, log = TRUE)))
  exp(sum(dnbinom(old_cases$R, mu = mean, size = disp, log = TRUE)))
}

R_values <- seq(0, 2, length = 100)
dispersion_values <- seq(0, 2, length = 100)
grid <- expand.grid(R = R_values,
                    dispersion = dispersion_values)
like_values <- apply(grid, 1, function(e) likelihood(e[1], e[2]))

R_ml <- grid[which.max(like_values), "R"]
R_ml
disp_ml <- grid[which.max(like_values), "dispersion"]
disp_ml

df_grid_like <- cbind.data.frame(grid, likelihood = like_values)

ggplot(df_grid_like) +
  geom_raster(aes(x = R, y = dispersion, fill = likelihood)) +
  large_txt +
  scale_fill_gradient(low = "#ffffff", high = "#800000") +
  labs(title = "Likelihood surface: \nnegative binomial distribution",
       x = "Mean reproduction number (R)",
       y = "Dispersion parameter")

```

The likelihood function is well-defined, with a maximum-likelihood estimate of
$R$ is `r round(R_ml, 2)`. How well does this model fit the data? We can compare
predictions from the Negative Binomial distribution (MLE) to the numbers of
secondary cases actually observed:

```{r expected_secondary}

n_cases <- 0:15
obs_freq <- sapply(n_cases, function(e) mean(old_cases$R == e))
pred_freq <- dnbinom(n_cases, mu = R_ml, size = disp_ml)

df_model <- data.frame(n_cases,
                       observed = obs_freq,
                       predicted = pred_freq) %>%
  gather(type, frequency, -1)

ggplot(df_model,
       aes(x = n_cases,
           y = frequency,
           fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  large_txt +
  scale_x_continuous(breaks = int_breaks) +
  scale_fill_manual("", values = c(observed = "#d27979", predicted = "#9999ff")) +
  labs(title = "Nombre de cas secondaires observés et prédits",
       x = "Nombre de cas secondaires",
       y = "Fréquence")

```



## Effective reproduction number in time

```{r R_in_time}

ggplot(R_df, aes(x = date_admission, y = R)) +
  geom_point() +
  geom_smooth() +
  large_txt +
  scale_weeks +
  rotate_x_text(45) +
  labs(title = "Nombre de reproduction effectif au cours du temps",
       x = "Date d'admission",
       y = "Nombre de cas secondaires par cas index")

```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Export tables {.tabset .tabset-fade .tabset-pills}

## Outline

We export R objects (the *epicontacts* object) as well as the main plot as a
separate html file.



## R objects

We export the *epicontacts* object.

```{r rds_exports}

if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}

saveRDS(x,
        file = file.path("produced_rds",
                         "epicontacts_rds"))
destination <- here("data", "clean", "epicontacts_rds")

file.copy(from = file.path("produced_rds",
                           "epicontacts_rds"),
          to = destination,
          overwrite = TRUE)

```






## Excel files

We export the table containing information on: 

* the cases (linelist info)
* the effective reproduction numbers ($R$)
* the identifiers of the transmission chains

```{r xlsx_exports}

if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

rio::export(R_df,
            file = file.path("produced_xlsx",
                             "linelist_with_R_and_chains.xlsx"))

```


**Links to the files:**

- [linelist_with_R_and_chains.xlsx](produced_xlsx/linelist_with_R_and_chains.xlsx)



## visNetwork visualisation

We save the plot of the transmission chain as a separate file called
`plot_chains.html`:

```{r export_html}

visNetwork::visSave(plot_x,
                    file = "plot_chains.html")

visNetwork::visSave(plot_x_active,
                    file = "plot_active_chains.html")

```




<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# System information {.tabset .tabset-fade .tabset-pills}

## Outline

The following information documents the system on which the document was
compiled.



## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```
