---
title: "`r paste('Compare VHF databases -', params$type)`"
author: "Thibaut Jombart"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 2
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

<div class="report_meta">
  
**Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)

**Code contributors:** Thibaut Jombart, Amy Gimma

**Data contributors:** ...

**Version:** 1.0.0

</div>



<br>

To compile this report with a specific data type:

1. double-click on the `open.Rproj` files at the root of the *reportfactory* to open Rstudio

2. in Rstudio, run the command: 

```{r eval = FALSE}

reportfactory::compile_report("compare_data_2019-10-14",
                              params = list(type = "xxx"),
                              clean_report_sources = TRUE)`

```

where you can replace `xxx` by a character string identifying the type of
dataset to use. Make sure the data to compare are stored in
`data/data_comparison/`.



<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Data preparation {.tabset .tabset-fade .tabset-pills}



<!-- ======================================================= -->
## Outline

### Data used

What is the source of your data? Does it need running some previous reports like
preliminary data cleaning? If so, list these reports here.

### Method

The data preparation involves the following steps, detailed in the following tabs:


* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Clean data**: this section contains *ad hoc* data cleaning, i.e. which is
  not used in other reports (otherwise cleaning should be done in a dedicated
  report); this section is also used to create new variables used in the
  analyses



<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->


## Load packages

This part loads required packages. Note that this could be outsourced in a
separate script in `scripts/`, but the advantage of putting pakages here is: i)
we explicitely know what is used and ii) `reportfactory::install_deps()` will
detect these dependencies and install them if needed.

```{r load_packages}

library(tidyverse) # overkill but easier
library(compareDF) # data.frame comparison
library(linelist) # data structure comparison
library(here) # find files
library(rio) # import / export
library(DT) # print interactive tables

```


## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* (optional) all global scripts, i.e. stored outside the factory in
  `../global_scripts/`

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$",
                     full.names = TRUE)
for (file in scripts_files) source(file, local = TRUE)

```




<!-- ======================================================= -->
## Load data

In this section, we:

* look for data files in the `data/data_comparison` folder
* find the two latest datasets
* read them into `x_old` and `x_new`


```{r load_data}

## load the data
data_comparison_folder <- here("data", "data_comparison")

## find order of the files
all_files <- list.files(data_comparison_folder,
                        pattern = params$type,
                        ignore.case = TRUE)
file_order <- all_files %>%
  guess_dates() %>%
  order(decreasing = TRUE)

new_file <- here("data", "data_comparison", all_files[file_order[1]])
old_file <- here("data", "data_comparison", all_files[file_order[2]])


x_new <- rio::import(new_file) %>% as_tibble()
x_old <- rio::import(old_file) %>% as_tibble()


## extract database date from the file name
new_file_date <- gsub("^[^.]+/", "", new_file) %>% guess_dates()
old_file_date <- gsub("^[^.]+/", "", old_file) %>% guess_dates()

```

The **completion dates** of the databases are:

* old data: **`r format(old_file_date, format = "%A %d %b %Y")`**.
* new data: **`r format(new_file_date, format = "%A %d %b %Y")`**.




<!-- ======================================================= -->
## Data cleaning

To keep the original data unchanged as much as possible, we only clean the dates
of `date_report`, needed for further subsetting of the data:

```{r clean}

x_new <- x_new %>% clean_data()
x_old <- x_old %>% clean_data()

```




<!-- ======================================================= -->
## Scales and colors

We define custom colors for the output of `compare_df`.

```{r colors}

comp_colors <- c(addition = "#1FC46F",
                 removal = "#F94444",
                 unchanged_cell = "#7C8AA4",
                 unchanged_row = "#7C8AA4")

```



<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Data comparison {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Explanations

The data comparison includes the following:

* **structure comparison:** look for differences in dimensions, and names,
  ordering and types of the variables
  
* **duplicates:** look for duplicated IDs, and identify the corresponding
  entries; this includes a comparison of the duplicates between old and new
  datasets
  
* **changes in cases:** look for changes in cases, showing changes in colors
  into a separate html table
  



<!-- ======================================================= -->
## Struture comparison

In this part we use *linelist*'s function `compare_data` to compare the
structures of the two datasets: which variables, of which types, did some
variable disappear etc.

```{r structure_comparison}

compare_data(x_old, x_new, use_values = FALSE)

```





<!-- ======================================================= -->
## Looking for duplicates: old data

In this section, we look for duplicated identifiers, and output a table of the
corresponding individuals.

```{r duplicates_old}

to_keep <- x_old %>%
  filter(duplicated(id)) %>%
  pull(id)

duplicates_old <- x_old %>%
  filter(id %in% to_keep)

if (nrow(duplicates_old) > 0) {
  duplicates_old %>%
    show_table()
}

```




<!-- ======================================================= -->
## Looking for duplicates: new data

In this section, we look for duplicated identifiers, and output a table of the
corresponding individuals.

```{r duplicates_new}

to_keep <- x_new %>%
  filter(duplicated(id)) %>%
  pull(id)

duplicates_new <- x_new %>%
  filter(id %in% to_keep)

if (nrow(duplicates_new) > 0) {
  duplicates_new %>%
    show_table()
}

```




<!-- ======================================================= -->
## Changes in duplicates

In this part we compare duplicated entries between the *old* and the *new*
dataset. Note that the table display in this document is sub-optimal. Click on
the link below to open the table.

**Color code**:

* <font color="#7C8AA4">No change</font>
* <font color="#F94444">Removal (in old file, not in the new one)</font>
* <font color="#1FC46F">Addition (in new file, not in old one)</font>



<a href="tables/comparison_duplicates_table.html" target="_blank" class="button" >Open comparison table</a>

<font color="#0B9A86">**Important:**the file will open in your web browser by
default; for even better visualisation, we recommend going to the `tables/`
folder and opening `comparison_duplicates_table.html` with **Excel** </font>

```{r comparison_duplicates, results = "asis"}

duplicates_in_both <- as.logical(
    nrow(duplicates_old) *
    nrow(duplicates_new))

if (duplicates_in_both) {

  comparison_duplicates <- compareDF::compare_df(
                                          duplicates_new,
                                          duplicates_old,
                                          group_col = "id",
                                          limit_html = 1000,
                                          color_scheme = comp_colors,
                                          keep_unchanged_rows = TRUE,
                                          stop_on_error = FALSE)
}

```







<!-- ======================================================= -->
## Changes in cases

In this section we look for changes in cases between the two versions of the
dataset, using the *compareDF* package to check which entries of the data have
changed. This is computer-intensive work. To make it run under reasonable time,
we restrict the comparisons to cases reported within the last 42 days (*Filter*
section).

Note that the table display in this document is sub-optimal. Click on
the link below to open the table.

**Color code**:

* <font color="#7C8AA4">No change</font>
* <font color="#F94444">Removal (in old file, not in the new one)</font>
* <font color="#1FC46F">Addition (in new file, not in old one)</font>

<a href="tables/comparison_table.html" target="_blank" class="button" >Open comparison table</a>

<font color="#0B9A86">**Important:** the file will open in your web browser by
default; for even better visualisation, we recommend going to the `tables/`
folder and opening `comparison_duplicates_table.html` with **Excel** </font>


```{r comparison, results = "asis"}

comparison <- compareDF::compare_df(
                             x_new,
                             x_old,
                             group_col = "id",
                             limit_html = 1000,
                             color_scheme = comp_colors,
                             stop_on_error = FALSE)

```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Export tables {.tabset .tabset-fade .tabset-pills}


<!-- ======================================================= -->
## Create output directory

We create a directory called `tables/` to store output files, if it does not
exist.

```{r create_tables_dir}

if (!dir.exists("tables")) {
  dir.create("tables")
}

```



<!-- ======================================================= -->
## Create output files

The following items are exported:

```{r exports}

if (duplicates_in_both) {
  writeLines(comparison_duplicates$html_output,
             con = file.path("tables",
                             "comparison_duplicates_table.html"),
             useBytes = TRUE)
}

writeLines(comparison$html_output,
           con = file.path("tables",
                            "comparison_table.html"),
           useBytes = TRUE)

if (nrow(duplicates_old) > 0) {
  rio::export(duplicates_old, file = file.path("tables",
                                               "duplicates_old.xlsx"))
}

if (nrow(duplicates_new) > 0) {
  rio::export(duplicates_new, file = file.path("tables",
                                               "duplicates_new.xlsx"))
}

```




<!-- ======================================================= -->
## Links to output files

Click on the links below to open items:

**Changes in cases:**

* [comparison_duplicates_table](tables/comparison_duplicates_table.html): html table of comparisons amongst old and new duplicates

* [comparison_table](tables/comparison_table.html): html table of comparisons amongst old and new cases


**Duplicates:**

* [duplicates_old](tables/duplicates_old.xlsx): Excel file containing duplicated entries (old dataset)

* [duplicates_new](tables/duplicates_new.xlsx): Excel file containing duplicated entries (new dataset)






<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# System information {.tabset .tabset-fade .tabset-pills}


<!-- ======================================================= -->
## Outline

The following information documents the system on which the document was
compiled.



<!-- ======================================================= -->
## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```



<!-- ======================================================= -->
## R environment

This provides information on the version of R used:

```{r R_session}

R.version

```




<!-- ======================================================= -->
## R packages

This provides information on the packages used:

```{r R_pkg}

sessionInfo()

```



<!-- ======================================================= -->
## Compilation parameters

This provides information on the parameters (passed through `params`) used for
compiling this document:

```{r params}

params

```
