---
title: "Investigation of alerts data: Goma"
author: "Thibaut Jombart, Flavio Finger, Christopher Jarvis for the analytic cell OEC Goma"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 1
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, fig.width = 8,
                      fig.height = 6, dpi = 150, warning = FALSE,
                      message = FALSE)
```




<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Data preparation {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

This report cleans and analyses the alert data of Goma. Input comes from an
`xlsx` file containing alerts in a specific format. Because all
sub-coordinations have different standards, each sub-coordination needs a
separate report.


The data preparation involves the following steps, detailed in the following tabs:


* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Clean data**: this section contains *ad hoc* data cleaning, i.e. which is
  not used in other reports (otherwise cleaning should be done in a dedicated
  report); this section is also used to create new variables used in the
  analyses




<!-- ======================================================= -->
## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all overall scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned VHF data stored as `x`


```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$",
                     full.names = TRUE)
for (file in scripts_files) source(file)

ggthemr("fresh")

```



<!-- ======================================================= -->
## Load data

We extract the completion date from the file name:

```{r load_data}

## load the data
current_goma
x_raw <- custom_import(current_goma)
glimpse(x_raw)

## extract database date from the file name
file_name <- gsub("^[^.]+/", "", current_goma)
database_date <- file_name %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.






<!-- ======================================================= -->
## Clean data

We use *linelist*'s function `clean_data()` to:

- remove all special characters from the data
- set all characters to lower case
- replace all accentuated and diacritic characters with their closest ascii
  match in the latin alphabet
- replace all separators with a single `_`
- replace all mis-spelling using a overall dictionary (see the file
  `cleaning_rules.xlsx` in `/dictionary/`, or type `cleaning_rules` in this R
  session, after compiling the document
- (optionally) detect date formats and convert data to dates, including cases
  where format varies within a column
  

This cleaning is achieved with:

```{r data_cleaning}

x <- x_raw %>%
  clean_data(guess_dates = FALSE,
             wordlists = cleaning_rules) %>%
  as_tibble()

x

```



<!-- ======================================================= -->
## Renaming variables and sanity checks

The following variables will be used, and are therefore checked:

- `date`: check no date is after the database completion
- sanitize `genre`
- make `origin` as a sanitized version of `source_notif`

```{r check_variables}

## date of notification
x <- x %>% mutate(date = guess_dates(date))
range(x$date, na.rm = TRUE)

## gender
table(x$sexe, useNA = "ifany")

## status
x <- mutate(x, status = result_invest)
table(x$status, useNA = "ifany")


## zone de sante
x <- mutate(x, zone_de_sante = factor(zone_sante))
table(x$zone_de_sante, useNA = "ifany")


## origin / source_notif
x <- mutate(x, origin = type_surveillance)
table(x$origin, useNA = "ifany")


x$origin %>% unique

## bleed
x <- mutate(x, bleed = saignements)
table(x$bleed, useNA = "ifany")


## contact connu
x <- mutate(x, contact_connu = lien_epi)
table(x$contact_connu, useNA = "ifany")


```


One last round of dictionary-based cleaning:

```{r last_cleaning}

x <- x %>%
  clean_variable_spelling(wordlists = cleaning_rules)
x

```


<!-- ======================================================= -->
## Outcomes

Outcome is defined for alerts which have either been validated or invalidated.

```{r outcomes}

## get only known outcomes
outcomes <- x %>%
  filter(status %in% c("validee", "invalidee")) %>%
  droplevels() %>% 
  mutate(validee = 1 * (status == "validee"))

```



<!-- ======================================================= -->
## Expected decisions

Alerts are supposed to be validated using the following key, depending on
whether the person has:

1. **known contact** with a case and at least **one symptom**
2. if not 1, **unexplained bleeding**
3. if not 2, **fever and 3 other symptoms**

Note that 3 de facto excludes bleeding as a candidate symptom. We create a new
variable which asserts these conditions:

```{r expected_result}

## elements of diagnostic
## criteria 1
has_contact <- outcomes$contact_connu %in% c("confirme", "oui")
has_one_symptom <- outcomes %>%
  select(fievre,
         nausee_vom,
         diarrhee,
         fatigue,
         anorexie,
         dlr_abdo,
         dlr_thorax,
         dlr_muscu,
         dlr_artic,
         cephalee,
         toux,
         dyspnee,
         dysphagie,
         odynophagie,
         hoquet,
         bleed) %>%
  apply(1, function(e) any(e == "oui", na.rm = TRUE))
fits_1 <- has_contact & has_one_symptom

## criteria 2
has_bleeding <- outcomes$bleed == "oui"
fits_2 <- !fits_1 & has_bleeding

## criteria 3
has_fever <- outcomes$fievre == "oui"
has_3_symptoms <- outcomes %>%
    select(nausee_vom,
         diarrhee,
         fatigue,
         anorexie,
         dlr_abdo,
         dlr_thorax,
         dlr_muscu,
         dlr_artic,
         cephalee,
         toux,
         dyspnee,
         dysphagie,
         odynophagie,
         hoquet,
         bleed) %>% 
  apply(1, function(e) sum(e == "oui", na.rm = TRUE)) >= 3
fits_3 <- !fits_1 & !fits_2 & has_fever & has_3_symptoms


## make sure criteria are exclusive - no number in the table below should exceed
## 1
table(fits_1 + fits_2 + fits_3)

## expected decisions
outcomes <- outcomes %>%
  mutate(admit_contact = fits_1,
         admit_bleeding = fits_2,
         admit_fever = fits_3) %>% 
  mutate(expected_decision = ifelse(
             fits_1 | fits_2 | fits_3,
             "validee",
             "invalidee"),
         decision_comparison = case_when(
             status == "validee" & expected_decision == "validee" ~
               "true_positive",
             status == "invalidee" & expected_decision == "invalidee" ~
               "true_negative",
             status == "validee" & expected_decision == "invalidee" ~
               "false_positive",
             status == "invalidee" & expected_decision == "validee" ~
               "false_negative",
             TRUE ~ NA_character_
             ),
         decision_comparison =
           factor(decision_comparison,
                  levels = c("true_positive",
                             "true_negative",
                             "false_positive",
                             "false_negative")))

## remove alerts with missing comparisons, from 2019
outcomes <- outcomes %>%
  filter(!is.na(decision_comparison),
         date >= as.Date("2019-01-01"))


## check that classification is well-made
outcomes %>%
  group_by(status, expected_decision, decision_comparison) %>%
  count()

```






<!-- ======================================================= -->
## Major health zones

We define major health zones as the ones having reported at least 100 alerts in
the outbreak, since the 1st January 2019.

```{r x_major}

major_zones <- x %>%
  group_by(zone_de_sante) %>%
  count() %>%
  filter(n >= 100) %>% 
  pull(zone_de_sante)

x_major <- x %>%
  filter(zone_de_sante %in% major_zones,
         date >= as.Date("2019-01-01"))

outcomes_major <- outcomes %>%
  filter(zone_de_sante %in% major_zones)


```




<!-- ======================================================= -->
## Last 21 days

We duplicate the previous datasets, retaining the 21 days leading up to the
current database date.

```{r subset_21_days}

last_date <- database_date - 21
x_recent <- filter(x, date >= last_date)
x_major_recent <- filter(x_major, date >= last_date)
outcomes_recent <- filter(outcomes, date >= last_date)
outcomes_major_recent <- filter(outcomes_major, date >= last_date)

```




<!-- ======================================================= -->
## Custom color scales

We define custom colors for some of the variables used in the plots.

```{r scales_fill}

scale_origins <- scale_fill_manual(
  "Origine",
  values = c(communautaire = "#ffcc00",
             recherche_active = "#c3c388",
             surveillance_passive = "#ff6699",
             point_entree = "#40bf80",
             autre = "#668cff",
             inconnu = "#8c8c8c"))

scale_decisions <- scale_fill_manual(
    "Décisions",
    values = c(true_positive = "#94b8b8",
               true_negative = "#8c8cd9",
               false_positive = "#ff8080",
               false_negative = "#b3003b"),
    labels = c(true_positive = "vrai positif",
               true_negative = "vrai négatif",
               false_positive = "faux positif",
               false_negative = "faux négatif"))

```











<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Alerts for the entire outbreak {.tabset .tabset-fade .tabset-pills}

## Outline

This section summarises alerts according to the following criteria:

* overall
* origins by health zone
* origins in time
* decisions by health zone
* decisions in time



<!-- ======================================================= -->
## Overall numbers

Here we report the overall numbers of alerts by health zone. Note that we cannot
use `geom_bar` for the counts as it drops zones with no alerts.

```{r overall}

ggplot(x, aes(x = zone_de_sante)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par zones de santé") +
  large_txt +
  rotate_x_text(45)

table_overall <- x %>%
  group_by(zone_de_sante) %>%
  count() %>%
  adorn_totals()

table_overall %>%
  knitr::kable()

```




<!-- ======================================================= -->
## Origins by health zone

This section focuses on the origin of alerts, by health zones.

```{r origins}

ggplot(x,
       aes(x = zone_de_sante, fill = origin)) +
  geom_bar(color = "white") +
  scale_x_discrete(drop = FALSE) +
  scale_origins +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par zones de santé") +
  large_txt +
  rotate_x_text(45)

table_origins <- x %>%
  group_by(zone_de_sante, origin) %>%
  count() %>% 
  spread(origin, n, fill = 0) %>% 
  adorn_totals(c("row", "col"))

table_origins %>%
  show_table()

```




<!-- ======================================================= -->
## Origins in time

This section focuses on the temporal evolution of origin of alerts by health
zones.

```{r origins_time}

ggplot(x_major, aes(x = date, fill = origin)) +
  geom_histogram(binwidth = 7, col = "white") +
  facet_grid(zone_de_sante ~ .) +
  scale_months +
  scale_origins +
  rotate_x_text(45) +
  large_txt +
  smaller_axis_txt +
  custom_vert_facet +
  theme(legend.position = "bottom") +
  labs(title = paste("Evolution des origines d'alertes au cours du temps",
                     "par zone de sante (>= 100 alertes)",
                     sep = "\n"),
       x = "",
       y = "Nombre d'alertes par semaine")

table_origins_time <- incidence(x$date,
                                "week",
                                groups = x$origin) %>%
  as.data.frame() %>%
  adorn_totals("col")

table_origins_time %>%
  show_table()

```






<!-- ======================================================= -->
## Decisions by health zone

This section focuses on the decision of alerts, by health zones.

When looking at decisions on the treatment of alerts, we identify the 4
following situations:

* **<font color = "#94b8b8">true positive</font>**: alerts were rightfully
  validated (patient tested)
* **<font color = "#8c8cd9">true negative</font>**: alerts were rightfully not
  validated (patient not tested)
* **<font color = "#ff8080">false positive</font>**: alerts were wrongly
  validated (patient tested, shoud not have been)
* **<font color = "#b3003b">false negative</font>**: alerts were wrongly not
  validated (patient not tested, should have been)

**False positive** create a waste of resources as well as un-necessary pressure
on the patient and community. **False negative** create a risk of missing cases.

```{r decisions}

ggplot(outcomes,
       aes(x = zone_de_sante, fill = decision_comparison)) +
  geom_bar(color = "white") +
  scale_x_discrete(drop = FALSE) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par zones de santé") +
  large_txt +
  scale_decisions +
  rotate_x_text(45)

table_decisions <- outcomes %>%
  group_by(zone_de_sante, decision_comparison) %>%
  count() %>%
  spread(decision_comparison, n, fill = 0) %>% 
  adorn_totals(c("row", "col")) %>%
  mutate(prop_false_positive_95ci =   prop_to_display_ci(false_positive, Total, dec = 2, perc = TRUE),
         prop_false_negative_95ci = prop_to_display_ci(false_negative, Total, dec = 2, perc = TRUE),
         prop_false_positive = prop_to_perc(false_positive / Total),
         lower_false_positive = prop_ci(false_positive, Total, "lower", TRUE),
         upper_false_positive = prop_ci(false_positive, Total, "upper", TRUE),
         prop_false_negative = prop_to_perc(false_negative / Total),
         lower_false_negative = prop_ci(false_negative, Total, "lower", TRUE),
         upper_false_negative = prop_ci(false_negative, Total, "upper", TRUE)
         )

table_decisions %>%
  select(-prop_false_positive,
          -lower_false_positive,
          -upper_false_positive,
          -prop_false_negative ,
          -lower_false_negative,
          -upper_false_negative
         ) %>% 
  show_table()

```


<!-- ======================================================= -->
## Incorrectly validated and incorrectly not validated


* Individuals incorrectly validated. They were validated and tested but did not
  meet the case definition.

* Individuals incorrectly not validated. They were not validated and tested but
  did meet the case definition.

```{r sens_spec}

table_sens_spec <- outcomes %>%
  group_by(zone_de_sante, decision_comparison) %>%
  count() %>%
  spread(decision_comparison, n, fill = 0) %>% 
  mutate(total_tested = true_positive + false_positive,
         total_not_tested = true_negative + false_negative) %>% 
  mutate(
      incorrectly_validated_95ci = prop_to_display_ci(
          false_positive,
          total_tested,
          dec = 2,
          perc = TRUE),
      incorrectly_not_validated_95ci = prop_to_display_ci(
          false_negative,
          total_not_tested,
          dec = 2, perc = TRUE)
         )
table_sens_spec %>%
  select(
    -false_negative,
    -true_negative
         ) %>% 
  show_table()

```





<!-- ======================================================= -->
## Decisions in time

This section focuses on the temporal evolution of decision of alerts by health
zones.

When looking at decisions on the treatment of alerts, we identify the 4
following situations:

* **<font color = "#94b8b8">true positive</font>**: alerts were rightfully
  validated (patient tested)
* **<font color = "#8c8cd9">true negative</font>**: alerts were rightfully not
  validated (patient not tested)
* **<font color = "#ff8080">false positive</font>**: alerts were wrongly
  validated (patient tested, shoud not have been)
* **<font color = "#b3003b">false negative</font>**: alerts were wrongly not
  validated (patient not tested, should have been)

**False positive** create a waste of resources as well as un-necessary pressure
on the patient and community. **False negative** create a risk of missing cases.

```{r decisions_time}

ggplot(outcomes_major, aes(x = date, fill = decision_comparison)) +
  geom_histogram(binwidth = 7, col = "white") +
  facet_grid(zone_de_sante ~ .) +
  scale_months +
  rotate_x_text(45) +
  large_txt +
  smaller_axis_txt +
  custom_vert_facet +
  scale_decisions +
  theme(legend.position = "bottom") +
  labs(title = paste("Evolution des decisions d'alertes au cours du temps",
                     "par zone de sante (>= 100 alertes)",
                     sep = "\n"),
       x = "",
       y = "Nombre d'alertes par semaine")

table_decisions_time <- incidence(outcomes$date,
                                  "week",
                                  groups = outcomes$decision_comparison) %>%
  as.data.frame() %>%
  adorn_totals("col")

table_decisions_time %>%
  show_table()

```












<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Alerts for the last 21 days {.tabset .tabset-fade .tabset-pills}


```{r deactivate_recent, include = FALSE}
no_recent <- nrow(x_recent) == 0
knitr::opts_chunk$set(eval = !no_recent)
```

```{r message_no_recent, echo = FALSE, include = TRUE, eval = TRUE, results = "asis"}

if (no_recent) {
  cat('\n<font color = "#cc0044" size = 6>No recent alerts reported in the database!</font>')
  cat(sprintf('\n<font color = "#cc0044" size = 6>Latest alert reported on the %s.</font>',
              format(max(x$date, na.rm = TRUE), "%A %d %B %Y")))
}

```



<!-- ======================================================= -->
## Outline

This section summarises alerts according to the following criteria, for the 21
days leading up to the current database, i.e. from the **`r format(last_date,
format = "%A %d %b %Y")`** to the **`r format(database_date, format = "%A %d %b
%Y")`**:

* overall
* origins by health zone
* origins in time
* decisions by health zone
* decisions in time

The next main section will reproduce these analyses identically, but for the
last 21 days.



<!-- ======================================================= -->
## Overall numbers

Here we report the overall numbers of alerts by health zone.

```{r overall_recent}

ggplot(x_recent, aes(x = zone_de_sante)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par zones de santé") +
  large_txt +
  rotate_x_text(45)

table_overall_recent <- x_recent %>%
  group_by(zone_de_sante) %>%
  count() %>%
  adorn_totals()

table_overall_recent %>%
  knitr::kable()

```




<!-- ======================================================= -->
## Origins by health zone

This section focuses on the origin of alerts, by health zones.

```{r origins_recent}

ggplot(x_recent,
       aes(x = zone_de_sante, fill = origin)) +
  geom_bar(color = "white") +
  scale_x_discrete(drop = FALSE) +
  scale_origins +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par zones de santé") +
  large_txt +
  rotate_x_text(45)

table_origins_recent <- x_recent %>%
  group_by(zone_de_sante, origin) %>%
  count() %>%
  spread(origin, n, fill = 0) %>% 
  adorn_totals(c("row", "col"))

table_origins_recent %>%
  show_table()

```




<!-- ======================================================= -->
## Origins in time

This section focuses on the temporal evolution of origin of alerts by health
zones.

```{r origins_time_recent}

ggplot(x_major_recent, aes(x = date, fill = origin)) +
  geom_histogram(binwidth = 1, col = "white") +
  facet_grid(zone_de_sante ~ .) +
  scale_weeks +
  scale_origins +
  rotate_x_text(45) +
  large_txt +
  smaller_axis_txt +
  custom_vert_facet +
  theme(legend.position = "bottom") +
  labs(title = paste("Evolution des origines d'alertes au cours du temps",
                     "par zone de sante (>= 100 alertes)",
                     sep = "\n"),
       x = "",
       y = "Nombre d'alertes par jour")

table_origins_time_recent <- incidence(x_recent$date,
                                1,
                                groups = x_recent$origin) %>%
  as.data.frame() %>%
  adorn_totals("col")

table_origins_time_recent %>%
  show_table()

```






<!-- ======================================================= -->
## Decisions by health zone

This section focuses on the decision of alerts, by health zones.

When looking at decisions on the treatment of alerts, we identify the 4
following situations:

* **<font color = "#94b8b8">true positive</font>**: alerts were rightfully
  validated (patient tested)
* **<font color = "#8c8cd9">true negative</font>**: alerts were rightfully not
  validated (patient not tested)
* **<font color = "#ff8080">false positive</font>**: alerts were wrongly
  validated (patient tested, shoud not have been)
* **<font color = "#b3003b">false negative</font>**: alerts were wrongly not
  validated (patient not tested, should have been)

**False positive** create a waste of resources as well as un-necessary pressure
on the patient and community. **False negative** create a risk of missing cases.

```{r decisions_recent}

ggplot(outcomes_recent,
       aes(x = zone_de_sante, fill = decision_comparison)) +
  geom_bar(color = "white") +
  scale_x_discrete(drop = FALSE) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par zones de santé") +
  large_txt +
  scale_decisions +
  rotate_x_text(45)

table_decisions_recent <- outcomes_recent %>%
  group_by(zone_de_sante, decision_comparison) %>%
  count() %>%
  spread(decision_comparison, n, fill = 0) %>% 
  adorn_totals(c("row", "col")) %>%
  mutate(prop_false_positive_95ci =   prop_to_display_ci(
             false_positive,
             Total,
             dec = 2,
             perc = TRUE),
         prop_false_negative_95ci = prop_to_display_ci(
             false_negative,
             Total,
             dec = 2,
             perc = TRUE),
         prop_false_positive = prop_to_perc(false_positive / Total),
         lower_false_positive = prop_ci(false_positive, Total, "lower", TRUE),
         upper_false_positive = prop_ci(false_positive, Total, "upper", TRUE),
         prop_false_negative = prop_to_perc(false_negative / Total),
         lower_false_negative = prop_ci(false_negative, Total, "lower", TRUE),
         upper_false_negative = prop_ci(false_negative, Total, "upper", TRUE)
         )

table_decisions_recent %>%
  select(-prop_false_positive,
          -lower_false_positive,
          -upper_false_positive,
          -prop_false_negative ,
          -lower_false_negative,
          -upper_false_negative
         ) %>% 
  show_table()

```





<!-- ======================================================= -->

## Incorrectly validated and incorrectly not validated

* Individuals incorrectly validated. They were validated and tested but did not
  meet the case definition.

* Individuals incorrectly not validated. They were not validated and tested but
  did meet the case definition.

```{r sens_spec_recent}

table_sens_spec_recent <- outcomes_recent %>%
  group_by(zone_de_sante, decision_comparison) %>%
  count() %>%
  spread(decision_comparison, n, fill = 0) %>% 
  mutate(total_tested = true_positive + false_positive,
         total_not_tested = true_negative + false_negative) %>% 
  mutate(
      incorrectly_validated_95ci = prop_to_display_ci(
          false_positive,
          total_tested,
          dec = 2, perc = TRUE),
      incorrectly_not_validated_95ci = prop_to_display_ci(
          false_negative,
          total_not_tested,
          dec = 2,
          perc = TRUE)
         )
table_sens_spec_recent %>%
  select(
    -false_negative,
    -true_negative
         ) %>% 
  show_table()

```




<!-- ======================================================= -->
## Decisions in time

This section focuses on the temporal evolution of decision of alerts by health
zones.

When looking at decisions on the treatment of alerts, we identify the 4
following situations:

* **<font color = "#94b8b8">true positive</font>**: alerts were rightfully
  validated (patient tested)
* **<font color = "#8c8cd9">true negative</font>**: alerts were rightfully not
  validated (patient not tested)
* **<font color = "#ff8080">false positive</font>**: alerts were wrongly
  validated (patient tested, shoud not have been)
* **<font color = "#b3003b">false negative</font>**: alerts were wrongly not
  validated (patient not tested, should have been)

**False positive** create a waste of resources as well as un-necessary pressure
on the patient and community. **False negative** create a risk of missing cases.

```{r decisions_time_recent}

ggplot(outcomes_major_recent, aes(x = date, fill = decision_comparison)) +
  geom_histogram(binwidth = 1, col = "white") +
  facet_grid(zone_de_sante ~ .) +
  scale_weeks +
  rotate_x_text(45) +
  large_txt +
  smaller_axis_txt +
  custom_vert_facet +
  scale_decisions +
  theme(legend.position = "bottom") +
  labs(title = paste("Evolution des decisions d'alertes au cours du temps",
                     "par zone de sante (>= 100 alertes)",
                     sep = "\n"),
       x = "",
       y = "Nombre d'alertes par jour")

table_decisions_time_recent <- incidence(outcomes_recent$date,
                                         1,
                                         groups = outcomes_recent$decision_comparison) %>%
  as.data.frame() %>%
  adorn_totals("col")

table_decisions_time_recent %>%
  show_table()

```










<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Export data and tables  {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

We export the clean data to the clean data folder and some of the relevant
tables, which will be placed in the current working directory.




<!-- ======================================================= -->
## R objects

We export some of the clean database, placed in `produced_rds/` as well as in
`data/clean/`:

```{r export_rds, eval = TRUE}

## check if a directory exists and if not then creates it
if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}

## create the text for the file name with the database date
rds_file_name <- sprintf("%sclean_%s.rds",
                     undated_file_name(current_goma),
                     format(database_date, "%Y-%m-%d"))
rds_file_name

## save the rds file in the produced_rds folder
rio::export(x,
            file.path("produced_rds", rds_file_name))

```

We copy these files to the `data/clean` folder:

```{r copy_rds, eval = TRUE}
# copy some files into `data/clean/`

# Provide the destination of where to copy the data
destination <- here("data",
                    "clean",
                    rds_file_name)
# Copy the rds data
file.copy(from = file.path("produced_rds", rds_file_name),
          to = destination,
          overwrite = TRUE)

```





<!-- ======================================================= -->
## Excel files

```{r exports}

if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

to_export <- c("table_decisions",
               "table_decisions_time", 
               "table_origins",
               "table_origins_time",
               "table_overall", 
               "table_sens_spec"
               )

for (e in to_export) {
  rio::export(get(e),
              file.path("produced_xlsx",
                        paste0(e, ".xlsx")))
}



```

Export recent tables (if there are recent alerts):

```{r exports_recent, eval = !no_recent}

to_export_recent <- c("table_decisions_recent",
               "table_decisions_time_recent",
               "table_origins_recent", 
               "table_origins_time_recent",
               "table_overall_recent",
               "table_sens_spec_recent"
               )

for (e in to_export_recent) {
  rio::export(get(e),
              file.path("produced_xlsx",
                        paste0(e, ".xlsx")))
}


```


Click on the following links to open the files (only works if the files above
have been generated and are in the same folder as this document):

Overall tables:

```{r links, results = "asis", echo = FALSE}

for (e in to_export) {
  txt <- sprintf("- [%s.xlsx](%s.xlsx)",
                 e,
                 file.path("produced_xlsx",
                           e))
  cat(txt, sep = "\n")
}

```

<br>
Recent tables:

```{r links_recent, results = "asis", echo = FALSE, eval = !no_recent}

for (e in to_export_recent) {
  txt <- sprintf("- [%s.xlsx](%s.xlsx)",
                 e,
                 file.path("produced_xlsx",
                           e))
  cat(txt, sep = "\n")
}

```







<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# System information {.tabset .tabset-fade .tabset-pills}

## Outline

The following information documents the system on which the document was
compiled.



<!-- ======================================================= -->
## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```



<!-- ======================================================= -->
## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```



<!-- ======================================================= -->
## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```



<!-- ===================================================== -->
## Compilation parameters

This shows which parameters were passed through `params` at compilation time:

```{r params}
params
```





<!-- ===================================================== -->
## Change log

### Version 1.0.0 (25 Oct 2019)

* revert back to version without incidence of alerts, as other version was
  broken for non-obvious reasons
  
* using new template for automated export of files

* fixing the appearance of facetted graphs

