---
title: "Linelist data cleaning"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    toc: TRUE
    toc_depth: 4
    toc_float: true
    toc_collapse: false
    number_sections: true
    highlight: pygments
    theme: spacelab
    code_folding: hide
    css: !expr here::here('css', 'style.css')
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE)
```


<div class="report_meta">
  <span class="notice">**Notice**: this is a **stable, routine report**. 
  **Do not touch it unless it is broken.** To make a contribution, **carefully read 
  the [README](../../../../../README.html) file**.</span>
  
  **Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)
  
  **Code contributors:** Thibaut Jombart, Flavio Finger, Christopher Jarvis, Madeleine Crowe, Jonathan Polonsky, Charlie Whittaker and the rest of the epi analysis cell, EOC Goma
  
  **Data contributors:**
  
  **Version:** 2.0.0
  
  **Reviewed by:** Amy Gimma
</div>


<!-- ================================= -->
<!-- ================================= -->
<!-- ================================= -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This is the universal cleaning script for master linelist database.

### Data used

Input data is the master linelist with one row per case. 
**All input files must be `xlsx` files.**

### Method

The data preparation involves the following steps, detailed in the following tabs:

* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded; scripts include the data-cleaning dictionary and the paths to the
  current data

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.


## Load scripts

These scripts will load:

* required packages
* a few helper functions
* a data-cleaning dictionary (`cleaning_rules`)
* the path to current version of raw data (`current_data`)

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names = TRUE)

for (file in scripts_files) source(file, local = TRUE)

```



## Load data

```{r load_data}

## record current date (time at which document is compiled)
today <- as.Date(Sys.time())

## get data file
## use excel and then the file encoding should not be an issue

current_data
check_file(current_data)

x_raw <- custom_import(current_data) %>%
  clean_data(guess_dates = FALSE) %>%
  as_tibble()

dim(x_raw)

```



## Completion date

We extract the completion date from the file name:

```{r database_date}

file_name <- extract_file_name(current_data)
database_date <- file_name %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.



<!-- ================================= -->
<!-- ================================= -->
<!-- ================================= -->

# Data cleaning {.tabset .tabset-fade .tabset-pills}

## Outline 

The data cleaning in this script must provide the data needed for all Linelist-based
reports. It involves the following steps:

* **data standardisation** using `clean_data`, which set all data to lower case,
  removes special characters, replaces non-ascii characters with their closest
  ascii match
  
* defining new variables from existing ones (mostly renaming) 

* cleaning dates

* using a dictionary-based cleaning: rules are defined in
  `../dictionary/cleaning_rules.xlsx`

* defining other variables, e.g. age categories or health-care worker status



## Aliases and variable renaming

We create aliases for some of the variables we'll use later (we use multiple
`mutate` to be able to track errors):

```{r aliases}

## data that would be good to add:
##
## - date_discharge from ETC

x <- x_raw %>%
  mutate(epicasedef = factor(statut,
                             levels = c("confirmed", "probable"))) %>%
  mutate(date_onset = onset_cnc) %>% 
  mutate(date_report = reportdate_cnc) %>% 
  mutate(date_admission = cte_date) %>%
  mutate(date_death = deathdischargedate) %>% 
  mutate(date_exposure_start = date_contact_confirmed_case1) %>% 
  mutate(date_exposure_end = enddate1_contact_confirmed) %>%
  mutate(date_vacc = vacc_date) %>%
  mutate(current_status = final_outcome) %>%
  mutate(report_status = statusreport_mdc) %>% 
  mutate(outcome = final_outcome) %>% 
  mutate(zone_de_sante = zs_calc) %>% 
  mutate(aire_de_sante = health_area) %>%
  mutate(age = age_cnc) %>%
  mutate(gender = sex) %>% 
  mutate(epiweek_report = aweek::date2week(date_report,
                                           week_start = "Monday", 
                                           floor_day = TRUE)) %>% 
  mutate(epiweek_report_label = aweek::week2date(epiweek_report,
                                                 week_start = "Monday")) %>% 
  mutate(epiweek_onset = aweek::date2week(date_onset,
                                          week_start = "Monday", 
                                          floor_day = TRUE)) %>% 
  mutate(epiweek_onset_label = aweek::week2date(epiweek_onset,
                                                week_start = "Monday")) %>%
  mutate(month_report = floor_date(date_report, unit = "month"),
         month_onset = floor_date(date_onset, unit = "month")) %>% 
  mutate(transmission_place_group = transmission_place) %>%
  mutate(place_death_crude = place_of_death)

```



## Clean dates

The date cleaning uses `guess_dates`, with some additional specifications of
priorities of formats when they can be ambiguous. This operation is
time-consuming.

Note that all sorts of issues happen with dates imported from Google Spreadsheet
or Excel. A sneaky one is that of dates represented as numbers but stored as
characters, e.g. `"43026"`. The conversion to `numeric` and then to `Date`
requires an origin, which differs in the native Excel format and a google
spreadsheet, even if it has been saved by Excel:

* Excel's origin is 1900-01-01 (for VHF)
* google spreadsheet's origin is 1889-12-30 (for MLL)

As the MLL comes from Google Spreadsheet, it will implicitely use the
`1889-12-30` as origin. We pass it to our tweak of `guess_dates()`.

```{r replace_dates}

x <- x %>% 
  mutate_at(
      vars(starts_with("date_")),
      guess_dates, origin = as.Date("1889-12-30"))


```



## Dictionary-based cleaning

Please see cleaning rules in `../dictionary/cleaning_rules.xlsx`

```{r dictionary_cleaning}

x <- x %>%
  clean_variable_spelling(wordlists = cleaning_rules) 

x <- x %>% 
  mutate(health_area = aire_de_sante)

```

## Define contact status

``` {r contact status}

x <- x %>%
  mutate(contact_status = case_when(
             contact_registered == "no" ~ "contact_inconnu",
             contact_registered == "yes" & 
             (contact_surveilled == "no" | 
              contact_surveilled == "unknown") ~ "connu_non_suivi",
             contact_surveilled == "yes" ~ "suivi",
             contact_registered == "unknown" ~ "non_enregistre")) %>% 
  mutate(contact_status = factor(contact_status, 
                                 levels = c("non_enregistre", 
                                            "contact_inconnu",
                                            "connu_non_suivi", 
                                            "suivi"))) 

```



## Define sub-coordinations

Sub-coordinations are determined from the `zone_de_sante` variable, using
cleaning rules to assign each zone to a sub-coordination.

```{r sous_coordination}

# clean_variable_spelling assigns sous_coordination from zs (relies on spelling)
x <- x %>%
  mutate(sous_coordination = zone_de_sante) %>%
  clean_variable_spelling(wordlists = cleaning_rules) 

# Check for NAs and mispellings
table(x$zone_de_sante, x$sous_coordination, useNA = "ifany")

```



## Identify community death

Community deaths are defined as deaths which occured outside ETCs and TCs. 

Values of the variable `type_death` will be defined as:

*  `community` if:
  1. place of death is community
  2. patient was dead on notification, place of death is unknown, and there is
     no date of cte admission
* `etc_tc` if patients died in ETCs / TCs
* `alive` for alive patients
* `unknown` in other situations (unknown outcomes)

``` {r type_death} 

table(x$type_death, useNA = "ifany")

x <- x %>%
  mutate(type_death =
           case_when(
               ## commu. death condition 1
               place_death_crude == "community" ~ "community",
               ##  commu. death condition 2
               report_status == "dead" &
               place_death_crude == "unknown" &
               is.na(cte_date) ~ "community",
               outcome == "dead" &
               place_death_crude == "etc" ~ "etc_tc",
               current_status %in% c("alive", "survived") ~ "alive",
               TRUE ~ "unknown"))

# Check for NAs
table(x$type_death, useNA = "ifany")

```

The breakdown amongst deaths is:

```{r type_death_breakdown}

x %>%
  filter(outcome == "dead") %>%
  mutate(has_cte_date = !is.na(cte_date)) %>% 
  count(type_death, place_death_crude, report_status, has_cte_date) %>%
  show_table()

```




## Define provinces

Province are defined from the *sub-coordinations*, which is almost exact except
for Mandima, which needs to be re-assigned to *Ituri*.

```{r provinces}

# clean_variable_spelling assigns province from sc (relies on spelling)

x <- x %>%
  mutate(province = zone_de_sante) %>% 
  clean_variable_spelling(wordlists = cleaning_rules) 

## check that provinces match

table(x$province, x$zone_de_sante, useNA = "ifany")

```



## Age and age classes

This section defines age classes - see code for details.

```{r clean_age}

x <- x %>%
  mutate(
      age = as.numeric(gsub("_", ".", age)),
      age_class = factor(
          case_when(
              age <= 5 ~ "<=5",
              age <= 10 ~ "6-10",
              age <= 17 ~ "11-17",
              age <= 25 ~ "18-25",
              age <= 35 ~ "26-35",
              age <= 45 ~ "36-45",
              age <= 55 ~ "46-55",
              age <= 65 ~ "56-65",
              is.finite(age) ~ "66+",
              TRUE ~ "unknown"
          ), levels = c(
                 "<=5",
                 "6-10",
                 "11-17",
                 "18-25",
                 "26-35",
                 "36-45",
                 "46-55",
                 "56-65",
                 "66+",
                 "unknown"
             )),
      age_class_plot = factor(
          age_class,
          levels = rev(levels(age_class))))

## check content
x %>%
  group_by(age_class) %>%
  summarise(n = n()) %>%
  adorn_totals("row") %>%
  show_table()

```




### Reporting delay

```{r reporting_delay}

x <- x %>%
  mutate(reporting_delay = as.integer(date_report - date_onset))

```


## Duration of hospitalisation

We add the duration of hospitalisation:

```{r duration_hospitalisation}

## rename dates of death and discharge
x <- x %>%
      # delay_outcome = as.integer(date_outcome - date_onset),
      mutate(delay_admission = as.integer(cte_date - date_onset))
      ## duration_hospitalisation = as.integer(
      ##     date_outcome - date_admission),
      ## documented_hospitalisation = !is.na(date_admission)
  

```


## Healthcare workers

```{r hcw}

table(x$hcw, useNA = "ifany")

```


## Reorder columns

```{r reorder}

x <- x %>%
  select(mllid,
         date_onset,
         date_report,
         date_admission,
         ## date_outcome,
         ## date_discharge,
         date_death,
         ## date_outcome,
         date_exposure_start,
         date_exposure_end,
         province,
         sous_coordination,
         zone_de_sante,
         aire_de_sante,
         age,
         age_class,
         hcw,
         ## delay_outcome,
         ## duration_hospitalisation,
         ## documented_hospitalisation,
         everything())

```





# Comparison with previous databases {.tabset .tabset-fade .tabset-pills}

## Outline

In this section, we compare the newly cleaned database to a previous one, used
as reference. The things that are compared include:

* the **dimensions** of the data (number of rows and columns)
* the **variable names**
* the **type** of variables present on both data


## Load reference database

We import the reference data based on older data (cleaned as well); this part
should be regularly updated so that the reference is a decently recent version
of linelist:

```{r load_ref, eval = FALSE}

previous_clean_data
check_file(previous_clean_data)
x_ref <- rio::import(previous_clean_data)
x_ref

```

## Compare cases counts

```{r previous_database_date}

## extract some dates
last_date_onset <- max(x$date_onset, na.rm= TRUE)
last_date_report <- max(x$date_report, na.rm= TRUE)


## compute respective case counts
to_keep <- c("confirmed", "probable")

case_count <- x %>% 
  dplyr::filter(epicasedef %in% to_keep) %>% 
  nrow()

```




<font color="#0f71a2">**Previous database:**</font>

<br>

<font color="#10a35e">**Current database:**</font> 

* Total rows: 
* Completition date: `r format(database_date, format =
"%A %d %b %Y")`
* Last date of onset:  `r format(last_date_onset, format =
"%A %d %b %Y")`.
* Last date of report: `r format(last_date_report, format =
"%A %d %b %Y")`.
* case counts: `r case_count` (confirmed and probable cases)






## Compare databases

```{r compared_data, eval = FALSE}

compare <- compare_data(x_ref, x,
                        columns = c(
                            "id",
                            "date_onset",
                            "date_report",
                            "date_admission",
                            "date_outcome",
                            "date_death",
                            "date_exposure_start",
                            "date_exposure_end",
                            "province",
                            "sous_coordination",
                            "zone_de_sante",
                            "aire_de_sante",
                            "age",
                            "age_class",
                            "gender",
                            "hcw"))

print(compare, diff_only = TRUE)



```




# Looking for duplicates {.tabset .tabset-fade .tabset-pills}

## Outline

In this section we identify duplicated identifiers, isolate the corresponding
entries and output a data table.


## Finding duplicates

```{r duplicates}
dup2 <- x %>% 
  filter(!is.na(vhf_code))


duplicates <- dup2$vhf_code[duplicated(dup2$vhf_code)]

duplicates 

```

There are `r length(duplicates)` duplicates in the data, corresponding to 
`r sum(x$id %in% duplicates)` rows in the linelist.


## Table of duplicates

```{r duplicates_table}

duplicates_table <- x %>%
  filter(x$vhf_code %in% duplicates) %>%
  select(vhf_code,
         surname,
         epicasedef,
         gender,
         age,
         date_report,
         date_onset,
         everything()) %>% 
  arrange(vhf_code,
          surname,
          epicasedef,
          gender,
          age,
          date_report,
          date_onset)

if (nrow(duplicates_table) > 0) {
  duplicates_table %>%
    show_table()
}

```




# Export data {.tabset .tabset-fade .tabset-pills}

## R objects

We export some of the clean database, placed in `produced_rds/` as well as in
`data/clean/`:

```{r export_rds}

## check if a directory exists and if not then creates it
if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}

## create the text for the file name with the database date
rds_file_name <- sprintf("%sclean_%s.rds",
                     undated_file_name(current_data),
                     format(database_date, "%Y-%m-%d"))
rds_file_name

## save the rds file in the produced_rds folder
rio::export(x,
            file.path("produced_rds", rds_file_name))

```

We copy these files to the `data/clean` folder:

```{r copy_rds}
# copy some files into `data/clean/`

# Provide the destination of where to copy the data
destination <- here("data",
                    "clean",
                    rds_file_name)
# Copy the rds data
file.copy(from = file.path("produced_rds", rds_file_name),
          to = destination,
          overwrite = TRUE)

```



## Excel spreadsheets

We export some of the clean database, placed in `produced_xlsx/` as well as in
`data/clean/`:

```{r export_excel}

## check if a directory exists and if not then creates it
if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

## create the text for the file name with the database date
xlsx_file_name <- sprintf("%sclean_%s.xlsx",
                     undated_file_name(current_data),
                     format(database_date, "%Y-%m-%d"))
xlsx_file_name

## save the excel file in the produced_xlsx folder

rio::export(x,
            file.path("produced_xlsx", xlsx_file_name))


## same procedure for duplicates table
dupli_file_name <- sprintf("%sduplicates_%s.xlsx",
                     undated_file_name(current_data),
                     format(database_date, "%Y-%m-%d"))
dupli_file_name

## save the excel file in the produced_rds folder

rio::export(duplicates_table,
            file.path("produced_xlsx", dupli_file_name))


```

We copy these files to the `data/clean` folder:

```{r copy_excel}

## copy the clean xlsx file to `data/clean/`
destination <- here("data",
                    "clean",
                    xlsx_file_name)
file.copy(from = file.path("produced_xlsx", xlsx_file_name),
          to = destination,
          overwrite = TRUE)


## copy the xlsx of duplicates file to `data/duplicates/`
destination <- here("data",
                    "duplicates",
                    dupli_file_name)
file.copy(from = file.path("produced_xlsx", dupli_file_name),
          to = destination,
          overwrite = TRUE)

```


```

## Links to file

Click on the link below to open the file:

- [`r file_name`](`r file.path("produced_xlsx", xlsx_file_name)`)
- [`r dupli_file_name`](`r file.path("produced_xlsx", dupli_file_name)`)



## Update the `current_clean_data.R` script

```{r update_script}

## path to the output file 
script_destination <- here::here("scripts",
                                 "current_clean_data.R")

## comments to say when this was updated
txt <- paste("## This file is generated automatically by `aaa_clean_linelist`",
             "## Do not edit it by hand!\n",
             sep = "\n")
cat(txt, file = script_destination, append = FALSE)

txt <- sprintf("\n## This file was last updated on the: %s\n",
               Sys.time())
cat(txt, file = script_destination, append = TRUE)

## actual content of the script
txt <- sprintf('\ncurrent_clean_data <- here::here("data",
                                 "clean",
                                 "%s")',
               rds_file_name)
cat(txt, file = script_destination, append = TRUE)

```


# System information {.tabset .tabset-fade .tabset-pills}

The following information documents the system on which the document was
compiled.

## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```

