---
title: "Incidence by aire de santé"
author: "Madeleine Crowe, Flavio Finger, Thibaut Jombart, and Jonathan Polonsky for the epi analytic cell North Kivu"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    toc: TRUE
    toc_depth: 4
    toc_float: true
    toc_collapse: false
    number_sections: true
    highlight: pygments
    theme: spacelab
    code_folding: hide
    css: !expr here::here('css', 'style.css')
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE)
```



<!-- ================================= -->
<!-- ================================= -->
<!-- ================================= -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This is the universal cleaning script for master linelist database.

### Data used

Input data is the master linelist with one row per case. 
**All input files must be `xlsx` files.**

### Method

The data preparation involves the following steps, detailed in the following tabs:

* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded; scripts include the data-cleaning dictionary and the paths to the
  current data

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.


## Load scripts

These scripts will load:

* required packages
* a few helper functions
* a data-cleaning dictionary (`cleaning_rules`)
* the path to current version of raw data (`current_data`)

```{r read_scripts}
## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names = TRUE)
for (file in scripts_files) source(file, local = TRUE)


ggthemr("pale")
```



## Load data

```{r load_data}

current_clean_data
x <- rio::import(current_clean_data)

file_name <- gsub("^[^.]+/", "", current_clean_data)
database_date <- file_name %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.


## Subsetting cases

We subset data to retain only confirmed and probable cases.

```{r subset cases}

x <- x %>%
  filter(epicasedef %in% c("confirmed", "probable"))

```

## Matching between health area and sub-coordination

The correspondance between health areas and sub-coordination will be needed recover the sub-coordinations information after stratifying by health areas. We define the matching based on the most frequent correspondence between health areas and sub-coordinations.


```{r health area sub coord}

aire_de_sante_info <- x %>% 
  group_by(aire_de_sante, sous_coordination) %>% 
  count() %>% 
  group_by(aire_de_sante) %>% 
  top_n(1, n) %>%
  select(-n)

aire_de_sante_info %>%
  show_table()

```

<!-- ================================= -->
<!-- ================================= -->
<!-- ================================= -->

# Analyses {.tabset .tabset-fade .tabset-pills}

## Computer incidence by health area

Daily incidence is computed for health areas using `incidence`, and will later be converted to a `data.frame` for further analyses:

```{r incidence by health area}

inc <- with(x, incidence(date_report, groups = aire_de_sante))

```


## Active health areas per day

Here we compute the number of active health areas per day, for the entire time period of the data. Active areas are defined as health areas with at least 1 case within the last 21 days of the date considered. We do this by using the `rollsum` function, which computes the cumulative number of cases/aire de sante for a given time window (here, 21 days). Note that the argument `align = "right"` specifies that the time window, on a given day, should be in the past, not in the future. The argument `fill = NA` is also (weirdly) needed for this to work.


```{r active health areas}

inc_df <- as.data.frame(inc, long = TRUE) %>% 
  rename(aire_de_sante = groups) %>% 
  left_join(aire_de_sante_info) 

timewindow <- 21

nb_as_active <- inc_df %>%
  group_by(aire_de_sante) %>%
  mutate(
      cases_last_21_days =zoo::rollsum(counts,
                                   timewindow,
                                   align = "right",
                                   fill = NA),
    has_cases = (cases_last_21_days > 0)
     ) %>%
  group_by(dates, sous_coordination) %>% 
  summarise(n_active_areas = sum(has_cases))
 

ggplot(nb_as_active) +
  geom_col(aes(x = dates, y = n_active_areas,
               fill = sous_coordination),
           width = 1) +
  labs(title = "Nombre d'aires de santé actives au cours du temps",
       x = "",
       y = paste("Nombre d'aires de santé avec au moins un cas",
                 "notifié dans les 21 derniers jours",
                 sep = "\n")) +
  scale_x_date(date_breaks = "1 month") +
  theme(legend.position = c(0.25, 0.85)) +
  scale_fill_discrete("Sous-coordination") +
  guides(fill = guide_legend(ncol = 2)) +
  large_txt +
  rotate_x_text(45)

```

```{r number as active}
nb_as_active %>% 
  group_by(dates, sous_coordination) %>% 
  summarise(n = sum(n_active_areas, na.rm = TRUE)) %>% 
  filter(!is.na(n), !sous_coordination %in% "other_unknown") %>%
  ggplot() +
  aes(x = dates, y = n, fill = sous_coordination, width = 1) +
  geom_col() +
  labs(
    title = "Nombre d'aires de santé actives au cours du temps",
    x = "",
    y = 
      paste(
        "Nombre d'aires de santé avec au moins un cas",
        "notifié dans les 21 derniers jours",
        sep = "\n"
      )
  ) +
  scale_x_date(date_breaks = "2 month") +
  theme(legend.position = "none") +
  scale_fill_discrete("Sous-coordination") +
  guides(fill = guide_legend(ncol = 2)) +
  large_txt +
  rotate_x_text(45) +
  # theme(axis.text.x = element_text(size = 6, angle = 90, hjust = .5)) +
  facet_wrap(~ sous_coordination)
```

```{r as active table}
nb_as_active %>%
  show_table()

```

## Current active areas - 21 days
In this part we identify health areas active on the date of completion of the database, and report the number of cases over the last 21 days, i.e. from the **`r database_date-21`** to the **`r database_date`**.


```{r current active areas}

current_active <- x %>%
  group_by(aire_de_sante) %>%
  summarise(n_cases_21_days = sum(date_report > (database_date - 20))) %>%
  filter(n_cases_21_days > 0) %>% 
  left_join(aire_de_sante_info)


ggplot(current_active, aes(x = aire_de_sante, y = n_cases_21_days)) +
  geom_col(aes(fill = sous_coordination)) +
  facet_grid(. ~ sous_coordination,
             scales = "free_x",
             space = "free") +
  custom_horiz_facet +
  large_txt +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1,
                                   size = 11)) +
  scale_fill_discrete("") +
  labs(title = sprintf("Aires de santé actives au %s",
                       format(database_date, "%d %b %Y")),
       x = "",
       y = "Nombre de cas au cours des 21 derniers jours")

```


```{r current active table}

current_active %>%
  show_table()

```

## Current active areas - 7 days
In this part we identify health areas active on the date of completion of the database, and report the number of cases over the last 7 days, i.e. from the **`r database_date-7`** to the **`r database_date`**.

```{r current active areas past 7 days}

current_active_7_days <- x %>%
  group_by(aire_de_sante) %>%
  summarise(n_cases_7_days = sum(date_report > (database_date - 7))) %>%
  filter(n_cases_7_days > 0) %>% 
  left_join(aire_de_sante_info)


ggplot(current_active_7_days, aes(x = aire_de_sante, y = n_cases_7_days)) +
  geom_col(aes(fill = sous_coordination)) +
  facet_grid(. ~ sous_coordination,
             scales = "free_x",
             space = "free") +
  custom_horiz_facet +
  large_txt +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1,
                                   size = 11)) +
  scale_fill_discrete("") +
  labs(title = sprintf("Aires de santé actives au %s",
                       format(database_date, "%d %b %Y")),
       x = "",
       y = "Nombre de cas au cours des 7 derniers jours")
```


```{r current active table 7 days}

current_active_7_days %>%
  show_table()

```


```


# Export data 

We export some of the relevant tables, which will be placed in the current working directory:

```{r export_xlsx}

if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

rio::export(inc_df,
            file.path("produced_xlsx",
                      "incidence_data_frame.xlsx"))

rio::export(nb_as_active,
            file.path("produced_xlsx",
                      "active_health_areas_21_days.xlsx"))

rio::export(current_active,
            file.path("produced_xlsx",
                      "current_active_health_areas.xlsx"))

```




# System information {.tabset .tabset-fade .tabset-pills}

The following information documents the system on which the document was
compiled.

## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```
