---
title: "Age sex distributions"
author: "Christopher Jarvis, Thibaut Jombart, Amy Gimma, Arthur Fitzmaurice for the analytic cell EOC Goma"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 2
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE)
```



<br>

<div class="report_meta">
  <span class="notice">**Notice**: this is a **stable, routine report**. 
  **Do not touch it unless it is broken.** To make a contribution, **carefully read 
  the [README](../../../../../README.html) file**.</span>
  
  **Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)
  
  **Code contributors:** Christopher Jarvis, Thibaut Jombart, Amy Gimma, Arthur Fitzmaurice
  
  **Data contributors:** Yannick Tutu, Richie Ngombo
  
  **Version:** 2.1.0
  
  **Reviewed by:** Thibaut Jombart
</div>






<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Data preparation {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

This report summarise the age and sex distribution of cases overall, and
stratified by geographic unit.

### Data used

The master linelist.

### Method

The data preparation involves the following steps, detailed in the following tabs:


* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Clean data**: this section contains *ad hoc* data cleaning, i.e. which is
  not used in other reports (otherwise cleaning should be done in a dedicated
  report); this section is also used to create new variables used in the
  analyses




<!-- ======================================================= -->
## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned VHF data stored as `x`

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$",
                     full.names = TRUE)
for (file in scripts_files) source(file)

ggthemr('grape')

```



<!-- ======================================================= -->
## Load data

We extract the completion date from the file name:

```{r load_data}

## load the data
current_clean_data
x <- rio::import(current_clean_data)

## extract database date from the file name
file_name <- gsub("^[^.]+/", "", current_clean_data)
database_date <- file_name %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.





<!-- ======================================================= -->
## Add variables

We add the following new variables 

* `recent`: cases reported within the last 42 days (from the database date)
* `top_zones`: 7 zones who reported the most cases over the last 42 days


```{r add_top_zones}

start_at <- database_date - 42

x <- x %>%
  mutate(recent = date_report >= start_at, 
         top_zones = top_values(zone_de_sante, 7, subset = recent))

```
  



<!-- ======================================================= -->
## Filter data

We retain only entries:

* corresponding to confirmed and probable cases
* whose sex is known
* whose age class is known
* from 2019-01-01 to present

```{r filter}

x <- x %>%
  filter(epicasedef %in% c("confirmed", "probable"),
         sex %in% c("male", "female"),
         age_class != "unknown",
         month_report >= as.Date("2019-01-01"))


```



## Custom colors

In this part we define custom color palettes for further plotting.

```{r custom_colors}

scale_sex <- scale_fill_manual(name = "Sexe", 
                               values = c(male = "#5DADE2",
                                          female = "#F08080"),
                               labels = c(male = "homme",
                                          female = "femme"))

```




<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Age-sex pyramids {.tabset .tabset-fade .tabset-pills}

## Outline 

We build age-sex pyramids for confirmed and probable cases, overall, and then
stratified.




<!-- ======================================================= -->
## Overall distribution

```{r overall}

counts_overall <- x %>% 
  group_by(sex, age_class) %>%
  count()

ggplot(counts_overall,
       aes(x = age_class,
           fill = sex)) +
  geom_bar(stat = "identity",
           size = 1,
           aes(y = ifelse(sex == "female", n, -n))) +
  theme_bw() +
  scale_sex +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  large_txt +
  labs(x = "",
       y = "Nombre de cas",
       title = "Distribution par 창ge des cas")

tab_overall <- counts_overall %>% 
  spread(key = "age_class", value = "n", fill = 0)

tab_overall %>%
  show_table()

```




<!-- ======================================================= -->
## By sub-coordination

```{r sub_coordination}

counts_sub_coordination <- x %>% 
  group_by(sous_coordination, sex, age_class) %>%
  count()

ggplot(counts_sub_coordination,
       aes(x = age_class,
           fill = sex)) +
  geom_bar(stat = "identity",
           size = 1,
           aes(y = ifelse(sex == "female", n, -n))) +
  theme_bw() +
  scale_sex +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  large_txt +
  facet_wrap(~ sous_coordination, scales = "free_x") +
  custom_vert_facet +
  labs(x = "",
       y = "Nombre de cas",
       title = "Distribution par 창ge des cas")

tab_sub_coordination <- counts_sub_coordination %>% 
  spread(key = "age_class", value = "n", fill = 0)

tab_sub_coordination %>%
  show_table()

```





<!-- ======================================================= -->
## By recent major health zone

```{r top_zones}

counts_top_zones <- x %>% 
  group_by(top_zones, sex, age_class) %>%
  count()

ggplot(counts_top_zones,
       aes(x = age_class,
           fill = sex)) +
  geom_bar(stat = "identity",
           size = 1,
           aes(y = ifelse(sex == "female", n, -n))) +
  theme_bw() +
  scale_sex +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  large_txt +
  facet_wrap(~ top_zones, scales = "free_x") +
  custom_vert_facet +
  labs(x = "",
       y = "Nombre de cas",
       title = "Distribution par 창ge des cas")

tab_top_zones <- counts_top_zones %>% 
  spread(key = "age_class", value = "n", fill = 0)

tab_top_zones %>%
  show_table()

```




<!-- ======================================================= -->
## By month

```{r month_report}

counts_month_report <- x %>% 
  group_by(month_report, sex, age_class) %>%
  count()

ggplot(counts_month_report,
       aes(x = age_class,
           fill = sex)) +
  geom_bar(stat = "identity",
           size = 1,
           aes(y = ifelse(sex == "female", n, -n))) +
  theme_bw() +
  scale_sex +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  large_txt +
  facet_wrap(~ month_report, scales = "free_x") +
  custom_vert_facet +
  labs(x = "",
       y = "Nombre de cas",
       title = "Distribution par 창ge des cas")

tab_month_report <- counts_month_report %>% 
  spread(key = "age_class", value = "n", fill = 0)

tab_month_report %>%
  show_table()

```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Export tables {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

Provide an outline of exported files.


<!-- ======================================================= -->
## Excel files

Explain what the exported files are using the following model:

* `produced_xlsx/tab_overall.xlsx`: tab_overall contains the overall distributions of cases by age group and sex.

* `produced_xlsx/tab_sub_coordination.xlsx`: tab_sub_coordination contains the overall distributions of 
cases by age group, sex, and sous coordination.

* `produced_xlsx/tab_top_zones.xlsx`: tab_top_zones contains the overall distributions of cases by age group and sex 
in health zones with the most cases during the last 42 days.

* `produced_xlsx/tab_month_report.xlsx`: tab_month_report contains the overall distributions of cases by 
age group, sex, and month of report. 


```{r xlsx_exports}
if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

rio::export(tab_overall,
    file.path("produced_xlsx", "tab_overall.xlsx"))
rio::export(tab_sub_coordination,
    file.path("produced_xlsx", "tab_sub_coordination.xlsx"))
rio::export(tab_top_zones,
    file.path("produced_xlsx", "tab_top_zones.xlsx"))
rio::export(tab_month_report,
    file.path("produced_xlsx", "tab_month_report.xlsx"))

```

Click on the following links to open the files (only works if the files above
have been generated and are in the same folder as this document):

- [tab_overal.xlsx](produced_xlsx/tab_overall.xlsx)
- [tab_sub_coordination.xlsx](produced_xlsx/tab_sub_coordination.xlsx)
- [tab_top_zones.xlsx](produced_xlsx/tab_top_zones.xlsx)
- [tab_month_report.xlsx](produced_xlsx/tab_month_report.xlsx)




<!-- ======================================================= -->
## R objects

Explain what the exported files are using the following model:


* `produced_rds/tab_overall.rds`: tab_overall contains the overall distributions of cases by age group and sex.

* `produced_rds/tab_sub_coordination.rds`: tab_sub_coordination contains the overall distributions of 
cases by age group, sex, and sous coordination. 

* `produced_rds/tab_top_zones.rds`: tab_top_zones contains the overall distributions of cases by age group and sex 
in health zones with the most cases during the last 42 days. 

* `produced_rds/tab_month_report.rds`: tab_month_report contains the overall distributions of cases by 
age group, sex, and month of report. 


```{r rds_exports}
# 
if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}

rio::export(tab_overall,
            file.path("produced_rds", "tab_overall.rds"))
rio::export(tab_sub_coordination,
            file.path("produced_rds", "tab_sub_coordination.rds"))
rio::export(tab_top_zones,
            file.path("produced_rds", "tab_top_zones.rds"))
rio::export(tab_month_report,
            file.path("produced_rds", "tab_month_report.rds"))

```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# System information {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

The following information documents the system on which the document was
compiled.



<!-- ======================================================= -->
## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```



<!-- ======================================================= -->
## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```



<!-- ======================================================= -->
## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```




<!-- ===================================================== -->
## Compilation parameters

This shows which parameters were passed through `params` at compilation time:

```{r params}
params
```



<!-- ===================================================== -->
## Change log

### Version 1.0.0

