---
title: "Key performance indicators for the DG weekly brief"
author: "Jonathan Polonsky, Patrick Keating, Thibaut Jombart for the epi analysis cell - EOC Goma"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    toc: TRUE
    toc_depth: 4
    toc_float: true
    toc_collapse: false
    number_sections: true
    highlight: pygments
    theme: spacelab
    code_folding: hide
    css: !expr here::here('css', 'style.css')
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  fig.width = 8,
  fig.height = 6,
  dpi = 150,
  warning = FALSE,
  message = FALSE
)
```


<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This report prepares a table of KPIs for the DG weekly brief

### Data used

The master linelist.

### Method

The data preparation involves the following steps, detailed in the following tabs:


* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Clean data**: this section contains *ad hoc* data cleaning, i.e. which is
  not used in other reports (otherwise cleaning should be done in a dedicated
  report); this section is also used to create new variables used in the
  analyses




## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned VHF data stored as `x`

```{r read_scripts}
## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names = TRUE)
for (file in scripts_files) source(file, local = TRUE)

```


## Load data

We extract the completion date from the file name:

```{r load_data}

## load the data
current_clean_data
x <- rio::import(current_clean_data)

## extract database date from the file name
database_date <- current_clean_data %>% 
  guess_dates()

database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.

## Clean data

Some optional data cleaning here, including the possible addition of new
variables used in the analyses.

### Data preparation

We filter the data to include only the most recent 6 complete epiweeks of data,
using dynamically created variables based on the database date.

We also add a new variable `top_zones` which retains the top 6 health zones (with
most cases since the beginning of the outbreak), and pool everything else into
'others':


```{r prep}

recent_week <- 
  ifelse(
    weekdays(database_date) %in% "Sunday", 
    aweek::date2week(database_date, floor_day = TRUE), 
    aweek::date2week(database_date - 7, floor_day = TRUE)
  )

start_week <-
  (aweek::week2date(recent_week) - (7 * 5)) %>% 
  aweek::date2week(floor_day = TRUE) %>% 
  as.character()

df_kpi <-
  x %>% filter(
            epiweek_report >= start_week,
            epiweek_report <= recent_week) %>% 
  mutate(top_zones = top_values(zone_de_sante, 6),
         registered_surveilled = 
           contact_registered == "yes" &
           contact_surveilled == "yes",
         delay_admission =
           if_else(delay_admission < 0,
                   NA_integer_,
                   delay_admission)) %>% 
  select(top_zones,
         epiweek_report,
         type_death,
         registered_surveilled,
         delay_admission)

## TODO: check if this is relevant
zs_order <- df_kpi$top_zones %>% levels()

```

## Functions

Here we define some functions to be used in the analysis:

```{r functions}

make_case_counts_total <- function(dat, zs) {
  
  df_crossing <- 
    df_kpi %>% 
    expand(top_zones, epiweek_report) %>% 
    filter(top_zones %in% zs)
  
  dat %>% 
    count(epiweek_report) %>% 
    right_join(df_crossing) %>% 
    replace_na(list(n = 0)) %>% 
    full_join(
      dat %>% 
        count() %>% 
        mutate(epiweek_report = "Overall")
    )
  
}

make_case_counts_var <- function(dat, zs, col_var) {
  
  col_var <- enquo(col_var)
  
  df_crossing <- 
    df_kpi %>% 
    expand(top_zones, epiweek_report, !!col_var) %>% 
    filter(top_zones %in% zs)
  
  dat %>%
    count(epiweek_report, !!col_var) %>% 
    right_join(df_crossing) %>% 
    replace_na(list(n = 0)) %>% 
    group_by(epiweek_report) %>% 
    mutate(freq = (n / sum(n) * 100) %>% round(1)) %>% 
    full_join(
      dat %>% 
        count(!!col_var) %>% 
        mutate(freq = (n / sum(n) * 100) %>% round(1)) %>% 
        mutate(epiweek_report = "Overall")
    ) %>% 
    filter(!!col_var %in% TRUE) %>% 
    mutate(!!col_var := str_c(n, " (", freq, "%)")) %>% 
    select(1:2)
  
}

make_delay <- function(dat, zs) {
  
  df_crossing <- 
    df_kpi %>% 
    expand(top_zones, epiweek_report) %>% 
    filter(top_zones %in% zs)
  
  dat %>% 
    right_join(df_crossing) %>% 
    replace_na(list(n = 0)) %>% 
    group_by(epiweek_report) %>%
    summarise(
      delay_med = median(delay_admission, na.rm = TRUE), 
      delay_low = quantile(delay_admission, probs = 0.25, na.rm = TRUE),
      delay_upp = quantile(delay_admission, probs = 0.75, na.rm = TRUE)
    ) %>% 
    full_join(
      dat %>% 
        summarise(
          delay_med = median(delay_admission, na.rm = TRUE), 
          delay_low = quantile(delay_admission, probs = 0.25, na.rm = TRUE),
          delay_upp = quantile(delay_admission, probs = 0.75, na.rm = TRUE)
        ) %>% 
        mutate(epiweek_report = "Overall")
    ) %>% 
    mutate(delay = str_c(delay_med, " [", delay_low, " - ", delay_upp, "]")) %>% 
    select(-c(2:4))
  
}  

show_kpi_tbl <- function(num) {
  
  tbl_kpi_trend[[num]] %>%
    knitr::kable(
      align = c("l", rep("r", 4)),
      caption = names(tbl_kpi_trend)[num]
    ) %>% 
    kable_styling(bootstrap_options = c("striped", "hover", "responsive"), font_size = 14) %>% 
    row_spec(nrow(tbl_kpi_trend[[num]]), bold = TRUE) %>% 
    column_spec(1, bold = TRUE)
  
}

```



## Analyses

Now we map these functions over a nested tibble to analyse the data by top zone: 

```{r calculations}

## df_kpi_nest <- 
##   df_kpi %>% 
##   group_by(top_zones) %>% 
##   nest()

## !! THIS CODE NO LONGER WORKS; COMPLAINS ABOUT TOP_ZONES USED AS GROUPING AND
## !! CANNOT BE ALTERED
## Renaming the variable triggers other errors. Code too obscure (no
## documentation of the functions) to be fixed quickly. This needs an extensive
## review.

## tbl_case_counts <- 
##   df_kpi_nest %>% 
##   mutate(
##     case_counts_total = pmap(list(data, top_zones), ~make_case_counts_total(.x, .y)),
##     case_counts_comm_deaths =
##       pmap(list(data, top_zones), ~make_case_counts_var(.x, .y, type_death)),
##     case_counts_registered_surveilled =
##       pmap(list(data, top_zones), ~make_case_counts_var(.x, .y, registered_surveilled)),
##     delay = pmap(list(data, top_zones), ~make_delay(.x, .y)))%>% #,
## ## top_zones = top_zones %>% fct_relevel(zs_order)
##   #) %>% 
##   select(-data) %>%
##   unnest %>% 
##   select(-matches("[0-9]$")) %>% 
##   set_names(
##       c(#"Zone de santé",
##         "Epiweek",
##         "# Cases reported",
##         "# Community deaths (%)",
##         "% Cases from contact lists and under follow up at time of detection",
##         "Median [IQR] days from onset to admission to ETC")
##   )



## !! This is a temporary patch. Need to overhaul the entire report.

tbl_case_counts <- df_kpi %>% 
  group_by(top_zones, epiweek_report) %>%
  summarise(n_cases = n(),
            n_comm_deaths = sum(type_death == "community"),
            denom_comm_deaths = sum(type_death != "unknown"),
            n_registered_surveilled = sum(registered_surveilled, na.rm = TRUE),
            denom_registered_surveilled = sum(!is.na(registered_surveilled)),
            median_delay_admission = median(delay_admission, na.rm = TRUE),
            q25_delay_admission = quantile(delay_admission, .25, na.rm = TRUE),
            q75_delay_admission = quantile(delay_admission, .75, na.rm = TRUE)
            ) %>%
  mutate("Zone de santé" = top_zones,
         "Epiweek" = epiweek_report,
         "# Cases reported" = n_cases,
         "# Community deaths (%)" =
           sprintf("%i (%.1f%%)",
                   n_comm_deaths,
                   prop_to_perc(n_comm_deaths / n_cases)), # to fix once the outcome data is fixed!!
         "% Cases from contact lists and under follow up at time of detection" =
           sprintf("%i (%.1f%%)",
                   n_registered_surveilled,
                   prop_to_perc(n_registered_surveilled / n_cases)), # to fix once denom fixed
         "Median [IQR] days from onset to admission to ETC" =
           sprintf("%.1f [%.1f-%.1f])",
                   median_delay_admission,
                   round(q25_delay_admission, 1),
                   round(q75_delay_admission, 1)
                   )
         ) %>%
  select("Zone de santé",
         "Epiweek",
         "# Cases reported",
         "# Community deaths (%)",
         "% Cases from contact lists and under follow up at time of detection",
         "Median [IQR] days from onset to admission to ETC")


```



<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# KPIs for weekly DG brief
## Outline 

Here we present a summary of the epidemiological KPIs for the DG weekly brief


## Snapshot of 6 most active health zones for epiweek `r recent_week`

First we filter the data to provide a summary of the most recent 3 complete epiweeks.

```{r overview_text}

count_cases <- nrow(x)

penultimate_epiweek_1 <- 
  (recent_week %>% aweek::week2date() - 1) %>% 
  aweek::date2week(floor_day = TRUE)

penultimate_epiweek_2 <- 
  (recent_week %>% aweek::week2date() - 8) %>% 
  aweek::date2week(floor_day = TRUE)

tbl_recent_cases <- 
  x %>% 
  filter(epiweek_report %in% c(recent_week, penultimate_epiweek_1, penultimate_epiweek_2)) %>% 
  count(epiweek_report) %>% 
  arrange(epiweek_report)

recent_week_start <- recent_week %>% aweek::week2date() %>% format("%d %b %Y")
recent_week_end <- (recent_week %>% aweek::week2date() + 6) %>% format("%d %b %Y")

epiweek_sentence <- 
  str_c(
    "For the epiweek ", recent_week, " (", recent_week_start, " - ", recent_week_end, "):\n",
    tbl_recent_cases$n[3], " cases were reported, compared to ", tbl_recent_cases$n[2], " cases in ", tbl_recent_cases$epiweek_report[2], ", and ", tbl_recent_cases$n[1], " cases in ", tbl_recent_cases$epiweek_report[1], "." 
    )

```


```{r community deaths counts and health zone counts}

comm_deathp <- x %>%
  select(type_death, epiweek_report) %>% 
  group_by_all() %>% 
  tally() %>% 
  spread(type_death, n, fill = 0) %>% 
  adorn_totals("col") %>% 
  mutate(prop_comm_death = prop_to_perc(community / Total))

penultimate_cd <-
  comm_deathp %>% 
  filter(epiweek_report == penultimate_epiweek_1) %>% 
  select(prop_comm_death) %>% 
  round() %>% 
  as.integer()

week_before_penultimate_cd <-
  comm_deathp %>% 
  filter(epiweek_report == penultimate_epiweek_2) %>% 
  select(prop_comm_death) %>% 
  round() %>% 
  as.integer()

recent_week_cd <-
  comm_deathp %>% 
  filter(epiweek_report == recent_week) %>% 
  select(prop_comm_death) %>% 
  round() %>% 
  as.integer()

change <- 
  if(penultimate_cd > recent_week_cd) {
    print("decreased")
  } else if(penultimate_cd == recent_week_cd){
    print("remained the same")
  }else{
    print("increased")
  }


health_zones_counts <- 
  x %>% 
  select(zone_de_sante, epiweek_report) %>% 
  mutate(zone_de_sante = str_to_title(zone_de_sante)) %>% 
  group_by_all() %>% 
  tally() %>% 
  filter(epiweek_report %in% c(recent_week, penultimate_epiweek_1, penultimate_epiweek_2)) %>% 
  spread(epiweek_report, n, fill = 0) %>%
  adorn_totals("col") %>% 
  arrange(desc(Total)) %>% 
  mutate(total_cases_all_zs = sum(Total)) %>% 
  mutate(prop = prop_to_perc(Total/total_cases_all_zs),
         prop = round(prop),
         prop = as.integer(prop)) 

# health_zone_counts_21_days <-
#   x %>%
#   select(zone_de_sante, date_report) %>% 
#   group_by_all() %>% 
#   tally() %>%
#   filter(date_report > (max(x$date_report) - 20)) %>%
#   spread(date_report, n, fill = 0) %>%
#   adorn_totals("col") %>%
#   arrange(desc(Total)) %>%
#   select(zone_de_sante, `2019-08-26`, `2019-08-27`, `2019-08-28`, `2019-08-29`, Total) %>%
#   mutate(total_4_days = `2019-08-26` + `2019-08-27` + `2019-08-28` + `2019-08-29`)
# 
# sum(health_zone_counts_21_days$Total)
# sum(health_zone_counts_21_days$total_4_days)

x %>%
  group_by(zone_de_sante) %>%
  count() %>%
  adorn_totals()

```

```{r active health areas in previous two weeks, include=FALSE}
# Date of penultimate week
penultimate_epiweek_1_date <- 
  (recent_week %>% aweek::week2date() - 1)


# Match health areas with subcoordination
aire_de_sante_info <-
  x %>% 
  group_by(aire_de_sante, sous_coordination) %>% 
  count() %>% 
  group_by(aire_de_sante) %>% 
  top_n(1, n) %>%
  select(-n)


# count active health areas in recent week (previous 21 days including the last week)
current_active <- 
  x %>%
  group_by(aire_de_sante) %>%
  summarise(n_cases_21_days = sum(date_report > (aweek::week2date(recent_week) - 17))) %>%
  filter(n_cases_21_days > 0) %>% 
  left_join(aire_de_sante_info)


# count active health areas in previous week (21 days from the previous week)
penultimate_week_active <- 
  x %>%
  group_by(aire_de_sante) %>%
  filter(date_report < aweek::week2date(recent_week)) %>% 
  summarise(n_cases_21_days = sum(date_report > (aweek::week2date(penultimate_epiweek_1_date) - 21))) %>%
  filter(n_cases_21_days > 0) %>% 
  left_join(aire_de_sante_info)


## New health areas
new_as <- current_active[!current_active$aire_de_sante %in% penultimate_week_active$aire_de_sante ,]

new_as

```


```{r hcw}

hcw_count <- 
x %>% 
  filter(epiweek_report %in% recent_week) %>% 
  filter(hcw == "hcw") %>% 
  tally()

hcw_count

# total HCW infections. 

x %>% 
  filter(hcw == "hcw") %>% 
  tally()

```


  * As of `r format(database_date, "%d %b %Y")`, the total number of cases was `r nrow(x)`. 
  * For the epidemiological week `r str_sub(max(x$epiweek_report),7)` `r paste0("(",format(database_date -6,"%d"), " - ", format(database_date, "%d %b"),")")`
    * `r nrow(x[x$epiweek_report == recent_week, "epiweek_report"])` cases were reported compared to `r nrow(x[x$epiweek_report == penultimate_epiweek_1, "epiweek_report"])` in week `r str_sub(penultimate_epiweek_1, 7)`, and `r nrow(x[x$epiweek_report == penultimate_epiweek_2, "epiweek_report"])` in week `r str_sub(penultimate_epiweek_2, 7)`
    * Total number of active health areas was `r nrow(current_active)` in week `r str_sub(max(x$epiweek_report),7)` compared to `r nrow(penultimate_week_active)` in week `r str_sub(penultimate_epiweek_1, 7)`
      * There were a total of `r nrow(new_as)` new health areas in week `r str_sub(max(x$epiweek_report),7)`
  * Proportion of community deaths `r change` from `r penultimate_cd `% week `r str_sub(penultimate_epiweek_1, 7)` to `r recent_week_cd`% week `r str_sub(recent_week, 7)`. For week `r str_sub(penultimate_epiweek_2, 7)`, the community death rate was `r week_before_penultimate_cd`%.
  * There were a total of `r hcw_count` new infections among health care workers in week `r str_sub(max(x$epiweek_report),7)`
  * During the last 3 weeks (as of `r format(database_date, "%d %b")`):
    * `r health_zones_counts[1,1]` health zone (HZ) was the main hot spot with `r health_zones_counts[1,7]`%  (`r health_zones_counts[1,5]`/`r health_zones_counts[1,6]`) of total cases over the time period.  
    * `r health_zones_counts[2,1]` HZ was the second hot spot with `r health_zones_counts[2,7]`% (`r health_zones_counts[2,5]`/`r health_zones_counts[2,6]`) of total cases over the time period.
    * `r health_zones_counts[3,1]` HZ was the third hot spot with `r health_zones_counts[3,7]`% (`r health_zones_counts[3,5]`/`r health_zones_counts[3,6]`) of total cases over the time period.  

```{r kpi_tbl_recent_week}

df_kpi_latest <- df_kpi %>% filter(epiweek_report %in% recent_week)

kpi_total_1 <- 
  df_kpi_latest %>% 
  count() %>% 
  mutate(top_zones = "Overall") %>% 
  select(top_zones, everything())

kpi_total_2 <- 
  df_kpi_latest %>% 
  count(type_death) %>% 
  mutate(freq = (n / sum(n) * 100) %>% round(1)) %>% 
  mutate(top_zones = "Overall") %>% 
  filter(type_death == "community") %>% 
  mutate(type_death = str_c(n, " (", freq, "%)")) %>% 
  select(top_zones, type_death)

kpi_total_3 <- 
  df_kpi_latest %>% 
  count(registered_surveilled) %>% 
  mutate(freq = (n / sum(n) * 100) %>% round(1)) %>% 
  mutate(top_zones = "Overall") %>% 
  filter(registered_surveilled) %>% 
  mutate(registered_surveilled = str_c(n, " (", freq, "%)")) %>% 
  select(top_zones, registered_surveilled)

kpi_total_4 <- 
  df_kpi_latest %>% 
  summarise(
    delay_med = median(delay_admission, na.rm = TRUE), 
    delay_low = quantile(delay_admission, probs = 0.25, na.rm = TRUE),
    delay_upp = quantile(delay_admission, probs = 0.75, na.rm = TRUE)
  ) %>% 
  mutate(top_zones = "Overall") %>% 
  mutate(delay = str_c(delay_med, " [", delay_low, " - ", delay_upp, "]")) %>% 
  select(4:5)

kpi_total_all <- 
  kpi_total_1 %>% 
  full_join(kpi_total_2) %>% 
  full_join(kpi_total_3) %>% 
  full_join(kpi_total_4) %>% 
  add_column(epiweek = recent_week) %>% 
  set_names(
    c("Zone de santé", "# Cases reported", "# Community deaths (%)", "% Cases from contact lists and under follow up at time of detection", "Median [IQR] days from onset to admission to ETC", "Epiweek")
  )

tbl_kpi_latest <- 
  tbl_case_counts %>% 
  filter(Epiweek %in% recent_week) %>% 
  bind_rows(kpi_total_all) %>% 
  mutate(`Zone de santé` = `Zone de santé` %>% fct_relevel(c(zs_order, "Overall"))) %>% 
  arrange(`Zone de santé`) %>% 
  select(-Epiweek)

tbl_kpi_latest %>% 
  kable(align = c("l", rep("r", 4))) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), font_size = 14) %>% 
  footnote(
    general = str_c("Data for epiweek ", recent_week),
    title_format = c("bold")
  ) %>% 
  row_spec(nrow(tbl_kpi_latest), bold = TRUE) %>% 
  column_spec(1, bold = TRUE)

```

## Trends for past 6 weeks in most active health zones

```{r kpi_tbl_trend, results="asis"}

tbl_kpi_trend <- 
  tbl_case_counts %>% 
  group_by(`Zone de santé`) %>% 
  group_split(keep = FALSE) %>% 
  set_names(zs_order)

seq_along(tbl_kpi_trend) %>% walk(~ show_kpi_tbl(.x) %>% cat())

```





# System information {.tabset .tabset-fade .tabset-pills}

## Outline

The following information documents the system on which the document was
compiled.



## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```
