---
title: "Estimating transmission intensity"
author: "Thibaut Jombart and Christopher Jarvis for the analytic cell EOC North Kivu"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    toc: TRUE
    toc_depth: 4
    toc_float: true
    toc_collapse: false
    number_sections: true
    highlight: pygments
    theme: spacelab
    code_folding: hide
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 10,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE)
```



<br>

**Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)

**Code contributors:** Thibaut Jombart, Chris Jarvis

**Data contributors:** Richy Ngombo, Yannick Tutu

**Version:** 1.0.0

**Reviewed by:** 

**Notice**: this is a **stable, routine report**. **Do not touch it unless it is
broken.** To make a contribution, carefully read the
[README](../../../../../README.html) file.




<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Data preparation {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

This report produces estimates of a infection intensity based on the expected
number of secondary cases produced over the last 21 days by different geographic
units, taking into account the infectious period and the delay to isolation of
each case. It uses the cleaned Master Linelist database.

### Data used

This report uses the latest *cleaned master linelist*.

### Method

The data preparation involves the following steps, detailed in the following
tabs:

* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc. The function `risk_calculator` is imported as an R
object (`.rds`) from a separate report. It relies on the distribution of the
*infectious period*, i.e. the delay between the *onset of symptom* of a source case
and the *time at which they infect new secondary cases*. This is currently based
on transmission chains from Butembo, but may need periodic updating.

* **Add new variables**: addition of new variables to the VHF database, like top
  affected health zones.

* **Filter the data**: keep only relevant confirmed and probable cases, possibly
  removing erroneous dates, for further analysis



<!-- ======================================================= -->
## Load scripts

These scripts will load:

* required packages
* a few helper functions
* a data-cleaning dictionary (`cleaning_rules`)
* the path to current version of raw data (`current_data`)

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names = TRUE)
for (file in scripts_files) source(file, local = TRUE)

ggthemr("grape")

```




<!-- ======================================================= -->
## Load / install *epirisk*

This report requires a specific new RECON package called *epirisk*. It is not
loaded as part of standard packages used in other reports, as only useful
here. If this package is not installed on your computer, you need to install it
using (this requires an internet connection):

```{r install_epirisk, eval = FALSE}

devtools::install_github("reconhub/epirisk")

```

Then we load the package using:

```{r epirisk}

library(epirisk)

```




<!-- ======================================================= -->
## Import data

We load clean **linelist** data:

```{r load_data}

## linelist dataset, clean
current_clean_data
linelist <- rio::import(current_clean_data) %>% 
  as_tibble()

```

We also check that the **epi risk calculator** has been loaded:

```{r load_risk_calculator}

risk_calculator

```



<!-- ======================================================= -->
## Completion date of the database

```{r database_date}

file_name <- gsub("^[^.]+/", "", current_clean_data)
database_date <- file_name %>%
  guess_dates()
database_date

```

The **completion date** of the linelist database is 
**`r format(database_date, format = "%A %d %B %Y")`**.
All **analyses are done with respect to the date of completion** 
(not the date at which the report was last compiled). 




<!-- ======================================================= -->
## Subset data

We retain all the following subset of the data:

* confirmed and probable cases
* reported in the last 21 days

```{r subset}

last_21 <- database_date - 21
last_21

cases <- linelist %>%
  filter(epicasedef %in% c("probable", "confirmed"),
         date_report >= last_21)

```




<!-- ======================================================= -->
## Custom colors

We define custom color scales to be used in subsequent plots.

```{r custom_colors}

color_low <- "#d6d6c2"
color_high <- "#999966"

score_fill <- scale_fill_manual("Transmissibilité",
                                 values = c(score_low = color_low,
                                            score_high = color_high),
                                 labels = c(score_low = "R = 1.1",
                                            score_high = "R = 1.5"))

score_colors <- scale_color_manual("Transmissibilité",
                                   values = c(score_low = color_low,
                                              score_high = color_high),
                                   labels = c(score_low = "R = 1.1",
                                              score_high = "R = 1.5"))

```






<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# The infection intensity {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Method 

The *infection intensity* is defined as the expected number of secondary cases
infected over the last 21 days. This estimation relies on:

* dates of *onset* and *isolation* of the recent cases (report/isolation within
  last 21 days)
  
* the distribution of the delay from *onset* of a case and the *dates of
  infection* of thei secondary cases

* the average *reproduction number* (*R*) derived from recent transmission
  chains; we also use, for comparison, an *R* of 1.5 approximately corresponding
  to the infectivity in absence of vaccination


```{r compute_scores}

## compute scores with low / high R
cases$score_low <- risk_calculator(
    onset = cases$date_onset,
    isolation = cases$date_report)

cases$score_high <- risk_calculator(
    onset = cases$date_onset,
    isolation = cases$date_report,
    R = 1.5)

```


<!-- ======================================================= -->
## Infection intensity by health zone

Here we compute and display the infection intensities by health zones by,
through the following steps:

1. compute scores by health zones and re-organise results by increasing values
   of *R*
   
2. generate a plot

3. generate a table displayed in the document, and later exported


```{r scores_by_health_zone}

## step 1
scores_health_zone <- cases %>%
  group_by(zone_de_sante) %>%
  summarise(cases = n(),
            score_low = round(sum(score_low, na.rm = TRUE), 2),
            score_high = round(sum(score_high, na.rm = TRUE), 2)) %>%
  arrange(desc(score_low)) %>%
  gather(R, score, -(1:2)) %>%
  mutate(R = factor(R, levels = c("score_low", "score_high")))
  

## step 2
ggplot(scores_health_zone,
       aes(x = zone_de_sante, y = score, fill = R)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  large_txt + rotate_x_text(45) +
  score_fill +
  labs(title = "Intensité de transmission par zone de santé",
       x = "",
       y = "Estimation des cas secondaires générés \n lors des 3 dernières semaines")


## step 3
table_scores_health_zone <- scores_health_zone %>%
  spread(R, score) %>% 
  adorn_totals()

table_scores_health_zone %>%
  show_table()

```





<!-- ======================================================= -->
## Infection intensity by health area

```{r scores_by_health_area, fig.width = 10}

## step 1
scores_health_area <- cases %>%
  group_by(zone_de_sante, aire_de_sante) %>%
  summarise(cases = n(),
            score_low = round(sum(score_low, na.rm = TRUE), 2),
            score_high = round(sum(score_high, na.rm = TRUE), 2)) %>%
  arrange(desc(score_low)) %>%
  gather(R, score, -(1:3)) %>%
  mutate(R = factor(R, levels = c("score_low", "score_high")))
   

## step 2
ggplot(scores_health_area,
       aes(x = aire_de_sante, y = score, fill = R)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  large_txt +
  rotate_x_text(45) +
  score_fill +
  facet_grid(. ~ zone_de_sante,
             scale = "free", space = "free") +
  custom_horiz_facet + 
  labs(title = "Intensité de transmission par aire de santé active",
       x = "",
       y = "Estimation des cas secondaires générés \n lors des 3 dernières semaines")


## step 3
table_scores_health_area <- scores_health_area %>%
  spread(R, score) %>% 
  adorn_totals()

table_scores_health_area %>%
  show_table()

```








<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Export outputs {.tabset .tabset-fade .tabset-pills}

We export some of the relevant tables, which will be placed in the current
working directory:

```{r exports}

if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}


to_export <- c("table_scores_health_zone",
               "table_scores_health_area")

for (e in to_export) {
  rio::export(get(e),
              file.path("produced_xlsx",
                        paste0(e, Sys.Date(), ".xlsx")))
}

```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# System information {.tabset .tabset-fade .tabset-pills}

The following information documents the system on which the document was
compiled.

<!-- ======================================================= -->
## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

<!-- ======================================================= -->
## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


<!-- ======================================================= -->
## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```




<!-- ===================================================== -->
## Compilation parameters

This shows which parameters were passed through `params` at compilation time:

```{r params}
params
```



<!-- ===================================================== -->
## Change log

### Version 1.0.0
