---
title: "Weekly presentation background"
author: "Thibaut Jombart, Christopher Jarvis and Patrick Keating for the epi analysis cell - EOC Goma"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: pygments
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 2
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "figures/")

```


# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This report produces the summary graphics and figures of the background slides of the weekly presentation

### Data used

This report uses the latest cleaned Master linelist data.

### Method

The data preparation involves the following steps, detailed in the following
tabs:

* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Add new variables**: addition of new variables to the Master linelist database, like top
  affected health zones.

* **Filter the data**: keep only relevant confirmed and probable cases, possibly
  removing erroneous dates, for further analysis



## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned MDC data stored as `x`

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$", full.names=TRUE)
for (file in scripts_files) source(file, local = TRUE)

ggthemr("grape")

```



## Load data

We extract the completion date from the file name:

```{r load_data}

current_clean_data
x <- rio::import(current_clean_data)

file_name <- gsub("^[^.]+/", "", current_clean_data)
database_date <- file_name %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.



## Add new variables

* We add a new variable `top_zones` which retains the top 7 health zones (with
most cases since the beginning of the outbreak), and pool everything else into
'others': 


```{r add_new_variables}

x <- x %>%
  mutate(top_zones = top_values(zone_de_sante, 6))

```


## Filter data 

We keep only data with:

- `epicasedef`: `confirmed` and `probable`

```{r filter}

to_keep <- c("confirmed", "probable")

x <- x %>%
  filter(epicasedef %in% to_keep) %>% 
  mutate(month= floor_date(date_report, "month")) 

```


# Plots for presentation {.tabset .tabset-fade .tabset-pills}

## Outline

This section produces the following graphs for the background to the presentation:  
* epicurve by date notification and epicasedef  
* epicurve by date notification and zone de sante (top zones)  
* plot of number of aire de sante actives  
* plot of proportion of cases unknown as contacts  
* plot of proportion of décès communautaire  
* plot of proportion of cases admitted to CTE that went within 3 days
* Newly active zone de santé and newly active aire de sante.


## Epicurves

### Figure: epicurve by report and epicasedef

```{r figure_report_epicasedef, fig.width = 10}

i_report_epicasedef <- incidence(x$date_report, 7, groups = x$epicasedef)
i_report_epicasedef

plot(i_report_epicasedef, border = "white") +
  scale_fill_discrete() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Statut du cas",
                      breaks = c("confirmed", "probable"),
                      labels = c("Confirmé","Probable")) +
  large_txt +
  theme(legend.title = element_blank(),
        legend.position = c(0, 1),
        legend.justification = c(-0.05, 1.1),
        legend.background = element_rect(
            fill = "transparent",
            color = "white"
        )) +
  labs(title = paste0("Courbe épidemique: cas confirmés et probables",
                     "\npar date de notification"),
       y = "Nombre de nouveaux cas par semaine") +
  scale_x_date(breaks = "1 month", date_labels = "%d %b %Y")

```

### Figure: epicurve by community death

```{r figure_report_commdeaths, fig.width = 10}

i_report_comdeath <- incidence(x$date_report, 7, groups = x$type_death)
i_report_comdeath

plot(i_report_comdeath, border = "white") +
  scale_fill_discrete() +
  rotate_x_text(45) +
  scale_fill_discrete(name = "Statut du cas",
                      breaks = c("community_death", "non_community_death"),
                      labels = c("Décèdes communitaire","Vivant ou non décèdes communitaire")) +
  large_txt +
  theme(legend.title = element_blank(),
        legend.position = c(0, 1),
        legend.justification = c(-0.05, 1.1),
        legend.background = element_rect(
            fill = "transparent",
            color = "white"
        )) +
  labs(title = paste0("Nombre de cas confirmés et probables",
                     "\npar semaine de notification et des décèdes communitaires"),
       x = "Semaine de notification",
       y = "Nombre de nouveaux cas par semaine") +
  scale_x_date(breaks = "1 month", date_labels = "%d %b %Y")

```


### Figure: epicurve by selected Zone de sante
```{r figure_report_commdeaths_zs, fig.width = 10}

zs_select <- c("mambasa", "mandima", "mabalako", "beni", "kalunguta", "katwa", "butembo")

x_comm_death_zs <- x %>%
    filter(zone_de_sante %in% zs_select) %>%
    select(epiweek_report_label, type_death, zone_de_sante) %>%
    group_by(epiweek_report_label, zone_de_sante, type_death) %>%
    tally()
    
x_comm_death_zs$zone_de_sante <- fct_relevel(x_comm_death_zs$zone_de_sante, c("mambasa", "mandima", "mabalako", "beni", "kalunguta", "katwa", "butembo"))

x_comm_death_zs$type_death <- fct_relevel(x_comm_death_zs$type_death, c("Vivant", "Décèdes communitaire"))


ggplot(x_comm_death_zs,
       aes(x = epiweek_report_label, y = n, fill = type_death)) +
  geom_bar(stat = "identity") +
  facet_grid(zone_de_sante ~ .) +
  rotate_x_text(45) +
  large_txt +
  # scale_fill_discrete(guide = FALSE) +
  theme(legend.position = "bottom") +
  labs(title = paste0("Nombre de cas confirmés et probables",
                      "\npar semaine de notification et des décèdes communitaires"),
       x = "Semaine de notification",
       y = "Nombre de nouveaux cas par semaine") +
   theme(strip.text.y = element_text(size = 12, angle = 0)) +
  scale_x_date(breaks = "1 month", date_labels = "%d %b %Y") +
  scale_fill_discrete(name = "", labels = c("Vivant", "Décèdes communitaire", "Non connu")) +
  guides(fill = guide_legend())

```


### Table: weekly case counts by report and epicasedef

```{r table_report_epicasedef}

table_report_epicasedef <- i_report_epicasedef %>%
  as.data.frame() %>%
  adorn_totals(c("row", "col"))

table_report_epicasedef %>%
  show_table()

```



### Figure: weekly case counts by major zone (variant)

```{r figure_report_top_zones_variant, fig.width = 10, fig.height = 8}


i_report_top_zones <- incidence(x$date_report, 7, groups = x$top_zones)


df_report_top_zones <- as.data.frame(i_report_top_zones, long = TRUE)

ggplot(df_report_top_zones,
       aes(x = dates, y = counts, fill = groups)) +
  geom_bar(stat = "identity") +
  facet_grid(groups ~ .) +
  rotate_x_text(45) +
  large_txt +
  scale_fill_discrete(guide = FALSE) +
  theme(legend.position = "bottom") +
  labs(title = paste0("Courbe épidemique: cas confirmés et probables",
                      "\npar date de début de notification"),
       x = "",
       y = "Nombre de nouveaux cas par semaine") +
   theme(strip.text.y = element_text(size = 12, angle = 0)) +
  scale_x_date(breaks = "1 month", date_labels = "%d %b %Y")

```

### Table: weekly case counts by major health zone

```{r table_report_top_zones}

table_report_top_zones <- i_report_top_zones %>%
  as.data.frame() %>%
  adorn_totals(c("row", "col"))

table_report_top_zones %>%
  show_table()

```

## New health areas 21 days from the previous week. 

```{r new_zs_as}

# Match health areas with subcoordination
aire_de_sante_info <-
  x %>% 
  group_by(aire_de_sante, sous_coordination) %>% 
  count() %>% 
  group_by(aire_de_sante) %>% 
  top_n(1, n) %>%
  select(-n)

recent_week <- 
  ifelse(
    weekdays(database_date) %in% "Sunday", 
    aweek::date2week(database_date, floor_day = TRUE), 
    aweek::date2week(database_date - 7, floor_day = TRUE)
  )

start_week <-
  (aweek::week2date(recent_week) - (7 * 5)) %>% 
  aweek::date2week(floor_day = TRUE) %>% 
  as.character()

# count active health areas in recent week (previous 21 days including the last week)
current_active <- 
  x %>%
  group_by(aire_de_sante) %>%
  summarise(n_cases_21_days = sum(date_report > (aweek::week2date(recent_week) - 15))) %>% 
  filter(n_cases_21_days > 0) %>% 
  left_join(aire_de_sante_info)

# Date of penultimate week
penultimate_epiweek_1_date <- 
  (recent_week %>% aweek::week2date() - 1)

penultimate_week_active <- x %>%
  group_by(aire_de_sante) %>%
  filter(date_report > (aweek::week2date(recent_week) - 21) &  date_report < (aweek::week2date(recent_week))) %>% 
  summarise(n_cases_21_days = sum(date_report > (aweek::week2date(penultimate_epiweek_1_date) - 21))) %>%
  filter(n_cases_21_days > 0) %>% 
  left_join(aire_de_sante_info)

## New health areas
new_as <- current_active[!current_active$aire_de_sante %in% penultimate_week_active$aire_de_sante ,]

# Match zone de sante with subcoordination
zone_de_sante_info <-
  x %>% 
  group_by(zone_de_sante, sous_coordination) %>% 
  count() %>% 
  group_by(zone_de_sante) %>% 
  top_n(1, n) %>%
  select(-n)

# count active health zones in recent week (previous 21 days including the last week)
current_active_zs <- 
  x %>%
  group_by(zone_de_sante) %>%
  summarise(n_cases_21_days = sum(date_report > (aweek::week2date(recent_week) - 15))) %>%
  filter(n_cases_21_days > 0) %>% 
  left_join(zone_de_sante_info)

penultimate_week_active_zs <- 
  x %>%
  group_by(zone_de_sante) %>%
  filter(date_report < aweek::week2date(recent_week)) %>% 
  summarise(n_cases_21_days = sum(date_report > (aweek::week2date(penultimate_epiweek_1_date) - 21))) %>%
  filter(n_cases_21_days > 0) %>% 
  left_join(zone_de_sante_info)


## New health areas
new_zn <- current_active_zs[!current_active_zs$zone_de_sante %in% penultimate_week_active_zs$zone_de_sante ,]
new_zn$zone_de_sante


# to calculate number of active health areas by sous coordination 
current_active_za_for_number <- 
  x %>%
  filter(date_report > (aweek::week2date(recent_week) - 15))
length(unique(current_active_za_for_number$aire_de_sante[current_active_za_for_number$sous_coordination == "beni"]))
length(unique(current_active_za_for_number$aire_de_sante[current_active_za_for_number$sous_coordination == "butembo"]))
length(unique(current_active_za_for_number$aire_de_sante[current_active_za_for_number$sous_coordination == "goma"]))
length(unique(current_active_za_for_number$aire_de_sante[current_active_za_for_number$sous_coordination == "komanda"]))
length(unique(current_active_za_for_number$aire_de_sante[current_active_za_for_number$sous_coordination == "mangina"]))

# 
# x %>%
#   select(zone_de_sante, aire_de_sante) %>%
#   filter(aire_de_sante %in% new_as_tbl) %>%
#   group_by(zone_de_sante, aire_de_sante)
#   
 
```

Total number of health areas affected across the entire course of the outbreak: `r length(unique(x$aire_de_sante))`

Number of health areas reporting cases within the past 21 days: `r length(current_active$aire_de_sante)` 

Health areas reporting cases within the past 21 days: `r current_active$aire_de_sante`

New or reactivating health areas reporting cases within the week: `r new_as$aire_de_sante`

Active health areas in the previous week: `r length(penultimate_week_active$aire_de_sante)`

Total number of health zones affected across the entire course of the outbreak: `r length(unique(x$zone_de_sante))`

Number of health zones reporting cases within the past 21 days: `r length(current_active_zs$zone_de_sante)`

Health zones active reporting cases within the past 21 days: `r current_active_zs$zone_de_sante`

New or reactivating health zones reporting cases within the week:`r new_zn$zone_de_sante`

Health zones affected in previous week: `r length(penultimate_week_active_zs$zone_de_sante)`

## Active health areas


### Matching between health area and sub-coordination

The correspondance between health areas and sub-coordination will be needed recover the sub-coordinations information after stratifying by health areas. We define the matching based on the most frequent correspondence between health areas and sub-coordinations.


```{r health area sub coord}

aire_de_sante_info <- x %>% 
  group_by(aire_de_sante, sous_coordination) %>% 
  count() %>% 
  group_by(aire_de_sante) %>% 
  top_n(1, n) %>%
  select(-n)

```

### Computer incidence by health area

Daily incidence is computed for health areas using `incidence`, and will later be converted to a `data.frame` for further analyses:

```{r incidence by health area}

inc <- with(x, incidence(date_report, groups = aire_de_sante))

```


### Figure:: active health areas per day (past 21 days)

Here we compute the number of active health areas per day, for the entire time period of the data. Active areas are defined as health areas with at least 1 case within the last 21 days of the date considered. We do this by using the `rollsum` function, which computes the cumulative number of cases/aire de sante for a given time window (here, 21 days). Note that the argument `align = "right"` specifies that the time window, on a given day, should be in the past, not in the future. The argument `fill = NA` is also (weirdly) needed for this to work.


```{r figure_active_health_areas}

ggthemr("pale")

inc_df <- as.data.frame(inc, long = TRUE) %>% 
  rename(aire_de_sante = groups) %>% 
  left_join(aire_de_sante_info) 

timewindow <- 21

nb_as_active <- inc_df %>%
  group_by(aire_de_sante) %>%
  mutate(
      cases_last_21_days =zoo::rollsum(counts,
                                   timewindow,
                                   align = "right",
                                   fill = NA),
    has_cases = (cases_last_21_days > 0)
     ) %>%
  group_by(dates, sous_coordination) %>% 
  summarise(n_active_areas = sum(has_cases))
 

ggplot(nb_as_active) +
  geom_col(aes(x = dates, y = n_active_areas,
               fill = sous_coordination),
           width = 1) +
  labs(title = "Nombre d'aires de santé actives au cours du temps",
       x = "",
       y = paste("Nombre d'aires de santé avec au moins un cas",
                 "notifié dans les 21 derniers jours",
                 sep = "\n")) +
  scale_x_date(date_breaks = "1 month") +
  theme(legend.position = c(0.25, 0.85)) +
  scale_fill_discrete("Sous-coordination") +
  guides(fill = guide_legend(ncol = 2)) +
  large_txt +
  rotate_x_text(45)

```

## Analyses of proportion of cases unknown as contact, by week
### Figure: proportion of cases unknown as contact, by week

```{r figure_unknown_contacts_week}
ggthemr("grape")

tab_week <- 
  x %>% 
  group_by(epiweek_report,
           epiweek_report_label,
           contact_registered) %>% 
  tally() %>% 
  spread(contact_registered, n, fill = 0) %>%
  mutate(total = no + yes,
         p   = prop_to_perc(no / total),
         lci = prop_ci(no, total, "lower", TRUE),
         uci = prop_ci(no, total, "upper", TRUE))    

## Overall
ggplot(tab_week, aes(x = epiweek_report_label)) +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.2) +
  geom_point(aes(y = p)) +
  geom_line(aes(y = p)) +
  scale_y_continuous(limits = c(0, 100)) +
  large_txt +
  rotate_x_text(90)+
  xlab("Semaine de notification") +
  ylab("Proportion de cas") +
  scale_x_date(date_breaks = "2 weeks", date_labels =  "%d %b") +
  theme(axis.text.x = element_text(vjust = .5, angle = 90)) +
  ggtitle("Proportion des cas non connus comme contact",
          subtitle = 'avec intervalle de confiance à 95%')

```

### Table: proportion of cases unknown as contact, by week
```{r table_proportion_cases_unknown_contact}

tab_week %>% 
  show_table()

```


## Deces communautaire proportion

### Figure: Proportion of community deaths over time

```{r fig_report_community_death_proportion}


proportion_community_death <- x %>% 
  group_by(epiweek_report_label, epiweek_report, type_death) %>% 
  count() %>% 
  spread(type_death, n, fill = 0) %>%
  mutate(total = alive + community + etc_tc) %>%
  mutate(prop_community_death = prop_to_perc(community / total),
         lci_prop_community_death = prop_ci(community, total, "lower", TRUE),
         uci_prop_community_death = prop_ci(community, total, "upper", TRUE)) %>%
  select(epiweek_report,
         prop_community_death,
         lci_prop_community_death,
         uci_prop_community_death,
         everything())


ggplot(proportion_community_death, 
       aes(x = epiweek_report_label, y = prop_community_death)) +
  geom_line(aes(x = epiweek_report_label, y = prop_community_death)) +
  geom_ribbon(aes(x = epiweek_report_label, ymin = lci_prop_community_death, 
                  ymax = uci_prop_community_death),
              alpha = 0.25) +
  ylim(0,100) +
  rotate_x_text(45) +
  scale_months +
  large_txt +
  labs(x = "",
       y = "Pourcentage de décès communautaires",
       title = paste("Décès communautaires parmi les cas confirmés et",
                     "probables par semaine de notification",
                     sep = "\n"))


```

### Table: Proportion of community deaths over time

```{r table_report_community_death_proportion}

table_proportion_community_death <- proportion_community_death %>%
  as_tibble() %>%
  select("Semaine de notification" = epiweek_report_label,
         "Semaine épi" = epiweek_report,
         "Décès communautaires" = community,
         "Décès CTE ou CT" = etc_tc,
         "Inconnu" = unknown,
         "Total connu" = total,
         "% décès communautaire" = prop_community_death,
         "IC 95% inf" = lci_prop_community_death,
         "IC 95% sup" = uci_prop_community_death)


table_proportion_community_death %>% 
  show_table()

```



## CTE admission within 3 days of onset
### Figure: Proportion of cases admitted to CTE within 3 days of onset of symptoms

```{r figure_proportion_cte_three_day_delay}

#Calculate the proportion of cases with a delay of <=3 days from onset to admission to CTE


proportion_delay_cte <- 
  x %>% 
  mutate(cte_delay = as.integer(date_admission - date_onset)) %>% 
  mutate(month = floor_date(date_report, "month")) %>% 
  filter(!is.na(date_onset),
         !is.na(date_admission),
         cte_delay >= 0) %>% 
  mutate(cte_delay_group = case_when(
    cte_delay >= 0 & cte_delay < 4 ~ "0-3",
    cte_delay >=4 & cte_delay < 7 ~ "4-6",
    cte_delay >= 7 & cte_delay < 10 ~ "7-9",
    is.na(delay_admission) ~ "inconnu",
    TRUE ~ ">=10"
  )) %>% 
  select(month, cte_delay_group) %>% 
  group_by_all() %>% 
  tally() %>% 
  spread(cte_delay_group, n, fill = 0) %>% 
  adorn_totals(c('col')) %>% 
  mutate(prop_under3 = prop_to_perc(`0-3`/Total),
         lci_prop_under3 = prop_ci(`0-3`, Total, "lower", TRUE),
         uci_prop_under3 = prop_ci(`0-3`, Total, "upper", TRUE)) 
  
  
  proportion_delay_cte_plot <- 
  ggplot(proportion_delay_cte)+
  geom_line(aes(x = month, y = prop_under3)) +
  geom_ribbon(aes(x = month, ymin = lci_prop_under3, 
                  ymax = uci_prop_under3),
              alpha = 0.25) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %d %Y") +
  rotate_x_text(45) +
  labs(x = "Mois de notification", y = "Proportion", title = "Proportion des cas admis au CTE dans les 3 jours après la date \nde début de symptômes par mois de notification") +
  ylim(0,100)


proportion_delay_cte_plot


```

### Table: proportion of cases admitted to CTE within 3 days of onset of symptoms

```{r table_proportion_cte_three_day_delay}

proportion_delay_cte %>% 
  show_table()

```

# Cases as contacts  {.tabset .tabset-fade .tabset-pills}

## Outline
•	Proportion of new cases listed and followed as contacts in the past 21 days, compared to previous 21 days: xx%
•	Proportion of new cases listed as contacts but not able to be followed as contacts in the past 21 days compared to previous 21 days: xx%


## Proportion of new cases listed and followed as contacts in the past 21 days
### Table: proportion of new cases listed and followed as contacts in the past 21 days

```{r table_proportion_new_cases_listed_followed_recent21days}
# NOTE THAT THIS IS A DUPLICATION OF WHAT'S IN THE WEEKLY CASES_UNKNOWN_CONTACT.RMD FILE 

past21days <- database_date - 20


number_cases_contacts_followed_past21days <- 
  x %>% 
  filter(contact_registered == "yes" & contact_surveilled == "yes") %>% 
  filter(date_report >= past21days ) %>% 
  select(contact_surveilled) %>% 
  tally()


number_cases_past21days <- 
   x %>% 
  filter(date_report >= past21days) %>% 
  tally()


prop_followed_past21days <- prop_to_perc(number_cases_contacts_followed_past21days/number_cases_past21days)

prop_followed_past21days %>% 
  show_table()

```


### Table: proportion of new cases listed and followed as contacts in the previous 21 days (i.e starting from one week earlier that current 21 days)

```{r table_proportion_new_cases_listed_followed previous 21 days}


enddate <- database_date - 7
startdate <- database_date -27


number_cases_contacts_followed_previous21days <- 
  x %>% 
  filter(contact_registered == "yes" & contact_surveilled == "yes") %>% 
  filter(date_report >= startdate & date_report <= enddate) %>% 
  select(contact_surveilled) %>% 
  tally()


number_cases_previous21days <- 
   x %>% 
  filter(date_report >= startdate & date_report <= enddate) %>% 
  tally()


prop_followed_previous21days <- prop_to_perc(number_cases_contacts_followed_previous21days/number_cases_previous21days)


prop_followed_previous21days %>% 
  show_table()

```




## Proportion of new cases listed and not followed up as contacts in the past 21 days
### Table: proportion of new cases listed and not followed as contacts in the past 21 days

```{r table_proportion_new_cases_not_followed_recent21days}

past21days <- database_date - 20


number_cases_contacts_not_followed_past21days <- 
  x %>% 
  filter(contact_registered == "yes" & contact_surveilled == "no") %>% 
  filter(date_report >= past21days ) %>% 
  select(contact_surveilled) %>% 
  tally()


number_cases_past21days <- 
   x %>% 
  filter(date_report >= past21days) %>% 
  tally()


prop_not_followed_past21days <- prop_to_perc(number_cases_contacts_not_followed_past21days/number_cases_past21days)

prop_not_followed_past21days %>% 
  show_table()

```


### Table: proportion of new cases listed and not followed as contacts in the previous 21 days (i.e starting from one week earlier that current 21 days)

```{r table_proportion_new_cases_listed_not_followed previous 21 days}


enddate <- database_date - 7
startdate <- database_date -27


number_cases_contacts_not_followed_previous21days <- 
  x %>% 
  filter(contact_registered == "yes" & contact_surveilled == "no") %>% 
  filter(date_report >= startdate & date_report <= enddate) %>% 
  select(contact_surveilled) %>% 
  tally()


number_cases_previous21days <- 
   x %>% 
  filter(date_report >= startdate & date_report <= enddate) %>% 
  tally()


prop_not_followed_previous21days <- prop_to_perc(number_cases_contacts_not_followed_previous21days/number_cases_previous21days)


prop_not_followed_previous21days %>% 
  show_table()

```

# Export tables {.tabset .tabset-fade .tabset-pills}

## Outline

We export the various tables produced in this report to excel files saved in a
folder `produced_xlsx`.

## Excel files

```{r exports, eval = FALSE, include = FALSE}

if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

```








# System information {.tabset .tabset-fade .tabset-pills}

The following information documents the system on which the document was
compiled.

## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```
