---
title: "Analyses of transmission chains"
author: "Thibaut, Maddy, Sam, Luc, Sonia, Elias, Xavier, the Butembo team, the analytic cell Goma"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 4
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE)
```



<br>

<div class="report_meta">
  <span class="notice">**Notice**: this is a **stable, routine report**. 
  **Do not touch it unless it is broken.** To make a contribution, **carefully read 
  the [README](../../../../../README.html) file**.</span>
  
  **Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)
  
  **Code contributors:** Thibaut Jombart, Amy Gimma
  
  **Data contributors:** transmission chains teams from Butembo, Mangina, Beni, HQ
  
  **Version:** 1.1.0
  
  **Reviewed by:**
</div>




<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline

This report focusses on transmission chains of the outbreak, derived from the
narratives. This work was pioneered by the Butembo team, who assembled specific
linelists with transmission information. To simplify this workflow for other
teams (with less data management resources), we streamline this process as
follows:

1. information on the cases are taken from the clean linelist data (the *master
   linelist*), thus not requiring additional work from the field teams
   
2. information on transmissions is stored in a separate file following the same
   model as the master linelist, i.e. a google spreadsheet which can be edited
   by different data managers; this file only contains information on the
   *transmissions*, not on the cases, e.g.: from whom, to whom, which kind of
   transmission context etc.
   
The two sources of data will be merged using *epicontacts*.



### Data used

We combine two data sources:

* the *master linelist*, containing information about the cases

* the *master transmission list*, describing transmission chains using unique
  identifiers for source cases and secondary cases
  

### Method

The data preparation involves the following steps, detailed in the following tabs:


* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Clean data**: this section contains *ad hoc* data cleaning, i.e. which is
  not used in other reports (otherwise cleaning should be done in a dedicated
  report); this section is also used to create new variables used in the
  analyses




## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned VHF data stored as `x`

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$",
                     full.names = TRUE)
for (file in scripts_files) source(file)

```



## Load data

Two types of data are used in this report:

1. the clean *master linelist*, called `linelist` in this report
2. the raw *master transmission list*, called `chains` in this report

```{r load_data}

## load the master linelist ('MLL')
current_linelist
linelist <- rio::import(current_linelist) %>%
    as_tibble()
linelist

## load the chains
current_chains
chains_raw <- custom_import(current_chains) %>%
    as_tibble()
chains_raw

## extract completion dates
linelist_file_name <- gsub("^[^.]+/", "", current_linelist)
linelist_date <- linelist_file_name %>%
  guess_dates()
linelist_date

chains_file_name <- gsub("^[^.]+/", "", current_chains)
chains_date <- chains_file_name %>%
  guess_dates()
chains_date

```

The **completion dates** of the databases are:

* master linelist: **`r format(linelist_date, format = "%A %d %b %Y")`**
* master transmission: **`r format(chains_date, format = "%A %d %b %Y")`**





## Clean data

As we use a clean master linelist, cleaning only applies here to the *master
transmission list* stored in the object called `chains_raw`. Once cleaned,
transmission data will be stored as `chains`.

We use *linelist*'s function `clean_data()` to:

- remove all special characters from the data
- set all characters to lower case
- replace all accentuated and diacritic characters with their closest ascii
  match in the latin alphabet
- replace all separators with a single `_`
- replace all mis-spelling using a overall dictionary (see the file
  `cleaning_rules.xlsx` in `/dictionary/`, or type `cleaning_rules` in this R
  session, after compiling the document
- (optionally) detect date formats and convert data to dates, including cases
  where format varies within a column
  

This cleaning is achieved with:

```{r data_cleaning}

chains <- chains_raw %>%
  clean_data(guess_dates = FALSE) %>%
  as_tibble()

chains

```




## Cleaning using dictionary

Here we correct typos etc using our dictionary. We then tabulate the categories
of key variables to make sure nothing has been left out. New items to clean
should be added to the dictionary in `dictionary/cleaning_rules.xlsx`.

```{r clean_dic}

## clean using dictionary
chains <- chains %>%
  clean_variable_spelling(wordlists = cleaning_rules)

## check variables
table(chains$type_contamination, useNA = "ifany")
table(chains$lien_cas_source, useNA = "ifany")

```




## Removing duplicate chains

As data entry in the master transmission list is performed by multiple people,
duplicated chains do occur (same combination of *source -> secondary case*). We
remove these from the `chains` data.

```{r remove_duplicate_chains}

chains <- chains %>%
  mutate(chain_txt = paste(source_case_id, case_id, sep = "->"),
         duplicate = duplicated(chain_txt)) %>%
  filter(!duplicate)

```



## Making an *epicontacts* object {.tabset .tabset-fade .tabset-pills}


We build the *epicontacts* object using:

* the *master linelist* stored in `linelist`, keeping only selected variables
  useful for further analyses, and entries for which we are missing VHF
  identifiers (`vhf_code`), as well as duplicates
  
* the *master transmission list* data stored in `chains`

```{r epicontacts}

select_linelist <- linelist %>%
  rename(mll_id = mllid,
         vhf_id = vhf_code) %>% 
  select("vhf_id",
         "mll_id",
         "epicasedef",
         "date_onset",
         "date_report",
         "date_death",
         "date_admission",
         "date_exposure_start",
         "date_exposure_end",
         "zone_de_sante",
         "aire_de_sante",
         "age",
         "age_class",
         "gender",
         "hcw",
         "firstname",
         "surname",
         "current_status",
         "report_status",
         "outcome",
         "hosp1_name",
         "sous_coordination") %>%
  filter(!is.na(vhf_id),
         !duplicated(vhf_id))

x <- make_epicontacts(linelist = select_linelist,
                      contacts = chains,
                      directed = TRUE,
                      id = "vhf_id",
                      from = "source_case_id",
                      to = "case_id")
x

```




## Identify clusters / chains

We define clusters as sets of cases connected by transmission events. In graph
terminology, these are the connected components of the graph. We also:

* identify **active clusters**, defined as clusters having at least one new case
notified within the last 21 days

* identify the main active clusters 

```{r clusters}

x <- get_clusters(x,
                  member_col = "cluster_id",
                  size_col = "cluster_size",
                  override = TRUE)

## re-label the cluster ids
x$linelist$cluster_id <- paste("chain", x$linelist$cluster_id, sep = "_")

## find which clusters are active
x$linelist <- x$linelist %>%
  group_by(cluster_id) %>%
  mutate(cluster_active = any(date_report > date_recent, na.rm = TRUE))


## add top_clusters
x$linelist <- x$linelist %>%
  mutate(top_cluster = top_values(cluster_id, 6, subset = x$linelist$cluster_active))

```









## Active chains on last 21 days

We define as active chains over the last 21 days as: 

A. cases notified within the last 21 days, i.e. 21 days prior to the completion
   date of the chain database (i.e. here, **`r format(chains_date - 42, format =
   "%A %d %b %Y")`**)
   
B. all the cases of the transmission chains linked to cases identified in A)


The code to do this is actually quite involved. The strategy is to:

1. identify connected components (i.e. clusters of linked cases)

2. identify the roots of the clusters, i.e. the first case of each cluster

3. name clusters after their roots

4. identify the most recent case of each cluster; retain only clusters whose
   most recent case has been notified within the *recent* time window
   
5. keep only singletons which are within the defined time window

The resulting chains are stored as a new object `x_active`.


```{r plot_chains_time_window}

## 1. identify transmission chains ("clusters" or "connected components" in
## graph terminology)


## 2. identify the roots, i.e. source case of the transmission chains; these are
## cases who have no edge coming in, and at least one coming out

in_degree <- data.frame(in_degree = get_degree(x, "in")) %>%
  rownames_to_column(var = "id")
out_degree <- data.frame(out_degree = get_degree(x, "out")) %>%
  rownames_to_column(var = "id")

# Adds in_degree and out_degree to x$linelist, using the id variable for the joins
# Also adds a variable in to determine whether or not an individual is a root or not
x$linelist <- x$linelist %>%
  left_join(in_degree) %>%
  left_join(out_degree) %>%
  mutate(is_root = (in_degree == 0),# & (out_degree >= 0),
         root = if_else(is_root, id, NULL))

## find roots of each transmission chain
roots <- split(x$linelist$root, x$linelist$cluster_id) %>%
  lapply(function(e) e[!is.na(e)]) %>%
  stack() %>%
  setNames(c("root", "chain_id")) %>%
  as_tibble() %>% 
  select("chain_id", "root")

roots


## 3. name clusters after their roots; this information to the linelist as
## `chain_id` for all instances where in cluster_id that match one of the names
## in chain_id column of roots, replaces the value in cluster_id with the value
## in root column of roots, and uses this as a new variable, chain_id

x$linelist <- x$linelist %>%
  mutate(chain_id = clean_spelling(cluster_id, wordlist = roots))

## set limit date
date_recent <- chains_date - 42


## 4. find transmission chains whose latest case is after `date_recent`

chains_to_keep <- x$linelist %>%
  group_by(chain_id) %>%
  summarise(most_recent = max(date_admission, na.rm = TRUE)) %>%
  filter(most_recent > date_limit) %>%
  pull(chain_id)

x_active <- subset(x,
                   node_attribute = list(chain_id = chains_to_keep)) %>%
  thin()



```








## Summary

This provides an overview of the content of the *epicontacts* object. Pay
attention, in particular, to cases that are in the transmission tree (called
"contacts" in *epicontacts* terminology) but not in the linelist. These could be
recovered by using a larger database for linelist, such as VHF or the new MLL
(Master Line List) currated in Goma.

```{r summary}
summary(x)
```



## General plot

This is a general visusalisation of the whole dataset. Because of its large
size, we do not show the full plot in this document, but ouput it in a separate
file [`plot_chains.html`](plot_chains.html). 

```{r plot_chains}

plot_x <- plot(x,
               node_color = "zone_de_sante",
               node_shape = "gender",
               shapes = c(female = "female",
                          male = "male",
                          unknown = "question-circle"),
               legend_max = 100,
               edge_color = "type_contamination"
               )

```



## Plot active chains

Graphs will be shown using the date of isolation (`date_admission`) if no value
is missing; otherwise the default plotting of transmissions will be used.


```{r plot_active_chains}
## check if some date_admission are missing
missing_date <- get_pairwise(x_active, "date_admission") %>% is.na()

if (!any(missing_date)) {
  x_axis_info <- "date_admission"
} else {
  x_axis_info <- NULL
}

## make the plot
plot_x_active <- plot(x_active,
                      node_color = "chain_id",
                      node_shape = "gender",
                      shapes = c(female = "female",
                                 male = "male",
                                 unknown = "question-circle"),
                      legend_max = 100,
                      edge_color = "type_contamination",
                      x_axis = x_axis_info,
                      edge_width = 5
                      )
plot_x_active

```



## Active chains on last 6 weeks days

We define as active chains, on a given limit date: 

1. cases that were isolated after the limit date, set up to be 6 weeks prior to
   the completion date of the chain database (i.e. here, **`r
   format(chains_date - 42, format = "%A %d %b %Y")`**)
   
2. all the cases of the transmission chains linked to 1)

Graphs will be shown using the date of isolation (`date_admission`) if no value
is missing; otherwise the default plotting of transmissions will be used.

```{r plot_chains_time_window}

## identify transmission chains ("clusters" or "connected components" in graph
## terminology)
x <- get_clusters(x, member_col = "cluster_id", size_col = "chain_size", override = TRUE)

## force character type
x$linelist$cluster_id <- paste("chain", x$linelist$cluster_id, sep = "_")

## find source case of the transmission chains; these are cases who have no edge
## coming in, and at least one coming out

in_degree <- data.frame(in_degree = get_degree(x, "in")) %>%
  rownames_to_column(var = "id")
out_degree <- data.frame(out_degree = get_degree(x, "out")) %>%
  rownames_to_column(var = "id")

# Adds in_degree and out_degree to x$linelist, using the id variable for the joins
# Also adds a variable in to determine whether or not an individual is a root or not
x$linelist <- x$linelist %>%
  left_join(in_degree) %>%
  left_join(out_degree) %>%
  mutate(is_root = (in_degree == 0) & (out_degree > 0),
         root = if_else(is_root, id, NULL))

## find roots of each transmission chain
roots <- split(x$linelist$root, x$linelist$cluster_id) %>%
  lapply(function(e) e[!is.na(e)]) %>%
  stack() %>%
  setNames(c("root", "chain_id")) %>%
  as_tibble() %>% 
  select("chain_id", "root")

roots

## add this information to the linelist as `chain_id`
# for all instances where in cluster_id that match one of the names in chain_id column of roots, replaces
# the value in cluster_id with the value in root column of roots, and uses this as a new variable, chain_id 
x$linelist <- x$linelist %>%
  mutate(chain_id = clean_spelling(cluster_id, wordlist = roots))

## set limit date
date_limit <- chains_date - 42

## find transmission chains whose latest case is after the limit date
chains_to_keep <- x$linelist %>%
  group_by(chain_id) %>%
  summarise(most_recent = max(date_admission, na.rm = TRUE)) %>%
  filter(most_recent > date_limit) %>%
  pull(chain_id)

x_active <- subset(x,
                   node_attribute = list(chain_id = chains_to_keep)) %>%
  thin("contacts")

## check if some date_admission are missing
missing_date <- get_pairwise(x_active, "date_admission") %>% is.na()

if (!any(missing_date)) {
  x_axis_info <- "date_admission"
} else {
  x_axis_info <- NULL
}

## make the plot
plot_x_active <- plot(x_active,
                      node_color = "chain_id",
                      node_shape = "gender",
                      shapes = c(female = "female",
                                 male = "male",
                                 unknown = "question-circle"),
                      legend_max = 100,
                      edge_color = "type_contamination",
                      x_axis = x_axis_info,
                      edge_width = 5
                      )
plot_x_active


```








<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Descriptive analyses {.tabset .tabset-fade .tabset-pills}

## Outline

In the following we provide some simple descriptions of the transmission
network. We focus on spatial patterns in transmissions, comparing the locations
of the sources cases and their secondary cases, and the relative frequencies of
these transmissions.




## By sex

```{r contacts_sex}

pairs_sex <- get_pairwise(x,
                          "gender",
                          function(from, to)
                            as.data.frame(table(from, to)))

ggplot(pairs_sex,
       aes(y = from, x = to, size = Freq)) +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 20),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(45) + 
  labs(title = "Transmissions between sex",
       y = "Source case",
       x = "Infectee")

```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_sex_circi}

## Chord diagram
circle_plot(pairs_sex, label_space = 2)

```




## By age class

```{r contacts_age}

pairs_age <- get_pairwise(x,
                          "age_class",
                          function(from, to)
                            as.data.frame(table(from, to)))

ggplot(pairs_age,
       aes(y = from, x = to, size = Freq)) +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 20),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(45) + 
  labs(title = "Transmissions between Age Groups",
       y = "Source case",
       x = "Infectee")

```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_age_circi}

## Chord diagram
circle_plot(pairs_age, label_space = 2)

```
 
 

## By health zones

We plot transmissions by health zones:

```{r contacts_zone_de_sante}

pairs_zone_de_sante <- get_pairwise(x,
                                    "zone_de_sante",
                                    function(from, to)
                                      as.data.frame(table(from, to)))

ggplot(pairs_zone_de_sante,
       aes(y = from, x = to, size = Freq)) +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 20),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(45) + 
  labs(title = "Transmissions across health zones",
       y = "Source case",
       x = "Infectee")

```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_zone_de_sante_circi}

## Chord diagram
circle_plot(pairs_zone_de_sante, label_space = 2)

```

We compute the proportion of transmission happening across different health
zones:

```{r move_zones}

move_zone <- get_pairwise(x,
                          "zone_de_sante",
                          function(from, to)
                            as.logical(na.omit(from != to)))
move_zone

## mean and 95% CI
mean(move_zone)
prop_ci(k = sum(move_zone), n = length(move_zone))

```




## By health areas

We plot transmissions by health areas:

```{r contacts_aire_de_sante, fig.width = 12}

pairs_aire_de_sante <- get_pairwise(x,
                                    "aire_de_sante",
                                    function(from, to)
                                      as.data.frame(table(from, to)))

ggplot(pairs_aire_de_sante,
       aes(y = from, x = to, size = Freq)) +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 10),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(90) + 
  labs(title = "Transmissions across health areas",
       y = "Source case",
       x = "Infectee")


```


We compute the proportion of transmission happening across different health
areas:

```{r move_areas}

move_area <- get_pairwise(x,
                          "aire_de_sante",
                          function(from, to)
                            as.logical(na.omit(from != to)))
move_area

## mean and 95% CI
mean(move_area)
prop_ci(k = sum(move_area), n = length(move_area))

```




## Sources and recipients of transmissions

In this section we characterise health areas in terms of the number of
transmissions given ("*sources*"), or received ("*recipients*").

```{r sources_recipients, fig.width = 10}

## count cases sourced by area
df_sources <- pairs_aire_de_sante %>%
  filter(as.character(from) != as.character(to), Freq > 0) %>%
  group_by(from) %>%
  summarise(n_sourced = sum(Freq)) %>%
  rename(aire_de_sante = from)

## count cases received by area
df_receivers <- pairs_aire_de_sante %>%
  filter(as.character(from) != as.character(to), Freq > 0) %>%
  group_by(to) %>%
  summarise(n_received = sum(Freq)) %>%
  rename(aire_de_sante = to)

## put the information together
df_sources_recipients <- full_join(df_sources, df_receivers) %>%
  filter(aire_de_sante != "inconnu") %>% 
  replace_na(list(n_sourced = 0, n_received = 0)) %>%
  arrange(desc(n_sourced), desc(n_received))

## add info on zones
zone_info <- linelist %>%
  group_by(aire_de_sante, zone_de_sante) %>%
  count() %>%
  group_by(aire_de_sante) %>%
  top_n(1, n) %>%
  ungroup() %>% 
  select(1:2)

df_sources_recipients <- right_join(zone_info, df_sources_recipients)

## display the table
df_sources_recipients %>%
  show_table()


df_sources_recipients_top <- df_sources_recipients %>%
  mutate(sum = n_sourced + n_received) %>%
  group_by(zone_de_sante) %>% 
  top_n(5, sum) %>%
  arrange(zone_de_sante, sum) %>%
  mutate(rank = 1:n()) %>% 
  top_n(4, rank)


## make the plot
ggplot(df_sources_recipients,
       aes(x = n_received,
           y = n_sourced,
           color = aire_de_sante)) +
  theme_bw() +
  geom_point() +
  geom_label_repel(data = df_sources_recipients_top,
                   aes(label = aire_de_sante), size = 4) +
  scale_x_continuous(breaks = int_breaks) +
  scale_y_continuous(breaks = int_breaks) +
  large_txt +
  theme(strip.text = element_text(size = 16)) +
  facet_wrap(~ zone_de_sante) +
  scale_color_discrete(guide = FALSE) +
  labs(title = "Sources and recipients of cases",
       x = "Number of cases received",
       y = "Number of cases sourced")

```



<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Analysis of transmissibility {.tabset .tabset-fade .tabset-pills}

## Reproduction number

We can assess the effective reproduction number ($R$) by looking at the numbers
of secondary cases by index case:

```{r R}

## in-degree
x %>% get_degree() %>% table()

## out-degree = reproduction number
x %>% get_degree("out") %>% table()

R_df <- x %>%
  get_degree("out") %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  rename("id" = "rowname", "R" = ".")

head(R_df)

```

We create a new `data.frame` with R values and case info from the linelist;
unlike the linelist, where each line is a case, here each line is a source case:

```{r R_df}

R_df <- left_join(R_df, x$linelist, by = "id") %>%
  as_tibble()

R_df %>%
  show_table()

regression <- lm(R ~ R_df$age_class, data = R_df)
summary(regression)

regression <- lm(R ~ (R_df$date_report - R_df$date_onset), data = R_df)
summary(regression)

```

Because of delays from primary to secondary onset, and from onset to
notification, we need to ignore the latest source cases to avoid
under-estimating the $R$. We use a conservative limit of 45 days:

```{r R_plots}

old_cases <- R_df %>%
  filter(date_onset < (chains_date - 45))

ggplot(old_cases, aes(x = R)) +
  geom_bar() +
  large_txt +
  labs(title = "Number of secondary cases by index case",
       x = "Observed number of secondary cases (R)",
       y = "Frequency")

```




## Estimating the distribution of R

In this part we estimate the underlying distribution of $R$, accounting for
superspreading using a Negative Binomial distibution. The two parameters that
will be estimated are:

* the reproduction number
* the dispersion parameter, which increases when super-spreading events increase
  in frequency.

```{r estimate_R}

## likelihood function for a Neg Bin distribution of R
## (m: the mean; disp: dispersion parameter)

likelihood <- function(mean, disp) {
  ## exp(sum(dpois(old_cases$R, m, log = TRUE)))
  exp(sum(dnbinom(old_cases$R, mu = mean, size = disp, log = TRUE)))
}

R_values <- seq(0, 2, length = 100)
dispersion_values <- seq(0, 2, length = 100)
grid <- expand.grid(R = R_values,
                    dispersion = dispersion_values)
like_values <- apply(grid, 1, function(e) likelihood(e[1], e[2]))

R_ml <- grid[which.max(like_values), "R"]
R_ml
disp_ml <- grid[which.max(like_values), "dispersion"]
disp_ml

df_grid_like <- cbind.data.frame(grid, likelihood = like_values)

ggplot(df_grid_like) +
  geom_raster(aes(x = R, y = dispersion, fill = likelihood)) +
  large_txt +
  scale_fill_gradient(low = "#ffffff", high = "#800000") +
  labs(title = "Likelihood surface: \nnegative binomial distribution",
       x = "Mean reproduction number (R)",
       y = "Dispersion parameter")

```

The likelihood function is well-defined, with a maximum-likelihood estimate of
$R$ is `r round(R_ml, 2)`. How well does this model fit the data? We can compare
predictions from the Negative Binomial distribution (MLE) to the numbers of
secondary cases actually observed:

```{r expected_secondary}

n_cases <- 0:15
obs_freq <- sapply(n_cases, function(e) mean(old_cases$R == e))
pred_freq <- dnbinom(n_cases, mu = R_ml, size = disp_ml)

df_model <- data.frame(n_cases,
                       observed = obs_freq,
                       predicted = pred_freq) %>%
  gather(type, frequency, -1)

ggplot(df_model,
       aes(x = n_cases,
           y = frequency,
           fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  large_txt +
  scale_x_continuous(breaks = int_breaks) +
  scale_fill_manual("", values = c(observed = "#d27979", predicted = "#9999ff")) +
  labs(title = "Nombre de cas secondaires observés et prédits",
       x = "Nombre de cas secondaires",
       y = "Fréquence")

```



## Effective reproduction number in time

```{r R_in_time}

ggplot(R_df, aes(x = date_admission, y = R)) +
  geom_point() +
  geom_smooth() +
  large_txt +
  scale_weeks +
  rotate_x_text(45) +
  labs(title = "Nombre de reproduction effectif au cours du temps",
       x = "Date d'admission",
       y = "Nombre de cas secondaires par cas index")

```







<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Delay distributions {.tabset .tabset-fade .tabset-pills}

In this section, we estimate delays characterising transmission pairs. The
procedure is the same for all delays:

1. compute delays from the data to obtain an **empirical distribution** (histograms)

2. inspect the empirical distribution for **outliers**, and remove them if needed

3. estimate the underlying distribution by **fitting a discretised Gamma**
   distribution to the empirical delays


## Serial interval distribution

The serial interval is defined as the delay between symptom onset between
primary (source) cases and secondary cases.

We first derive the empirical distribution:

```{r serial_interval_empirical}

serial_interval_empirical <- x %>%
  get_pairwise("date_onset") %>%
  na.omit() %>%
  as.integer()

serial_interval_base <- serial_interval_empirical %>%
  tibble() %>%
  rename(serial_interval = ".") %>%
  ggplot(aes(x = serial_interval)) +
  geom_histogram(aes(y = stat(count / sum(count))),
                 color = "white", fill = "navyblue", alpha = .3) +
  geom_rug(color = "navyblue", alpha = .3) +
  theme_bw() +
  large_txt +
  labs(title = "Intervalle sériel: distribution empirique",
       x = "Jours entre l'apparition des symptômes \ndes cas primaires et secondaires",
       y = "Nombre de paires de transmission")
serial_interval_base

```

There are no indication of outliers, so we can go ahead and fit a discretized
Gamma to the data:

```{r serial_interval_fit}

serial_interval_fit <- serial_interval_empirical %>%
  fit_disc_gamma(mu_ini = mean(serial_interval_empirical))

serial_interval_fit

## check that the fitting worked
if (!serial_interval_fit$converged) {
  stop("Fitting of serial interval did not converged")
}

## make plot with fitting
serial_interval_base +
  stat_function(fun = serial_interval_fit$distribution$d,
                color = "#cc6699", lwd = 1) +
  labs(title = "Distribution de l'intervalle sériel")

```

The serial interval is characterised by:

* mean of **`r round(serial_interval_fit$mu, 1)`** days
* standard deviation of **`r round(serial_interval_fit$sd, 1)`**
* **50% of the delays** are between **`r serial_interval_fit$distribution$q(.25)`** 
and **`r serial_interval_fit$distribution$q(.75)`** days
* **95% of the delays** are between **`r serial_interval_fit$distribution$q(.025)`** 
and **`r serial_interval_fit$distribution$q(.975)`** days






## Reporting interval distribution

This distribution is a variant of the serial interval, where dates of reporting
are used instead of dates of symptom onset.

We first derive the empirical distribution:

```{r reporting_interval_empirical}

reporting_interval_empirical <- x %>%
  get_pairwise("date_report") %>%
  na.omit() %>%
  as.integer()

reporting_interval_base <- reporting_interval_empirical %>%
  tibble() %>%
  rename(reporting_interval = ".") %>%
  ggplot(aes(x = reporting_interval)) +
  geom_histogram(aes(y = stat(count / sum(count))),
                 color = "white", fill = "navyblue", alpha = .3) +
  geom_rug(color = "navyblue", alpha = .3) +
  theme_bw() +
  large_txt +
  labs(title = "Intervalle de notification: distribution empirique",
       x = "Jours entre la notification \ndes cas primaires et secondaires",
       y = "Nombre de paires de transmission")
reporting_interval_base

```

There are no indication of outliers, so we can go ahead and fit a discretized
Gamma to the data:

```{r reporting_interval_fit}

reporting_interval_fit <- reporting_interval_empirical %>%
  fit_disc_gamma(mu_ini = mean(reporting_interval_empirical))

## check that the fitting worked
if (!reporting_interval_fit$converged) {
  stop("Fitting of reporting interval did not converged")
}

## make plot with fitting
reporting_interval_base +
  stat_function(fun = reporting_interval_fit$distribution$d,
                color = "#cc6699", lwd = 1) +
  labs(title = "Distribution de l'intervalle de notification")

```


The reporting interval is characterised by:

* mean of **`r round(reporting_interval_fit$mu, 1)`** days
* standard deviation of **`r round(reporting_interval_fit$sd, 1)`**
* **50% of the delays** are between **`r reporting_interval_fit$distribution$q(.25)`** 
and **`r reporting_interval_fit$distribution$q(.75)`** days
* **95% of the delays** are between **`r reporting_interval_fit$distribution$q(.025)`** 
and **`r reporting_interval_fit$distribution$q(.975)`** days




## Comparison of distributions

In this section, we compare the different distributions estimated from the
transmission chains.

```{r comparison}

max_delay <- max(serial_interval_empirical,
                 reporting_interval_empirical) + 5
delays <- 0:max_delay

distributions <- tibble(
    delays,
    serial_interval = serial_interval_fit$distribution$d(delays),
    reporting_interval = reporting_interval_fit$distribution$d(delays)) %>%
  gather(distribution, proba, -1)


ggplot(distributions,
       aes(x = delays, y = proba, color = distribution)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  theme_bw() +
  large_txt +
  labs(title = "Comparaison des distributions",
       x = "Nombre de jours",
       y = "Probabilité") +
  theme(legend.position = c(.8, .8))


```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Export outputs {.tabset .tabset-fade .tabset-pills}

## Outline

We save several outputs to separate files, including a `html` files for the
transmission chain plots, and R objects (`.rds` format) for fitted delay
distributions.



## R objects

We export the following items as R objects, saved into `produced_rds`:

* **epicontacts.rds**: the *epicontacts* object

* **serial_interval_fit.rds**: the serial interval distribution

* **reporting_interval_fit.rds**: the reporting interval distribution


```{r rds_exports}

if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}

## epicontact object
saveRDS(x,
        file = file.path("produced_rds",
                         "epicontacts.rds"))
destination <- here("data", "clean", "epicontacts.rds")

file.copy(from = file.path("produced_rds",
                           "epicontacts.rds"),
          to = destination,
          overwrite = TRUE)


## serial interval distribution
saveRDS(serial_interval_fit,
        file = file.path("produced_rds",
                         "serial_interval_fit.rds"))
destination <- here("data", "clean", "serial_interval_fit.rds")

file.copy(from = file.path("produced_rds",
                           "serial_interval_fit.rds"),
          to = destination,
          overwrite = TRUE)


## reporting interval distribution
saveRDS(reporting_interval_fit,
        file = file.path("produced_rds",
                         "reporting_interval_fit.rds"))
destination <- here("data", "clean", "reporting_interval_fit.rds")

file.copy(from = file.path("produced_rds",
                           "reporting_interval_fit.rds"),
          to = destination,
          overwrite = TRUE)


```






## Excel files

We export the table containing information on: 

* the cases (linelist info)
* the effective reproduction numbers ($R$)
* the identifiers of the transmission chains

```{r xlsx_exports}

if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

rio::export(R_df,
            file = file.path("produced_xlsx",
                             "linelist_with_R_and_chains.xlsx"))

```


**Links to the files:**

- [linelist_with_R_and_chains.xlsx](produced_xlsx/linelist_with_R_and_chains.xlsx)



## visNetwork visualisation

We save the plot of the transmission chain as a separate file called
`plot_chains.html`:

```{r export_html}

visNetwork::visSave(plot_x,
                    file = "plot_chains.html")

visNetwork::visSave(plot_x_active,
                    file = "plot_active_chains.html")

```




<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# System information {.tabset .tabset-fade .tabset-pills}

## Outline

The following information documents the system on which the document was
compiled.



## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```



## ChangeLog

### version 1.0.0

* initial version of the report


### version 1.1.0

* filter out duplicated chains

* no longer plotting the whole dataset in the report, but still outputting to
  `plot_chains.html`
  
* added a recent dataset (last 6 weeks)
