---
title: "Analyses of transmission chains"
author: "Thibaut, Maddy, Sam, Luc, Sonia, Elias, Xavier, the Butembo team, the analytic cell Goma"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 4
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE)
```



<br>

<div class="report_meta">
  <span class="notice">**Notice**: this is a **stable, routine report**. 
  **Do not touch it unless it is broken.** To make a contribution, **carefully read 
  the [README](../../../../../README.html) file**.</span>
  
  **Maintainer:** Thibaut Jombart (thibautjombart@gmail.com)
  
  **Code contributors:** Thibaut Jombart, Amy Gimma
  
  **Data contributors:** transmission chains teams from Butembo, Mangina, Beni, HQ
  
  **Version:** 1.1.0
  
  **Reviewed by:**
</div>




<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Data preparation {.tabset .tabset-fade .tabset-pills}

<!-- ============================================ -->
## Outline

This report focusses on transmission chains of the outbreak, derived from the
narratives. This work was pioneered by the Butembo team, who assembled specific
linelists with transmission information. To simplify this workflow for other
teams (with less data management resources), we streamline this process as
follows:

1. information on the cases are taken from the clean linelist data (the *master
   linelist*), thus not requiring additional work from the field teams
   
2. information on transmissions is stored in a separate file following the same
   model as the master linelist, i.e. a google spreadsheet which can be edited
   by different data managers; this file only contains information on the
   *transmissions*, not on the cases, e.g.: from whom, to whom, which kind of
   transmission context etc.
   
The two sources of data will be merged using *epicontacts*.



### Data used

We combine two data sources:

* the *master linelist*, containing information about the cases

* the *master transmission list*, describing transmission chains using unique
  identifiers for source cases and secondary cases
  

### Method

The data preparation involves the following steps, detailed in the following tabs:


* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Clean data**: this section contains *ad hoc* data cleaning, i.e. which is
  not used in other reports (otherwise cleaning should be done in a dedicated
  report); this section is also used to create new variables used in the
  analyses




<!-- ============================================ -->
## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`
* the path to the cleaned VHF data stored as `x`

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$",
                     full.names = TRUE)
for (file in scripts_files) source(file, local = TRUE)

```




<!-- ============================================ -->
## Load data

Two types of data are used in this report:

1. the clean *master linelist*, called `linelist` in this report
2. the raw *master transmission list*, called `chains` in this report

```{r load_data}

## load the master linelist ('MLL')
current_linelist
linelist <- rio::import(current_linelist) %>%
    as_tibble()
linelist

## load the chains
current_chains
chains_raw <- custom_import(current_chains) %>%
    as_tibble()
chains_raw

## extract completion dates
linelist_file_name <- gsub("^[^.]+/", "", current_linelist)
linelist_date <- linelist_file_name %>%
  guess_dates()
linelist_date

chains_file_name <- gsub("^[^.]+/", "", current_chains)
chains_date <- chains_file_name %>%
  guess_dates()
chains_date

```

The **completion dates** of the databases are:

* master linelist: **`r format(linelist_date, format = "%A %d %b %Y")`**
* master transmission: **`r format(chains_date, format = "%A %d %b %Y")`**





<!-- ============================================ -->
## Clean data

As we use a clean master linelist, cleaning only applies here to the *master
transmission list* stored in the object called `chains_raw`. Once cleaned,
transmission data will be stored as `chains`.

We use *linelist*'s function `clean_data()` to:

- remove all special characters from the data
- set all characters to lower case
- replace all accentuated and diacritic characters with their closest ascii
  match in the latin alphabet
- replace all separators with a single `_`
- replace all mis-spelling using a overall dictionary (see the file
  `cleaning_rules.xlsx` in `/dictionary/`, or type `cleaning_rules` in this R
  session, after compiling the document
- (optionally) detect date formats and convert data to dates, including cases
  where format varies within a column
  

This cleaning is achieved with:

```{r data_cleaning}

chains <- chains_raw %>%
  clean_data(guess_dates = FALSE) %>%
  as_tibble()

chains

```




<!-- ============================================ -->
## Cleaning using dictionary

Here we correct typos etc using our dictionary. We then tabulate the categories
of key variables to make sure nothing has been left out. New items to clean
should be added to the dictionary in `dictionary/cleaning_rules.xlsx`.

```{r clean_dic}

## clean using dictionary
chains <- chains %>%
  clean_variable_spelling(wordlists = cleaning_rules)

## check variables
table(chains$type_contamination, useNA = "ifany")
table(chains$lien_cas_source, useNA = "ifany")

```




<!-- ============================================ -->
## Removing duplicate chains

As data entry in the master transmission list is performed by multiple people,
duplicated chains do occur (same combination of *source -> secondary
case*). Here we:

1. identify duplicates

2. output a table of duplicates which is later exported to `xlsx`

3. remove duplicates


```{r duplicates}

## step 1
chains <- chains %>%
  mutate(chain_txt = paste(source_case_id, case_id, sep = "->"),
         duplicate = duplicated(chain_txt))

## step 2
table_duplicates <- chains %>% 
  filter(chain_txt %in% chain_txt[duplicate]) %>%
  arrange(case_id, source_case_id) %>%
  select(-(chain_txt:duplicate))

table_duplicates  %>%
  show_table()

## step 3
chains <- chains %>% 
  filter(!duplicate)

```





<!-- ============================================ -->
## Removing self-infections

We also identified, at some point, some self-infections. As for duplicates, we:

1. identify these

2. report the self-infections in a table to be later exported

3. remove them from the data

```{r self_infections}

## step 1
chains <- chains %>%
  mutate(self_infection = case_id == source_case_id)

## step 2
table_self_infections <- chains %>% 
  filter(self_infection) %>%
  select(-(chain_txt:self_infection))
  
table_self_infections  %>%
  show_table()

## step 3
chains <- chains %>% 
  filter(!self_infection)

```






<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Building transmission chains {.tabset .tabset-fade .tabset-pills}

<!-- ============================================ -->
## Outline

In this section, we merge MTL and MLL data into an *epicontacts* object,
summarise the resulting object and provide some interactive visualisation.



<!-- ============================================ -->
## Making an *epicontacts* object

We build the *epicontacts* object using:

* the *master linelist* stored in `linelist`, keeping only selected variables
  useful for further analyses, and entries for which we are missing VHF
  identifiers (`vhf_code`), as well as duplicates
  
* the *master transmission list* data stored in `chains`

```{r epicontacts}

select_linelist <- linelist %>%
  rename(mll_id = mllid,
         vhf_id = vhf_code) %>% 
  select("vhf_id",
         "mll_id",
         "epicasedef",
         "date_onset",
         "date_report",
         "date_death",
         "date_admission",
         "date_exposure_start",
         "date_exposure_end",
         "zone_de_sante",
         "aire_de_sante",
         "age",
         "age_class",
         "gender",
         "hcw",
         "firstname",
         "surname",
         "current_status",
         "report_status",
         "outcome",
         "hosp1_name",
         "sous_coordination") %>%
  filter(!is.na(vhf_id),
         !duplicated(vhf_id))

x <- make_epicontacts(linelist = select_linelist,
                      contacts = chains,
                      directed = TRUE,
                      id = "vhf_id",
                      from = "source_case_id",
                      to = "case_id")
x

```




<!-- ============================================ -->
## Identify clusters / chains

We define clusters as sets of cases connected by transmission events. In graph
terminology, these are the connected components of the graph. We also:

* identify **active clusters**, defined as clusters having at least one new case
notified within the last 21 days

* identify the main active clusters 

```{r clusters}

x <- get_clusters(x,
                  member_col = "cluster_id",
                  size_col = "cluster_size",
                  override = TRUE)

## re-label the cluster ids
x$linelist$cluster_id <- paste("chain", x$linelist$cluster_id, sep = "_")

```




<!-- ============================================ -->
## Add degrees

Here we compute the in- and out-degrees for every case, defined as:

* *in-degree*: number of source cases for the case; zero indicates a *root*

* *out-degree*: number of secondary cases for the case, i.e. effective
  reproduction number (*R*)


```{r add_degrees}

in_degree <- data.frame(in_degree = get_degree(x, "in")) %>%
  rownames_to_column(var = "id")
out_degree <- data.frame(out_degree = get_degree(x, "out")) %>%
  rownames_to_column(var = "id")

# Adds in_degree and out_degree to x$linelist, using the id variable for the joins
# Also adds a variable in to determine whether or not an individual is a root or not
x$linelist <- x$linelist %>%
  left_join(in_degree) %>%
  left_join(out_degree) %>%
  mutate(is_root = (in_degree == 0),# & (out_degree >= 0),
         root = if_else(is_root, id, NULL))

```




<!-- ============================================ -->
## Active chains (6 weeks)

We define as *active chains* the transmission chains which have reported at
least one case in the last 42 days (prior to completion of the master
linelist), here corresponding to the 
**`r format(chains_date - 42, format = "%A %d %b %Y")`**).
   

The resulting chains are stored as a new object `x_active`.

```{r active_chains}


## set limit date
date_recent <- linelist_date - 42

## find which clusters are active
x$linelist <- x$linelist %>%
  group_by(cluster_id) %>%
  mutate(active_cluster = any(date_report > date_recent, na.rm = TRUE))

x_active <- x
x_active$linelist <- x_active$linelist %>%
  filter(active_cluster)

x_active <- x_active %>%
  thin(what = "contacts")

```





<!-- ============================================ -->
## Summary

This provides an overview of the content of the *epicontacts* object. Pay
attention, in particular, to cases that are in the transmission tree (called
"contacts" in *epicontacts* terminology) but not in the linelist. These could be
recovered by using a larger database for linelist, such as VHF or the new MLL
(Master Line List) currated in Goma.

```{r summary}
summary(x)
```




<!-- ============================================ -->
## General plot

This is a general visusalisation of the whole dataset. Because of its large
size, we do not show the full plot in this document, but ouput it in a separate
file [`plot_chains.html`](plot_chains.html). 

```{r plot_chains}

plot_x <- plot(x,
               node_color = "zone_de_sante",
               node_shape = "gender",
               shapes = c(female = "female",
                          male = "male",
                          unknown = "question-circle"),
               legend_max = 100,
               edge_color = "type_contamination"
               )

```




<!-- ============================================ -->
## Plot active chains

Graphs will be shown using the date of isolation (`date_admission`) if no value
is missing; otherwise the default plotting of transmissions will be used.


```{r plot_active_chains_2}
## check if some date_admission are missing
missing_date <- get_pairwise(x_active, "date_admission") %>% is.na()

if (!any(missing_date)) {
  x_axis_info <- "date_admission"
} else {
  x_axis_info <- NULL
}

## make the plot
plot_x_active <- plot(x_active,
                      node_color = "zone_de_sante",
                      node_shape = "gender",
                      shapes = c(female = "female",
                                 male = "male",
                                 unknown = "question-circle"),
                      legend_max = 100,
                      edge_color = "type_contamination",
                      x_axis = x_axis_info,
                      edge_width = 5
                      )
plot_x_active

```









<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Descriptive analyses {.tabset .tabset-fade .tabset-pills}

<!-- ============================================ -->
## Outline

In the following we provide some simple descriptions of the transmission
network. We focus on spatial patterns in transmissions, comparing the locations
of the sources cases and their secondary cases, and the relative frequencies of
these transmissions.




<!-- ============================================ -->
## By sex

```{r contacts_sex}

pairs_sex <- get_pairwise(x,
                          "gender",
                          function(from, to)
                            as.data.frame(table(from, to)))

ggplot(pairs_sex,
       aes(y = from, x = to, size = Freq)) +
  theme_bw() +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 20),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(45) + 
  labs(title = "Transmissions between sex",
       y = "Source case",
       x = "Infectee")

```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_sex_circi}

## Chord diagram
circle_plot(pairs_sex, label_space = 2)

```




<!-- ============================================ -->
## By age class

```{r contacts_age}

pairs_age <- get_pairwise(x,
                          "age_class",
                          function(from, to)
                            as.data.frame(table(from, to)))

ggplot(pairs_age,
       aes(y = from, x = to, size = Freq)) +
  theme_bw() +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 20),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(45) + 
  labs(title = "Transmissions between Age Groups",
       y = "Source case",
       x = "Infectee")

```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_age_circi}

## Chord diagram
circle_plot(pairs_age, label_space = 2)

```
 
 

<!-- ============================================ -->
## By health zones

We plot transmissions by health zones:

```{r contacts_zone_de_sante}

pairs_zone_de_sante <- get_pairwise(x,
                                    "zone_de_sante",
                                    function(from, to)
                                      as.data.frame(table(from, to)))

ggplot(pairs_zone_de_sante,
       aes(y = from, x = to, size = Freq)) +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 20),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  rotate_x_text(45) + 
  labs(title = "Transmissions across health zones",
       y = "Source case",
       x = "Infectee")

```

We provide the alternative circular visualisation; colors of the flows are
indicative of the source.

```{r contacts_zone_de_sante_circi}

## Chord diagram
circle_plot(pairs_zone_de_sante, label_space = 2)

```

We compute the proportion of transmission happening across different health
zones:

```{r move_zones}

move_zone <- get_pairwise(x,
                          "zone_de_sante",
                          function(from, to)
                            as.logical(from != to))
move_zone

## mean and 95% CI
p_move_health_zone <- round(100 * mean(move_zone, na.rm = TRUE))
p_move_health_zone
ci_move_health_zone <- prop_ci(k = sum(move_zone, na.rm = TRUE),
                               n = sum(!is.na(move_zone)),
                               perc = TRUE,
                               dec = 0)
ci_move_health_zone

```

Overall, `r p_move_health_zone`% (IC95%: `r ci_move_health_zone[1]`%-`r ci_move_health_zone[2]`%) of transmissions happen cross different health zones.





<!-- ============================================ -->
## By health areas

We plot transmissions by health areas:

```{r contacts_aire_de_sante, fig.width = 12}

pairs_aire_de_sante <- get_pairwise(x,
                                    "aire_de_sante",
                                    function(from, to)
                                      as.data.frame(table(from, to)))

ggplot(pairs_aire_de_sante,
       aes(y = from, x = to, size = Freq)) +
  theme_bw() +
  geom_point(color = "navy", alpha = .5) +
  scale_size_continuous("Number of \ntransmissions",
                        range = c(1, 10),
                        limits = c(1, NA),
                        breaks = c(1, 5, 10, 20, 30, 40)) +
  large_txt +
  smaller_axis_txt +
  rotate_x_text(90) + 
  labs(title = "Transmissions across health areas",
       y = "Source case",
       x = "Infectee")


```


We compute the proportion of transmission happening across different health
areas:

```{r move_areas}

move_area <- get_pairwise(x,
                          "aire_de_sante",
                          function(from, to)
                            as.logical(na.omit(from != to)))
move_area

## mean and 95% CI
p_move_health_area <- round(100 * mean(move_area))
p_move_health_area
ci_move_health_area <- prop_ci(k = sum(move_area),
                               n = length(move_area),
                               perc = TRUE,
                               dec = 0)
ci_move_health_area

```

Overall, `r p_move_health_area`% (IC95%: `r ci_move_health_area[1]`%-`r ci_move_health_area[2]`%) of transmissions happen cross different health areas.





<!-- ============================================ -->
## Sources and recipients of transmissions

In this section we characterise health areas in terms of the number of
transmissions given ("*sources*"), or received ("*recipients*").

```{r sources_recipients, fig.height = 8, fig.width = 12}

## count cases sourced by area
df_sources <- pairs_aire_de_sante %>%
  filter(as.character(from) != as.character(to), Freq > 0) %>%
  group_by(from) %>%
  summarise(n_sourced = sum(Freq)) %>%
  rename(aire_de_sante = from)

## count cases received by area
df_receivers <- pairs_aire_de_sante %>%
  filter(as.character(from) != as.character(to), Freq > 0) %>%
  group_by(to) %>%
  summarise(n_received = sum(Freq)) %>%
  rename(aire_de_sante = to)

## put the information together
table_sources_recipients <- full_join(df_sources, df_receivers) %>%
  filter(aire_de_sante != "inconnu") %>% 
  replace_na(list(n_sourced = 0, n_received = 0)) %>%
  arrange(desc(n_sourced), desc(n_received))

## add info on zones
zone_info <- linelist %>%
  group_by(aire_de_sante, zone_de_sante) %>%
  count() %>%
  group_by(aire_de_sante) %>%
  top_n(1, n) %>%
  ungroup() %>% 
  select(1:2)

table_sources_recipients <- right_join(zone_info, table_sources_recipients)

table_sources_recipients_top <- table_sources_recipients %>%
  mutate(sum = n_sourced + n_received) %>%
  group_by(zone_de_sante) %>% 
  top_n(5, sum) %>%
  arrange(zone_de_sante, sum) %>%
  mutate(rank = 1:n()) %>% 
  top_n(4, rank)


## make the plot
ggplot(table_sources_recipients,
       aes(x = n_received,
           y = n_sourced,
           color = aire_de_sante)) +
  theme_bw() +
  geom_point() +
  geom_label_repel(data = table_sources_recipients_top,
                   aes(label = aire_de_sante), size = 4) +
  scale_x_continuous(breaks = int_breaks) +
  scale_y_continuous(breaks = int_breaks) +
  large_txt +
  smaller_axis_txt +
  theme(strip.text = element_text(size = 16)) +
  facet_wrap(~ zone_de_sante) +
  scale_color_discrete(guide = FALSE) +
  labs(title = "Donneur et receveurs de cas",
       x = "Nombre de cas reçus",
       y = "Nombre de cas sourcés")

## display the table
table_sources_recipients %>%
  show_table()


```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Analysis of transmissibility {.tabset .tabset-fade .tabset-pills}

<!-- ============================================ -->
## Reproduction number

We can assess the effective reproduction number ($R$) by looking at the numbers
of secondary cases by index case:

```{r R}

## in-degree
x %>%
  get_degree() %>%
  table()

## out-degree = reproduction number
x %>%
  get_degree("out") %>%
  table()

R_df <- x %>%
  get_degree("out") %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  rename("id" = "rowname", "R" = ".")

head(R_df)

```

We create a new `data.frame` with R values and case info from the linelist;
unlike the linelist, where each line is a case, here each line is a source case:

```{r R_df}

R_df <- left_join(R_df, x$linelist, by = "id") %>%
  as_tibble()

```

Because of delays from primary to secondary onset, and from onset to
notification, we need to ignore the latest source cases to avoid
under-estimating the $R$. We use a conservative limit of 45 days:

```{r R_plots}

old_cases <- R_df %>%
  filter(date_onset < (chains_date - 45))

## graph
ggplot(old_cases, aes(x = R)) +
  theme_bw() +
  geom_bar() +
  large_txt +
  labs(title = "Nombre de cas secondaire par cas index",
       x = "Nombre de cas secondaires observés (R)",
       y = "Fréquence")


## show table
R_df %>%
  show_table()

```




<!-- ============================================ -->
## Effective reproduction number in time

```{r R_in_time}

ggplot(R_df, aes(x = date_report, y = R)) +
  geom_point() +
  geom_smooth() +
  theme_bw() +
  large_txt +
  scale_months +
  rotate_x_text(45) +
  labs(title = "Nombre de reproduction effectif au cours du temps",
       x = "Date de notification",
       y = "Nombre de cas secondaires par cas index")

```







<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Delay distributions {.tabset .tabset-fade .tabset-pills}

In this section, we estimate delays characterising transmission pairs. The
procedure is the same for all delays:

1. compute delays from the data to obtain an **empirical distribution** (histograms)

2. inspect the empirical distribution for **outliers**, and remove them if needed

3. estimate the underlying distribution by **fitting a discretised Gamma**
   distribution to the empirical delays


## Serial interval distribution

The serial interval is defined as the delay between symptom onset between
primary (source) cases and secondary cases.

We first derive the empirical distribution:

```{r serial_interval_empirical}

## extract empirical data
serial_interval_empirical <- x %>%
  get_pairwise("date_onset") %>%
  na.omit() %>%
  as.integer()

## remove negative delays and delays > 50 days
to_keep <- serial_interval_empirical > 0 & serial_interval_empirical < 50
serial_interval_empirical <- serial_interval_empirical[to_keep]

## make graph
serial_interval_base <- serial_interval_empirical %>%
  tibble() %>%
  rename(serial_interval = ".") %>%
  ggplot(aes(x = serial_interval)) +
  geom_bar(aes(y = stat(count / sum(count))),
           color = "white", fill = "navyblue", alpha = .3) +
  theme_bw() +
  large_txt +
  labs(title = "Intervalle sériel: distribution empirique",
       x = "Jours entre l'apparition des symptômes \ndes cas primaires et secondaires",
       y = "Nombre de paires de transmission")
serial_interval_base

```

We fit a discretized Gamma distribution to the data:

```{r serial_interval_fit}

serial_interval_fit <- serial_interval_empirical %>%
  fit_disc_gamma(mu_ini = mean(serial_interval_empirical))

serial_interval_fit

## check that the fitting worked
if (!serial_interval_fit$converged) {
  stop("Fitting of serial interval did not converged")
}

## make plot with fitting
serial_interval_base +
  stat_function(fun = serial_interval_fit$distribution$d,
                color = "#cc6699", lwd = 1) +
  labs(title = "Distribution de l'intervalle sériel")

```

The serial interval is characterised by:

* mean of **`r round(serial_interval_fit$mu, 1)`** days
* standard deviation of **`r round(serial_interval_fit$sd, 1)`**
* **50% of the delays** are between **`r serial_interval_fit$distribution$q(.25)`** 
and **`r serial_interval_fit$distribution$q(.75)`** days
* **95% of the delays** are between **`r serial_interval_fit$distribution$q(.025)`** 
and **`r serial_interval_fit$distribution$q(.975)`** days






<!-- ============================================ -->
## Reporting interval distribution

This distribution is a variant of the serial interval, where dates of reporting
are used instead of dates of symptom onset.

We first derive the empirical distribution:

```{r reporting_interval_empirical}

## extract empirical data
reporting_interval_empirical <- x %>%
  get_pairwise("date_report") %>%
  na.omit() %>%
  as.integer()


## remove negative delays and delays > 50 days
to_keep <- reporting_interval_empirical > 0 & reporting_interval_empirical < 50
reporting_interval_empirical <- reporting_interval_empirical[to_keep]

## make graph
reporting_interval_base <- reporting_interval_empirical %>%
  tibble() %>%
  rename(reporting_interval = ".") %>%
  ggplot(aes(x = reporting_interval)) +
  geom_bar(aes(y = stat(count / sum(count))),
            color = "white", fill = "navyblue", alpha = .3) +
  theme_bw() +
  large_txt +
  labs(title = "Intervalle de notification: distribution empirique",
       x = "Jours entre la notification \ndes cas primaires et secondaires",
       y = "Nombre de paires de transmission")
reporting_interval_base

```

There are no indication of outliers, so we can go ahead and fit a discretized
Gamma to the data:

```{r reporting_interval_fit}

reporting_interval_fit <- reporting_interval_empirical %>%
  fit_disc_gamma(mu_ini = mean(reporting_interval_empirical))

## check that the fitting worked
if (!reporting_interval_fit$converged) {
  stop("Fitting of reporting interval did not converged")
}

## make plot with fitting
reporting_interval_base +
  stat_function(fun = reporting_interval_fit$distribution$d,
                color = "#cc6699", lwd = 1) +
  labs(title = "Distribution de l'intervalle de notification")

```


The reporting interval is characterised by:

* mean of **`r round(reporting_interval_fit$mu, 1)`** days
* standard deviation of **`r round(reporting_interval_fit$sd, 1)`**
* **50% of the delays** are between **`r reporting_interval_fit$distribution$q(.25)`** 
and **`r reporting_interval_fit$distribution$q(.75)`** days
* **95% of the delays** are between **`r reporting_interval_fit$distribution$q(.025)`** 
and **`r reporting_interval_fit$distribution$q(.975)`** days




## Comparison of distributions

In this section, we compare the different distributions estimated from the
transmission chains.

```{r comparison}

max_delay <- max(serial_interval_empirical,
                 reporting_interval_empirical) + 5
delays <- 0:max_delay

distributions <- tibble(
    delays,
    serial_interval = serial_interval_fit$distribution$d(delays),
    reporting_interval = reporting_interval_fit$distribution$d(delays)) %>%
  gather(distribution, proba, -1)


ggplot(distributions,
       aes(x = delays, y = proba, color = distribution)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  theme_bw() +
  large_txt +
  labs(title = "Comparaison des distributions",
       x = "Nombre de jours",
       y = "Probabilité") +
  theme(legend.position = c(.8, .8))

```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Export outputs {.tabset .tabset-fade .tabset-pills}

<!-- ============================================ -->
## Outline

We save several outputs to separate files, including a `html` files for the
transmission chain plots, and R objects (`.rds` format) for fitted delay
distributions.



<!-- ============================================ -->
## R objects

We export the following items as R objects, saved into `produced_rds`:

* **epicontacts.rds**: the *epicontacts* object

* **serial_interval_fit.rds**: the serial interval distribution

* **reporting_interval_fit.rds**: the reporting interval distribution


```{r rds_exports}

if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}

## epicontact object
saveRDS(x,
        file = file.path("produced_rds",
                         "epicontacts.rds"))
destination <- here("data", "clean", "epicontacts.rds")

file.copy(from = file.path("produced_rds",
                           "epicontacts.rds"),
          to = destination,
          overwrite = TRUE)


## serial interval distribution
saveRDS(serial_interval_fit,
        file = file.path("produced_rds",
                         "serial_interval_fit.rds"))
destination <- here("data", "clean", "serial_interval_fit.rds")

file.copy(from = file.path("produced_rds",
                           "serial_interval_fit.rds"),
          to = destination,
          overwrite = TRUE)


## reporting interval distribution
saveRDS(reporting_interval_fit,
        file = file.path("produced_rds",
                         "reporting_interval_fit.rds"))
destination <- here("data", "clean", "reporting_interval_fit.rds")

file.copy(from = file.path("produced_rds",
                           "reporting_interval_fit.rds"),
          to = destination,
          overwrite = TRUE)


```






<!-- ============================================ -->
## Excel files

We export several tables generated in this document.

```{r xlsx_exports}

if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

to_export <- c("table_duplicates",
               "table_self_infections",
               "table_sources_recipients")



for (e in to_export) {
  rio::export(get(e),
              file.path("produced_rds",
                        paste0(e, ".rds")))
}

```

Click on the following links to open the files (only works if the files above
have been generated and are in the same folder as this document):


```{r links, results = "asis", eval = FALSE}

for (e in to_export) {
  txt <- sprintf("- [%s.rds](%s.rds)",
                 e,
                 file.path("produced_rds",
                           e))
  cat(txt, sep = "\n")
}

```




<!-- ============================================ -->
## visNetwork visualisation

We save the plot of the transmission chain as a separate file called
`plot_chains.html`:

```{r export_html}

visNetwork::visSave(plot_x,
                    file = "plot_chains.html")

visNetwork::visSave(plot_x_active,
                    file = "plot_active_chains.html")

```






<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# System information {.tabset .tabset-fade .tabset-pills}

<!-- ============================================ -->
## Outline

The following information documents the system on which the document was
compiled.



## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```



<!-- ============================================ -->
## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```



<!-- ============================================ -->
## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```



<!-- ============================================ -->
## ChangeLog

### version 1.0.0

* initial version of the report


### version 1.1.0

* identify, output and filter out duplicated chains

* identify, output and filter out self-infections

* no longer plotting the whole dataset in the report, but still outputting to
  `plot_chains.html`
  
* added 'recent active chains' (chains with a case reported in the last 42 days)

* active chains are no longer named after their root, as there may not be a
  single one, or any
  
* removed estimations of reproduction number

* polished plots of interval distributions (using `geom_bar` rather than
  `geom_histogram`)

* batch-export of `xlsx` and automated insertion of links to files
