---
title: "Investigation of alerts data: Komanda"
author: "Charlie Whittaker, Christopher Jarvis,  Thibaut Jombart, Flavio Finger, Jonathan Polonsky, Patrick Keating, Amy Gimma and Aminata Ndiaye for the analytic cell OEC Goma"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document: 
    code_folding: hide
    highlight: zenburn
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 1
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6, 
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "figures/")
```


<br>

**Maintainer:** Christopher Jarvis (christopher.jarvis@lshtm.ac.uk)

**Code contributors:** Chris Jarvis, Charlie Whittaker

**Data contributors:** Surveillance team

**Version:** 1.0.0 

**Reviewed by:** Thibaut Jombart

**Notice**: this is a **stable, routine report**. **Do not touch it unless it is
broken.** To make a contribution, carefully read the
[README](../../../../../README.html) file.




<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Data preparation {.tabset .tabset-fade .tabset-pills}


<!-- ======================================================= -->
## Outline

This report cleans and analyses the alert data of Komanda. Input comes from an
`xlsx` file containing alerts in a specific format. Because all
sub-coordinations have different standards, each sub-coordination needs a
separate report.

The data preparation involves the following steps, detailed in the following 
tabs:

* **Load scripts**: loads libraries and useful scripts used in the analyses; all
  `.R` files contained in `scripts` at the root of the factory are automatically
  loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.

* **Clean data**: this section contains *ad hoc* data cleaning, i.e. which is
  not used in other reports (otherwise cleaning should be done in a dedicated
  report); this section is also used to create new variables used in the
  analyses

## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all Overall scripts, i.e. stored outside the factory in `../scripts/`

```{r read_scripts}

## read scripts
path_to_scripts <- here::here("scripts")
scripts_files <- dir(path_to_scripts, pattern = ".R$",
                     full.names = TRUE)
for (file in scripts_files) source(file, local = TRUE)

ggthemr("fresh")

```


<!-- ======================================================= -->
## Load alerts data

We extract the completion date from the file name:

```{r load_alerts_data}

## load the data
current_komanda
x_raw <- custom_import(current_komanda)
glimpse(x_raw)

## extract database date from the file name
file_name <- gsub("^[^.]+/", "", current_komanda)
database_date <- file_name %>%
  guess_dates()
database_date

```

The **completion date** of the database is **`r format(database_date, format =
"%A %d %b %Y")`**.

<!-- ======================================================= -->
## Clean data

We use *linelist*'s function `clean_data()` to:

- remove all special characters from the data
- set all characters to lower case
- replace all accentuated and diacritic characters with their closest ascii
  match in the latin alphabet
- replace all separators with a single `_`
- replace all mis-spelling using a Overall dictionary (see the file
  `cleaning_rules.xlsx` in `/dictionary/`, or type `cleaning_rules` in this R
  session, after compiling the document
- (optionally) detect date formats and convert data to dates, including cases
  where format varies within a column
  

*NOTE:* Exceptionally for Komanda, additional data management was done. Given the clear unreliability of the variable `origin`, rules were made to decide when the alert were most liekly to be a _surveillance passive_ or a _recherche active_ using information in other variables as 

This cleaning is achieved with:

```{r data_cleaning}

# x <- x_raw %>%
#   mutate(origin_not_cleaned = origine_alerte) %>%
#   as_tibble()

x <- x_raw %>%
  clean_data(guess_dates = FALSE,
             wordlists = cleaning_rules) %>%
  as_tibble()


#Now modify the type of transmission of the alert (by using info in type and origin) with lot of unknown and lot of "recherche_active" even though the  origin in many cases precise a healthcare place 
##When there is PS or CS or UTI, HGR, POE: it is surv passive ie healthcare services 

#first cretate a cat var that says whether or not a healthcare place was implicated
x <- x %>%
  mutate(hc_place= ifelse(
                    grepl(c("cm|ps|cs|poe|hgr|uti|clini|infir"), origine_alerte, ignore.case=F), "oui", "non")
  )


x <- x %>%
  mutate(origin_cleaned = ifelse(
    typ_transmission %in% c(NA,"inconnu") & (hc_place=="oui" | origine_alerte =="sbc"), "surveillance_passive", typ_transmission)
  ) 

# %>%
#   select(-origine_alerte)


```
<!-- ======================================================= -->
## Renaming variables and sanity checks

The following variables will be used, and are therefore checked:

- `date`: check no date is after the database completion, filtered for NA values after running `guess_dates`
- sanitize `genre`
- make `origin` as a sanitized version of `source_notif`

```{r check_variables}

## date of notification
x <- x %>% mutate(date = guess_dates(date))
range(x$date, na.rm = TRUE)

# REVIEW: Remove NA dates
date_na <- sum(is.na(x$date))
x <- x %>% filter(!is.na(date))
# Check that na dates are removed

## gender
table(x$sexe, useNA = "ifany")

## contact connu
x <- mutate(x, contact_connu = lien_epi)
table(x$contact_connu, useNA = "ifany")

## zone de sante
x <- mutate(x, zone_de_sante = as.character(zone_sante)) 
table(x$zone_de_sante, useNA = "ifany")

## aire de sante
x <- mutate(x, aire_de_sante = as.character(aire_sante)) 
table(x$aire_de_sante, useNA = "ifany")

## origin / source_notif
x <- mutate(x, origin = origin_cleaned)
table(x$origin, useNA = "ifany")

# create variable for alert validation
x <- x %>% mutate(
  status = ifelse(result_invest == "inconnu", "statut_inconnu", result_invest)) %>%
  mutate(status = factor(status, levels = c("statut_inconnu", 
                                            "invalidee", 
                                            "validee")))
table(x$status, useNA = "ifany")

# another round of cleaning 
x <- x %>%
  clean_variable_spelling(wordlists = cleaning_rules)

```


There are `r length(date_na)` missing dates in the data.


<!-- ======================================================= -->
## Variable creation

The following variables are created: 

- `bleed` - which describes whether the alert displayed any of the bleeding related
symptoms.
- `epiweek_report` & `epiweek_report_label` - which describe the epiweek the alert
was reported in.
- `top_zones` - which describes the 3 most active Zone de Santes (other Zone de 
Santes) are grouped into "other".
- `top_aires` - which describes the 14 most active Aire de Santes (out of all the
legitimate Aire de Santes belonging to Komanda present in the database). Note - All 
other Aire de Santes are aggregated into the category "other".

- `final_outcome` - which describes the final status of the alert after a test has (most likely) been done.


Note - All other Aire de Santes are aggregated into the category "other".

``` {r variable_creation}

## bleed
bleed <- x %>%
  select(contains("saignement")) %>%
  apply(1, function(e) ifelse(any(e == "oui"), "oui", "non"))
x <- mutate(x, bleed = bleed)
table(x$bleed, useNA = "ifany")

## epiweek report
x <- x %>%
  mutate(epiweek_report = 
           aweek::date2week(date, week_start = "Monday", floor_day = TRUE)) %>%
  mutate(epiweek_report_label = 
           aweek::week2date(date, week_start = "Monday", floor_day = TRUE)) 



x <- x %>% 
  mutate(top_zones = top_values(zone_de_sante, 5)) %>%
  mutate(top_aires = top_values(aire_de_sante, 14))



#create a variable for the final status of the alert
x <- x %>%
  mutate(final_outcome = classif_final) %>%
  mutate(final_outcome = factor(final_outcome, levels = c("confirme", 
                                                          "non_cas", 
                                                          "suspect")))

```

One last round of dictionary-based cleaning:

```{r last_cleaning}

x <- x %>%
  clean_variable_spelling(wordlists = cleaning_rules)

```

Remove data from before 2019-01-01 as not as reliable. 


```{r remove_pre_2019}

x <- x %>%
  filter(date >= "2019-01-01")

```

<!-- ======================================================= -->
## Outcomes

Outcome is defined for alerts which have either been validated or invalidated.

```{r outcomes}

## get only known outcomes
outcomes <- x %>%
  filter(status %in% c("validee", "invalidee")) %>%
  droplevels() %>% 
  mutate(validee = 1 * (status == "validee"))

```

<!-- ======================================================= -->
## Check admin areas are correct

```{r load_area_names_data}

## load the data
current_spatial
area_names <- rio::import(current_spatial) %>%
  clean_data(guess_dates = FALSE)

as_names <- unique(area_names$as)
zs_names <- unique(area_names$zs)

```

### Identify unknown areas  

```{r identify_area_names_data}

## load the data
n_rows <- nrow(x)

unknown_as <- x %>% 
  filter(!aire_de_sante %in% as_names)

most_recent_unknown_as <- max(unknown_as$date, na.rm = TRUE)

start_date <- database_date - 20
unknown_as_3weeks <- x %>% 
  filter(!aire_de_sante %in% as_names,
         date >= start_date)

```


There are `r nrow(unknown_as)` alerts with unknown aire de santes with the latest being `r most_recent_unknown_as`. There are `r nrow(unknown_as_3weeks)` alerts in the last 3 weeks with an unknown aire de santes in the database.

```{r show_unknown_area_names}

table_unknown_as <- unknown_as %>% 
  count(zone_de_sante, aire_de_sante) 
  
table_unknown_as %>% 
  show_table()

```

<!-- ======================================================= -->
## Filter to known aire de santes

```{r known_aire_de_sante}

x <- x %>% 
  filter(aire_de_sante %in% as_names)

```


<!-- ======================================================= -->
## Add population data

Load the population data

```{r load_pop_data}

## load the data
current_pop 
pop_data <- rio::import(current_pop)
glimpse(pop_data)

pop_data <- pop_data %>%
  clean_data(guess_dates = FALSE) %>%
  as_tibble()
```

### Check for area not in alerts

```{r filter_by_alerts}

## load the data
pop_data <- pop_data %>% 
  filter(aire_de_sante %in% x$aire_de_sante,
         zone_de_sante %in% x$zone_de_sante)

pop_data

## aire de sante population
aire_de_sante_pop <- pop_data %>%
  mutate(as_population = population) %>%
  select(zone_de_sante, aire_de_sante, as_population)

## zone de sante population
zone_de_sante_pop <- pop_data %>%
  group_by(zone_de_sante) %>%
  summarise(zs_population = sum(population, na.rm = TRUE))

total_population <- sum(unique(pop_data$population), na.rm = TRUE)

## adding populations to main data
x <- x %>%
  left_join(aire_de_sante_pop, 
            by = c("aire_de_sante", "zone_de_sante")) %>%
  left_join(zone_de_sante_pop, 
            by = c("zone_de_sante")) 
  
if(!allNA(x$as_population)){
  x <-   x %>% 
    mutate(total_population = total_population)
}


```


<!-- ======================================================= -->
## Check population linkage (review regularly)

```{r check_population_linkage}

# Calculating number of individuals successfully linked to population data
number_individuals <- nrow(x)
number_individuals_aire_linked <- sum(!is.na(x$as_population))
prop_linked <- sum(number_individuals_aire_linked / number_individuals)

# Compare these two quantities - ensure there aren't any mispellings we're missing
table(x$aire_de_sante[is.na(x$as_population)]) # aire de santes in database we're not matching

# Identifying Aire de Santes with missing population data
komanda_ha_pop <- pop_data$population
index <- which(is.na(komanda_ha_pop))
non_pop_data <- pop_data$aire_de_sante[index]

# Identifying number of alerts linked to Aire de Sante with 
# missing population data
alerts_no_pop <- sum(x$aire_de_sante %in% non_pop_data)

```

There are `r length(non_pop_data)` aire_de_santes without population data. Rate will no be calculated for these areas. 

<!-- ======================================================= -->
## Expected decisions


Exceptionally for Komanda, we remove  data before Sept,2nd 2019 for the analysis on expected decisions: date until which they retrospectively filled the symptoms variables in different columns (before thatd ate, empty symptoms which may lead to incorrect rule decisions).


Alerts are supposed to be validated using the following key, depending on
whether the person has:

1. **known contact** with a case and at least **one symptom**
2. if not 1, **unexplained bleeding**
3. if not 2, **fever and 3 other symptoms**

Note that 3 de facto excludes bleeding as a candidate symptom. We create a new
variable which asserts these conditions:

```{r expected_result}

## elements of diagnostic
## criteria 1
has_contact <- outcomes$contact_connu %in% c("confirme", "oui")
has_one_symptom <- outcomes %>%
  select(fievre,
         nausee_vom,
         diarrhee,
         fatigue,
         anorexie,
         dlr_abdo,
         dlr_thorax,
         dlr_muscu,
         dlr_artic,
         cephalee,
         toux,
         dyspnee,
         dysphagie,
         odynophagie, 
         hoquet,
         bleed) %>%
  apply(1, function(e) any(e == "oui", na.rm = TRUE))
fits_1 <- has_contact & has_one_symptom

## criteria 2
has_bleeding <- outcomes$bleed == "oui"
fits_2 <- !fits_1 & has_bleeding

## criteria 3
has_fever <- outcomes$fievre == "oui"
has_3_symptoms <- outcomes %>%
    select(nausee_vom,
         diarrhee,
         fatigue,
         anorexie,
         dlr_abdo,
         dlr_thorax,
         dlr_muscu,
         dlr_artic,
         cephalee,
         toux,
         dyspnee,
         dysphagie,
         odynophagie, 
         hoquet,
         bleed) %>% 
  apply(1, function(e) sum(e == "oui", na.rm = TRUE)) >= 3
fits_3 <- !fits_1 & !fits_2 & has_fever & has_3_symptoms


## make sure criteria are exclusive - no number in the table below should exceed
## 1. i.e this table should only zero or one as categories.
table(fits_1 + fits_2 + fits_3)

## expected decisions
outcomes <- outcomes %>%
  mutate(admit_contact = fits_1,
         admit_bleeding = fits_2,
         admit_fever = fits_3) %>% 
  mutate(expected_decision = ifelse(
             fits_1 | fits_2 | fits_3,
             "validee",
             "invalidee"),
         decision_comparison = case_when(
             status == "validee" & expected_decision == "validee" ~
               "true_positive",
             status == "invalidee" & expected_decision == "invalidee" ~
               "true_negative",
             status == "validee" & expected_decision == "invalidee" ~
               "false_positive",
             status == "invalidee" & expected_decision == "validee" ~
               "false_negative",
             TRUE ~ NA_character_
             ),
         decision_comparison =
           factor(decision_comparison,
                  levels = c("true_positive",
                             "true_negative",
                             "false_positive",
                             "false_negative")))

## remove alerts with missing comparisons, from 2019
outcomes <- outcomes %>%
  filter(!is.na(decision_comparison),
         date >= as.Date("2019-09-02"))


## check that classification is well-made
outcomes %>%
  group_by(status, expected_decision, decision_comparison) %>%
  count()


```


<!-- ======================================================= -->
## Last 21 days

We duplicate the previous datasets, retaining the 21 days leading up to the
current database date.

```{r subset_21_days}

start_date <- database_date - 21
x_recent <- filter(x, date > start_date)
outcomes_recent <- filter(outcomes, date > start_date)

```


<!-- ======================================================= -->
## Alerts counts split

Calculate number of alerts greater than or equal to 100 or less than 100 over all time. Greater than or less than 20 for recent. There are many aire de santes and this helps to restrict what can be seen. 

```{r alert_count_split}

greater_100 <- x %>% count(zone_de_sante,
               aire_de_sante) %>% 
  filter(n >= 100) %>% 
  pull(aire_de_sante)

less_100 <- x %>% count(zone_de_sante,
               aire_de_sante) %>% 
  filter(n < 100) %>% 
  pull(aire_de_sante)

less_100[less_100 %in% greater_100] <- NA
less_100 <- less_100[complete.cases(less_100)]

greater_20_recent <- x_recent %>% count(zone_de_sante,
               aire_de_sante) %>% 
  filter(n >= 20) %>% 
  pull(aire_de_sante)
less_20_recent <- x_recent %>% count(zone_de_sante,
               aire_de_sante) %>% 
  filter(n < 20) %>% 
  pull(aire_de_sante)

less_20_recent[less_20_recent %in% greater_20_recent] <- NA
less_20_recent <- less_20_recent[complete.cases(less_20_recent)]
  
```

<!-- ======================================================= -->
## Custom color scales

We define custom colors for some of the variables used in the plots.

```{r scales_fill}

scale_origins <- scale_fill_manual(
  "Origine",
  values = c(communautaire = "#ffcc00",
             recherche_active = "#c3c388",
             surveillance_passive = "#ff6699",
             point_entree = "#40bf80",
             autre = "#668cff",
             inconnu = "grey",
             check_cleaning_rules = "grey"))

scale_decisions <- scale_fill_manual(
    "Décisions",
    values = c(true_positive = "#94b8b8",
               true_negative = "#8c8cd9",
               false_positive = "#ff8080",
               false_negative = "#b3003b"),
    labels = c(true_positive = "Validation correcte",
               true_negative = "Invalidation correcte",
               false_positive = "Fausse alerte",
               false_negative = "Alerte manquée"))

scale_validations <- scale_fill_manual(
    "Outcome",
    values = c(statut_inconnu = "#BCB4A4", 
               validee = "#D56F3E", 
               invalidee = "#F2C69B"),
    labels = c(validee = "Validée", 
               invalidee = "Invalidée", 
               statut_inconnu = "Statut Inconnu"))


scale_final_outcome <- scale_fill_manual(
    name = "Statut final de l'alerte",
    values = c(confirme = "#ff9999",
               suspect = "#264d73",
               non_cas = "#8cb3d9"),
    labels = c(confirme = "Cas confirme",
               suspect= "Cas suspect",
               non_cas = "Non-cas"
               
               ))

scale_final_outcome_perc <- scale_fill_manual(
    name = "Statut final de l'alerte",
    values = c(perc_confirme = "#ff9999",
               perc_suspect = "#264d73",
               perc_non_cas = "#8cb3d9"),
    labels = c(perc_confirme = "Cas confirme",
              perc_suspect= "Cas suspect",
               perc_non_cas = "Non-cas"
               
               ))


```


# Validation status {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

This section summarises alerts over time according to their validation status i.e. whether they were validated or invalidated.

**Note: All known Aire de Santes are reported in tables**
**the top 14 Aire de Santes are plotted when it comes to plotting over time**
**(due to space constraints).**

<!-- ======================================================= -->
## Overall 

### Weekly, since database start

```{r sous_coord_time}

ggplot(x, aes(x = date, fill = status)) +
  geom_histogram(binwidth = 7, col = "white") +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par semaine",
       title = paste0("Nombre d'alertes par statut de validation - Komanda")) +
  large_txt +
  rotate_x_text(45) + 
  scale_months +
  theme(legend.position = "bottom")

```


### Table - weekly since database start

```{r sous_coord_time_table, fig.keep = "all"}

table_validation_overall_time <- x %>%
  count(epiweek_report, status) %>%
  spread(status, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_validation_overall_time %>%
 show_table()
  
```

<!-- ======================================================= -->
## Overall past 3 weeks

### Daily, past 3 weeks

```{r sous_coord_recent}

ggplot(x_recent, aes(x = date, fill = status)) +
  geom_histogram(binwidth = 1, col = "white") +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par jour",
       title = paste0("Nombre d'alertes par statut de validation - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  scale_weeks +
  theme(legend.position = "bottom")

```

### Table - daily past 3 weeks 

```{r sous_coord_recent_table, fig.keep = "all"}

table_validation_overall_past_3_weeks <- x_recent %>%
  count(date, status) %>%
  spread(status, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_validation_overall_past_3_weeks %>%
 show_table()
  
```

<!-- ======================================================= -->
## Overall proportion validated

```{r sous_co_proportion_validated}

x_prop <- x %>%
  count(epiweek_report_label, status) %>%
  spread(status, n, fill = 0) %>% 
  mutate(total = invalidee + validee,
         p   = prop_to_perc(validee / total),
         lci = prop_ci(validee, total, "lower", TRUE),
         uci = prop_ci(validee, total, "upper", TRUE)) %>%
  select(epiweek_report_label, p, lci, uci)

ggplot(x_prop, aes(x = epiweek_report_label, col = "#D56F3E")) +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.2, colour = NA) +
  geom_point(aes(y = p), size = 2) +
  geom_line(aes(y = p), size = 1) + 
  scale_color_discrete(guide = FALSE) +
  ylim(c(0, 100)) +
  labs(x = "",
       y = "Pourcentage d'alertes validées \npar semaine",
       title = paste0("Proportion d'alertes validées - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_months

```


<!-- ======================================================= -->
## Health Zone 

### Total since database start

``` {r health_zone_total}

ggplot(x, aes(x = top_zones, fill = status)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  scale_validations +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par statut de validation et \nzone de santé - Komanda") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)

```

### Table - total since database start  

```{r table_hz_total_validation, fig.keep = "all"}

table_hz_total_validation <- x %>%
  count(top_zones, status) %>%
  spread(status, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_hz_total_validation %>%
 show_table()
  
```

### Weekly since database start

```{r health_zone_time, fig.width = 12, fig.height = 8}

ggplot(x, aes(x = date, fill = status)) +
  geom_histogram(binwidth = 7, col = "white") +
  facet_wrap( ~ top_zones, scale = "free_y") +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par semaine",
       title = paste0("Nombre d'alertes par statut de validation ",
                      " et zone de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom", strip.text = element_text(size=17)) +
  scale_months

```

### Table - weekly since database start

```{r table_hz_over_time_validation, fig.keep = "all"}

table_hz_over_time_validation <- x %>%
  count(epiweek_report, top_zones, status) %>%
  spread(status, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_hz_over_time_validation %>%
 show_table()
  
```


<!-- ======================================================= -->
## Health Zone - past 3 weeks

### Total past 3 weeks

``` {r health_zone_total_recent}

ggplot(x_recent, aes(x = top_zones, fill = status)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  scale_validations +
  labs(x = "",
       y = "Nombre d'alertes",
       title = paste("Nombre d'alertes par statut de validation",
                     "et \nzone de santé - Komanda"),
      subtitle = "Données des 3 dernières semaines ") +  theme(legend.position = "bottom") +
  large_txt 

```

### Table - total past 3 weeks

```{r table_hz_total_recent_validation, fig.keep = "all"}

table_hz_total_recent_validation <- x_recent %>%
  count(top_zones, status) %>%
  spread(status, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_hz_total_recent_validation %>%
 show_table()
  
```


### Daily past 3 weeks 

```{r health_zone_time_recent, fig.width = 12, fig.height = 8}

ggplot(x_recent, aes(x = date, fill = status)) +
  geom_histogram(binwidth = 1, col = "white") +
  facet_wrap( ~ top_zones, scale = "free_y") +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par jour",
       title = paste0("Nombre d'alertes par statut de validation ",
                      "et zone de sante - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom", strip.text = element_text(size=17)) +
  scale_weeks

```

### Table - daily past 3 weeks 

```{r table_hz_over_time_recent_validation, fig.width = 12, fig.height = 5}

table_hz_over_time_recent_validation <- x_recent %>%
  count(date, top_zones, status) %>%
  spread(status, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_hz_over_time_recent_validation %>%
 show_table()
  
```

<!-- ======================================================= -->
## Health Area 

### Total since database start

Graphs are split by greater or less than 100 alerts removed all are kept in table 

```{r health_area_total_greater100, fig.width = 14}

x_validations <- x %>%
  count(status, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% greater_100)

ggplot(x_validations, aes(x = aire_de_sante, y = n, fill = status)) +
  geom_col() +
  scale_x_discrete(drop = TRUE) +
  scale_validations +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par statut de validation et aire de santé - Komanda",
       subtitle = "avec plus de 100 alertes") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```

```{r health_area_total_less100, fig.width = 14}

x_validations <- x %>%
  count(status, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% less_100)

ggplot(x_validations, aes(x = aire_de_sante, y = n, fill = status)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_validations +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par statut de validation et aire de santé - Komanda",
       subtitle = "avec moins de 100 alertes") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```

### Table - total since database start  

```{r table_ha_total_validation, fig.keep = "all"}

table_ha_total_validation <- x %>%
  count(aire_de_sante, status) %>%
  spread(status, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_ha_total_validation %>%
 show_table()
  
```

### Weekly since database start

```{r health_area_time, fig.width = 12, fig.height = 8}

ggplot(x, aes(x = epiweek_report_label, fill = status)) +
  geom_bar() +
  facet_wrap( ~ top_aires) +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par semaine",
       title = paste0("Nombre d'alertes par statut de validation",
                      " \net aire de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b") 

```

### Table - weekly since database start

```{r table_ha_over_time_validation, fig.keep = "all"}

table_ha_over_time_validation <- x %>%
  count(epiweek_report, top_aires, status) %>%
  spread(status, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_ha_over_time_validation %>%
 show_table()
  
```



<!-- ======================================================= -->
## Health Area - past 3 weeks

### Total past 3 weeks

Graphs are split by greater or less than 20 alerts removed all are kept in table 

```{r health_area_total_recent_greater20, fig.width = 14}

x_validations <- x_recent %>%
  count(status, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% greater_20_recent)  


ggplot(x_validations, aes(x = aire_de_sante, y = n, fill = status)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_validations +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = paste("Nombre d'alertes par statut de validation et aire de santé",
                     "\n(avec plus de 20 alertes) - Komanda"),
        subtitle = "Données des 3 dernières semaines") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```


```{r health_area_total_recent_less20, fig.width = 14}

x_validations <- x_recent %>%
  count(status, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% less_20_recent)


ggplot(x_validations, aes(x = aire_de_sante, y = n, fill = status)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_validations +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = paste("Nombre d'alertes par statut de validation et aire de santé",
                     "\n(avec moins de 20 alertes) - Komanda"),
        subtitle = "Données des 3 dernières semaines") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```


### Table - total past 3 weeks 

```{r table_ha_total_recent_validation, fig.keep = "all"}

table_ha_total_recent_validation <- x_recent %>%
  count(aire_de_sante, status) %>%
  spread(status, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_ha_total_recent_validation %>%
 show_table()
  
```

### Daily, past 3 weeks

```{r health_area_time_recent, fig.width = 12, fig.height = 8}

x_validations <- x_recent %>%
  filter(top_aires != "other") %>%
  count(date, status, top_aires)

ggplot(x_validations, aes(x = date, y = n, fill = status)) +
  geom_col() +
  facet_wrap( ~ top_aires) +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par jour",
       title = paste0("Nombre d'alertes par statut de validation",
                      " et aire de santé - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "1 week", date_labels =  "%d %b") 
  
```


### Table - daily past 3 weeks

```{r table_ha_over_time_recent_validation , fig.width = 12, fig.height = 5}

table_ha_over_time_recent_validation <- x_recent %>%
  count(date, top_aires, status) %>%
  spread(status, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_ha_over_time_recent_validation %>%
 show_table()
  
```


<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Origins {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

This section summarises alerts over time according to their origin 

**Note: All known Aire de Santes are reported in tables**
**the top 14 Aire de Santes are plotted when it comes to plotting over time**
**(due to space constraints).**



<!-- ======================================================= -->
## Overall

### Weekly, since database start

```{r sous_coord_origins}

ggplot(x, aes(x = epiweek_report_label, fill = origin)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  scale_origins +
  guides(fill = guide_legend(ncol = 2)) +
  labs(x = "",
       y = "Nombre d'alertes par semaine",
       title = "Nombre d'alertes par origine - Komanda") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45) +
  scale_months

```

### Table - weekly since database start

```{r sous_coord_time_table_origins, fig.keep = "all"}

table_origins <- x %>%
  count(epiweek_report, origin) %>%
  spread(origin, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_origins %>%
 show_table()
  
```

<!-- ======================================================= -->
## Overall past 3 weeks

### Daily, past 3 weeks

```{r origins_time_recent}

ggplot(x_recent, aes(x = date, fill = origin)) +
  geom_histogram(binwidth = 1, col = "white") +
  theme(strip.text.y = element_text(size = 12, angle = 0)) +
  scale_months +
  scale_origins +
  guides(fill = guide_legend(ncol = 2)) +
  rotate_x_text(45) +
  large_txt +
  theme(legend.position = "bottom",
        panel.spacing.y = unit(1, "lines")) +
  labs(title = "Nombre d'alertes par origine - Komanda",
       subtitle = "Données des 3 dernières semaines ",
       x = "",
       y = "Nombre d'alertes par jour") +
  scale_weeks

```

### Table - daily past 3 weeks

```{r table_origins_recent, fig.keep = "all"}

table_origins_recent <- x_recent %>%
  count(date, origin) %>%
  spread(origin, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_origins_recent %>%
 show_table()
  
```

<!-- ======================================================= -->
## Health Zone 

### Total since database start

``` {r health_zone_total_origin}

ggplot(x, aes(x = top_zones, fill = origin)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  scale_origins +
  guides(fill = guide_legend(ncol = 2)) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par origine et zone de santé - Komanda") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)

```

### Table - total since database start  

```{r table_hz_total_origins, fig.keep = "all"}

table_hz_total_origins <- x %>%
  count(top_zones, origin) %>%
  spread(origin, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_hz_total_origins %>%
 show_table()
  
```

### Weekly since database start

```{r health_zone_time_origin, fig.width = 12, fig.height = 8}

ggplot(x, aes(x = date, fill = origin)) +
  geom_histogram(binwidth = 7, col = "white") +
  facet_wrap( ~ top_zones, scale = "free_y") +
  scale_origins + 
  labs(x = "",
       y = "Nombre d'alertes par semaine",
       title = paste0("Nombre d'alertes par origine de validation ",
                      "et zone de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_months

```

### Table - weekly since database start

```{r table_hz_over_time_origins, fig.keep = "all"}

table_hz_over_time_origins <- x %>%
  count(epiweek_report, top_zones, origin) %>%
  spread(origin, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_hz_over_time_origins %>%
 show_table()
  
```

<!-- ======================================================= -->
## Health Zone - Past 3 Weeks

### Total past 3 weeks

``` {r health_zone_total_recent_origin_recent}

ggplot(x_recent, aes(x = top_zones, fill = origin)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  scale_origins +
  guides(fill = guide_legend(ncol = 2)) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = paste("Nombre d'alertes par origine et zone de santé - Komanda"), 
        subtitle = "Données des 3 dernières semaines ") +
  theme(legend.position = "bottom") +
  large_txt 

```

### Table - total past 3 weeks

```{r table_hz_total_recent_origins, fig.keep = "all"}

table_hz_total_recent_origins <- x_recent %>%
  count(top_zones, origin) %>%
  spread(origin, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_hz_total_recent_origins %>%
 show_table()
  
```


### Daily past 3 weeks 

```{r health_zone_time_recent_origins, fig.width = 12, fig.height = 8}

ggplot(x_recent, aes(x = date, fill = origin)) +
  geom_histogram(binwidth = 1, col = "white") +
  facet_wrap( ~ top_zones, scale = "free_y") +
  scale_origins + 
  labs(x = "",
       y = "Nombre d'alertes par jour",
       title = paste0("Nombre d'alertes par origine",
                      " et zone de santé - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_weeks 

```

### Table - daily past 3 weeks 

```{r table_hz_over_time_recent_origins, fig.width = 12, fig.height = 5}

table_hz_over_time_recent_origins <- x_recent %>%
  count(date, top_zones, origin) %>%
  spread(origin, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_hz_over_time_recent_origins %>%
 show_table()
  
```

<!-- ======================================================= -->
## Health Area 

### Total since database start

Graphs are split by greater or less than 100 alerts removed all are kept in table. 

```{r health_area_total_origins_greter100, fig.width = 14}

x_origins <- x %>%
  count(origin, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% greater_100) 


ggplot(x_origins, aes(x = aire_de_sante, y = n, fill = origin)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_origins +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par origine et aire de santé - Komanda",
       subtitle = "avec plus de 100 alertes") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```


```{r health_area_total_origins_less100, fig.width = 14}

x_origins <- x %>%
  count(origin, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% less_100)


ggplot(x_origins, aes(x = aire_de_sante, y = n, fill = origin)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_origins +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par origine et aire de santé - Komanda",
       subtitle = "avec moins de 100 alertes") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```

### Table - total since database start  


```{r table_ha_total_origins, fig.keep = "all"}

table_ha_total_origins <- x %>%
  count(aire_de_sante, origin) %>%
  spread(origin, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_ha_total_origins %>%
 show_table()
  
```

### Weekly since database start

```{r health_area_time_origin, fig.width = 12, fig.height = 8}

x_origins <- x %>%
  filter(top_aires != "other") 

ggplot(x_origins, aes(x = epiweek_report_label, fill = origin)) +
  geom_bar() +
  facet_wrap( ~ top_aires) +
  scale_origins + 
  labs(x = "",
       y = "Nombre d'alertes par semaine",
       title = paste0("Nombre d'alertes par origine",
                      " et zone de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b") 

```

### Table - weekly since database start

```{r table_ha_over_time_origins, fig.keep = "all"}

table_ha_over_time_origins <- x %>%
  count(epiweek_report, top_aires, origin) %>%
  spread(origin, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_ha_over_time_origins %>%
 show_table()
  
```


<!-- ======================================================= -->
## Health Area - past 3 weeks

### Total past 3 weeks

Graphs are split by greater or less than 20 alerts removed all are kept in table. 

``` {r health_area_total_recent_origin_greater20, fig.width = 14}

x_origins <- x_recent %>%
  count(origin, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% greater_20_recent)

ggplot(x_origins, aes(x = aire_de_sante, y = n, fill = origin)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_origins +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = paste("Nombre d'alertes par origine et aire de santé",
                     "\n(avec plus de 20 alertes) - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```


``` {r health_area_total_recent_origin_less20, fig.width = 14}

x_origins <- x_recent %>%
  count(origin, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% less_20_recent)  

ggplot(x_origins, aes(x = aire_de_sante, y = n, fill = origin)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_origins +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = paste("Nombre d'alertes par origine et aire de santé",
                     "\n(avec moins de 20 alertes) - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```


### Table - Total past 3 weeks 

```{r table_ha_total_recent_origins, fig.keep = "all"}

table_ha_total_recent_origins <- x_recent %>%
  count(aire_de_sante, origin) %>%
  spread(origin, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_ha_total_recent_origins %>%
 show_table()
  
```

### Daily, past 3 weeks

```{r health_area_time_recent_origin, fig.width = 12, fig.height = 8}

x_origins_recent <- x_recent %>%
  filter(top_aires != "other") %>%
  count(date, origin, top_aires) 

ggplot(x_origins_recent, aes(x = date, y = n, fill = origin)) +
  geom_col() +
  facet_wrap( ~ top_aires) +
  scale_origins + 
  labs(x = "",
       y = "Nombre d'alertes par jour",
       title = paste0("Nombre d'alertes par origine",
                      " et aire de santé - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "1 week", date_labels =  "%d %b") 
  
```


### Table - daily past 3 weeks

```{r table_ha_over_time_recent_origins, fig.width = 12, fig.height = 5}

table_ha_over_time_recent_origins <- x_recent %>%
  count(date, top_aires, origin) %>%
  spread(origin, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_ha_over_time_recent_origins %>%
 show_table()
  
```


<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Validation decisions {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

This section focuses on the decision of alerts, by health zones.

When looking at decisions on the treatment of alerts, we identify the 4
following situations:

* **<font color = "#94b8b8">true positive</font>**: alerts were rightfully
  validated (patient tested)
* **<font color = "#8c8cd9">true negative</font>**: alerts were rightfully not
  validated (patient not tested)
* **<font color = "#ff8080">false positive</font>**: alerts were wrongly
  validated (patient tested, shoud not have been)
* **<font color = "#b3003b">false negative</font>**: alerts were wrongly not
  validated (patient not tested, should have been)

**False positive** create a waste of resources as well as un-necessary pressure
on the patient and community. **False negative** create a risk of missing cases.

For better communication purposes, false-positives will be called "Fausse alerte" and false-negatives will be called  "Alerte manquée".


**Note: All known Aire de Santes are reported in tables**
**the top 14 Aire de Santes are plotted when it comes to plotting over time**
**(due to space constraints).**


<!-- ======================================================= -->
## Overall

### Weekly, since database start

```{r decisions}

ggplot(outcomes, aes(x = epiweek_report_label, fill = decision_comparison)) +
  geom_bar(color = "white") +
  scale_x_discrete(drop = FALSE) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par décision de validation - Komanda") +
  theme(legend.position = "bottom") +
  guides(fill=guide_legend(ncol=2)) +
  large_txt +
  scale_decisions +
  rotate_x_text(45) +
  scale_months

```

### Table - weekly since database start 

``` {r table_decisions}

table_decisions <- outcomes %>%
  count(epiweek_report, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>% 
  adorn_totals(c("row", "col")) %>%
  mutate(prop_false_positive_95ci =   
           prop_to_display_ci(false_positive, Total, dec = 2, perc = TRUE),
         prop_false_negative_95ci = 
           prop_to_display_ci(false_negative, Total, dec = 2, perc = TRUE),
         prop_false_positive = prop_to_perc(false_positive / Total),
         lower_false_positive = prop_ci(false_positive, Total, "lower", TRUE),
         upper_false_positive = prop_ci(false_positive, Total, "upper", TRUE),
         prop_false_negative = prop_to_perc(false_negative / Total),
         lower_false_negative = prop_ci(false_negative, Total, "lower", TRUE),
         upper_false_negative = prop_ci(false_negative, Total, "upper", TRUE))

table_decisions %>%
    select(-prop_false_positive,
          -lower_false_positive,
          -upper_false_positive,
          -prop_false_negative ,
          -lower_false_negative,
          -upper_false_negative
         ) %>% 
  show_table()

```


### Table - incorrectly validated/not validated

* Individuals incorrectly validated. They were validated and tested but did not meet the case definition.
* Individuals incorrectly not validated. They were not validated and tested but did meet the case definition.

```{r table_sens_spec}

table_sens_spec <- outcomes %>%
  count(top_zones, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>% 
  mutate(total_tested = true_positive + false_positive,
         total_not_tested = true_negative + false_negative) %>% 
  mutate(incorrectly_validated_95ci = 
           prop_to_display_ci(false_positive, total_tested, 
                              dec = 2, perc = TRUE),
         incorrectly_not_validated_95ci = 
           prop_to_display_ci(false_negative, total_not_tested, 
                              dec = 2, perc = TRUE))
table_sens_spec %>% 
  show_table()

```

`incorrectly validated` refers to the proportion of validated alerts which should not have been validated. `incorrectly invalidated` refers to the proportion of invalidated alerts that should have been validated.  

**Note:** The `incorrectly validated` variable will be over-estimated if alerts forms are poorly documented/incomplete - i.e. if patients had symptoms not reported in the form but that were used in the decision making process. 


<!-- ======================================================= -->
## Overall past 3 weeks

### Daily, past 3 weeks

```{r decisions_time}

ggplot(outcomes_recent, aes(x = date, fill = decision_comparison)) +
  geom_histogram(binwidth = 1, col = "white") +
  theme(strip.text.y = element_text(size = 12, angle = 0)) +
  scale_months +
  rotate_x_text(45) +
  large_txt +
  scale_decisions +
  theme(legend.position = "bottom") +
  guides(fill=guide_legend(ncol=2)) +
  labs(title = paste("Nombre d'alertes par décision de validation - Komanda"),
       subtitle = "Données des 3 dernières semaines",
       x = "",
       y = "Nombre d'alertes par jour") +
  scale_weeks

```

### Over time table

``` {r table_decisions_recent}

table_decisions_recent <- incidence(outcomes_recent$date, "day", 
                            groups = outcomes_recent$decision_comparison) %>%
  as.data.frame() %>%
  adorn_totals(where = c("row", "col"))

table_decisions_recent %>%
  show_table()

```


<!-- ======================================================= -->
## Health Zone 

### Total since database start

``` {r health_zone_total_decision}

ggplot(outcomes, aes(x = top_zones, fill = decision_comparison)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  scale_decisions +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par décision et zone de santé - Komanda") +
  theme(legend.position = "bottom") +
  guides(fill=guide_legend(ncol=2)) +
  large_txt +
  rotate_x_text(45)

```

### Table - Total since database start  

```{r table_hz_total_decisions, fig.keep = "all"}

table_hz_total_decisions <- outcomes %>%
  count(top_zones, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_hz_total_decisions %>%
 show_table()
  
```

### Weekly since database start

```{r health_zone_time_decision, fig.width = 12, fig.height = 8}

ggplot(outcomes, aes(x = date, fill = decision_comparison)) +
  geom_histogram(binwidth = 7, col = "white") +
  facet_wrap( ~ top_zones, scale = "free_y") +
  scale_decisions + 
  labs(x = "",
       y = "Nombre d'alertes par semaine",
       title = paste0("Nombre d'alertes par décision de validation ",
                      "et zone de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  guides(fill=guide_legend(ncol=2)) +
  scale_months

```

### Table - weekly since database start

```{r table_hz_over_time_decisions, fig.keep = "all"}

table_hz_over_time_decisions <- outcomes %>%
  count(epiweek_report, top_zones, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_hz_over_time_decisions %>%
 show_table()
  
```

<!-- ======================================================= -->
## Health Zone - past 3 weeks

### Total past 3 weeks

``` {r health_zone_total_recent_decision}

ggplot(outcomes_recent, aes(x = top_zones, fill = decision_comparison)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  scale_decisions +
  labs(x = "",
       y = "Nombre d'alertes",
       title = paste("Nombre d'alertes par décision et", " \nzone de santé - Komanda"),
        subtitle = "Données des 3 dernières semaines ") +
  theme(legend.position = "bottom") +
  guides(fill=guide_legend(ncol=2)) +
  large_txt

```

### Table - total past 3 weeks

```{r table_hz_total_recent_decisions, fig.keep = "all"}

table_hz_total_recent_decisions <- outcomes_recent %>%
  count(top_zones, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_hz_total_recent_decisions %>%
 show_table()
  
```

### Daily past 3 weeks 

```{r health_zone_time_recent_decisions, fig.width = 12, fig.height = 8}

ggplot(outcomes_recent, aes(x = date, fill = decision_comparison)) +
  geom_histogram(binwidth = 1, col = "white") +
  facet_wrap( ~ top_zones, scale = "free_y") +
  scale_decisions + 
  labs(x = "",
       y = "Nombre d'alertes par jour",
       title = paste0("Nombre d'alertes par décision de validation ",
                      "et zone de santé - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  guides(fill=guide_legend(ncol=2)) +
  scale_weeks 

```

### Table - daily past 3 weeks 

```{r table_hz_over_time_recent_decisions, fig.width = 12, fig.height = 5}

table_hz_over_time_recent_decisions <- outcomes_recent %>%
  count(date, top_zones, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_hz_over_time_recent_decisions %>%
 show_table()
  
```


<!-- ======================================================= -->
## Health Area 

### Total since database start

Graphs are split by greater or less than 100 alerts removed all are kept in table. 

```{r health_area_total_decisionsgreater100, fig.width = 14}

x_decisions <- outcomes %>%
  count(decision_comparison, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% greater_100)  

ggplot(x_decisions, 
       aes(x = aire_de_sante, y = n, fill = decision_comparison)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_decisions +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par décision de validation", " \net aire de santé - Komanda",
       subtitle = "avec plus de 100 alertes") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```

```{r health_area_total_decisionsless100, fig.width = 14}

x_decisions <- outcomes %>%
  count(decision_comparison, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% less_100)

ggplot(x_decisions, 
       aes(x = aire_de_sante, y = n, fill = decision_comparison)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_decisions +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = "Nombre d'alertes par décision de validation", " \net aire de santé - Komanda",
       subtitle = "avec moins de 100 alertes") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```

### Table - total since database start  


```{r table_ha_total_decision, fig.keep = "all"}

table_ha_total_decisions <- outcomes %>%
  count(aire_de_sante, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_ha_total_decisions %>%
 show_table()
  
```

### Weekly since database start

```{r health_area_time_decision, fig.width = 12, fig.height = 8}

x_decisions <- outcomes %>%
  filter(top_aires != "other")

ggplot(x_decisions, aes(x = epiweek_report_label, fill = decision_comparison)) +
  geom_bar() +
  facet_wrap( ~ top_aires) +
  scale_decisions + 
  labs(x = "",
       y = "Nombre d'alertes par semaine",
       title = paste0("Nombre d'alertes par décision de validation",
                      " et aire de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b") 

```

### Table - weekly since database start

```{r table_ha_over_time_decisions, fig.keep = "all"}

table_ha_over_time_decisions <- outcomes %>%
  count(epiweek_report, top_aires, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_ha_over_time_decisions %>%
 show_table()
  
```


<!-- ======================================================= -->
## Health Area - past 3 weeks

### Total past 3 weeks

Graphs are split by greater or less than 20 alerts removed all are kept in table.  

```{r health_area_total_recent_decision_greater20, fig.width = 14}

x_decisions <- outcomes_recent %>%
  count(decision_comparison, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% greater_20_recent)  

ggplot(x_decisions, 
       aes(x = aire_de_sante, y = n, fill = decision_comparison)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_decisions +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = paste0("Nombre d'alertes par décision de validation", " \net aire de santé (avec plus de 20 alertes) - Komanda"),
       subtitle = "Données des 3 dernières semaines") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```

```{r health_area_total_recent_decision_less20, fig.width = 14}

x_decisions <- outcomes_recent %>%
  count(decision_comparison, zone_de_sante, aire_de_sante) %>%
  filter(aire_de_sante %in% less_20_recent)    

ggplot(x_decisions, 
       aes(x = aire_de_sante, y = n, fill = decision_comparison)) +
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  scale_decisions +
  facet_grid(. ~ zone_de_sante, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes",
       title = paste0("Nombre d'alertes par décision de validation", " \net aire de santé (avec moins de 20 alertes) - Komanda"),
       subtitle = "Données des 3 dernières semaines") +
  theme(legend.position = "bottom") +
  large_txt +
  rotate_x_text(45)


```


### Table - total past 3 weeks 

```{r table_ha_total_recent_decisions, fig.keep = "all"}

table_ha_total_recent_decisions <- outcomes_recent %>%
  count(aire_de_sante, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>% 
  adorn_totals(where = c("row", "col"))

table_ha_total_recent_decisions %>%
 show_table()
  
```

### Daily, past 3 weeks

```{r health_area_time_recent_decision_recent, fig.width = 12, fig.height = 8}

x_decisions_recent <- outcomes_recent %>%
  filter(top_aires != "other",
         date > start_date)

ggplot(x_decisions_recent, aes(x = date, fill = decision_comparison)) +
  geom_bar() +
  facet_wrap( ~ top_aires) +
  scale_decisions + 
  labs(x = "",
       y = "Nombre d'alertes par jour",
       title = paste0("Nombre d'alertes par décision de validation",
                      " et aire de santé - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "1 week", date_labels =  "%d %b") 
  
```


### Table - daily past 3 weeks

```{r table_ha_over_time_recent_decisions}

table_ha_over_time_recent_decisions <- outcomes_recent %>%
  count(date, top_aires, decision_comparison) %>%
  spread(decision_comparison, n, fill = 0) %>%
  adorn_totals(where = c("row", "col"))

table_ha_over_time_recent_decisions %>%
 show_table()
  
```


<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Alert Rates {.tabset .tabset-fade .tabset-pills}

## Outline

This section integrates information on the number of alerts with population data to estimate the number of validated/invalidated alerts per 10,000 people.

**Note:** Rate are only calculated for areas with population data. There 
are `r length(non_pop_data)` Aire de Santes without population data, which totals
 **`r alerts_no_pop`** alerts.

**Note: All known Aire de Santes are reported in tables**



```{r reactivate_alert_rates, include = FALSE}

if(allNA(x$as_population)){
  knitr::opts_chunk$set(eval = FALSE)
  no_pop_data <- TRUE
  cat("There is no population data therefore alert rates can not be calculated.")
}

```


<!-- ======================================================= -->
## Overall

### Weekly, Since Database Start 

```{r alert_rates}

x_alert_rate_sous_co <- x %>%
  count(epiweek_report_label, status, total_population) %>%
  mutate(alert_rate = 10000 * n / total_population) 

ggplot(data = x_alert_rate_sous_co, 
       aes(x = epiweek_report_label, y = alert_rate, fill = status)) +
  geom_col() + 
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par semaine \net par 10,000 personnes",
       title = paste0("Nombre d'alertes par 10000 personnes -  Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_months

```


### Table - Weekly Since Database Start

```{r table_alert_rates_per_week, fig.keep = "all"}

check_iconnu <- max(x_alert_rate_sous_co$status == "statut_inconnu")

table_alert_rates_per_week <- x_alert_rate_sous_co %>%
  select(-alert_rate) %>% 
  spread(status, n, fill = 0) %>% 
  mutate(total = ifelse(check_iconnu,
              statut_inconnu + invalidee + validee,
              invalidee + validee),
         alert_rate = 10000 * total / total_population,
         alert_rate = round(alert_rate, 4))

table_alert_rates_per_week %>%
  show_table()

```


<!-- ======================================================= -->
## Overall Past 3 Weeks

### Daily, Past 3 Weeks 

``` {r alert_rates_recent}

x_alert_rate_sous_co <- x_recent %>%
  count(date, status, total_population) %>%
  mutate(alert_rate = 10000 * n / total_population) 

ggplot(data = x_alert_rate_sous_co, aes(x = date, 
                                        y = alert_rate, fill = status)) +
  geom_col() + 
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par jour \net par 10,000 personnes",
       title = paste0("Nombre d'alertes par 10000 personnes -  Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_weeks

```

### Table - Weekly Since Database Start

```{r table_alert_rates_recent, fig.keep = "all"}

check_iconnu <- max(x_alert_rate_sous_co$status == "statut_inconnu")


table_alert_rates_recent <- x_alert_rate_sous_co %>%
  select(-alert_rate) %>% 
  spread(status, n, fill = 0) %>% 
  mutate(total = ifelse(check_iconnu,
              statut_inconnu + invalidee + validee,
              invalidee + validee),
         alert_rate = 10000 * total / total_population,
         alert_rate = round(alert_rate, 4))

table_alert_rates_recent %>%
 show_table()
  
```


<!-- ======================================================= -->
## Health Zone

### Total, Since Database Start

``` {r alert_rate_zone_de_sante_total_validation_status}

date_span <- database_date - min(x$date, na.rm = TRUE)
number_weeks <- as.numeric(date_span/7)

x_alert_rate_zone_sante <- x %>%
  filter(top_zones != "other",
         zs_population > 0) %>%
  count(status, zone_de_sante, zs_population) %>%
  mutate(alert_rate = 10000 * n / (zs_population * number_weeks))

ggplot(data = x_alert_rate_zone_sante, aes(x = zone_de_sante, 
                                           y = alert_rate, fill = status)) +
  geom_col() +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par semaine \net par 10,000 personnes",
       title = paste0("Nombre d'alertes par 10000 personnes par statut\n", 
                      "de validation et zone de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom")


```

### Weekly, Since Database Start

``` {r alert_rate_zone_de_sante_weekly_status, fig.height = 5, fig.width = 10}

x_alert_rate_zone_sante <- x %>%
  filter(top_zones != "other",
         zs_population > 0) %>%
  count(epiweek_report_label, status, top_zones, zs_population) %>%
  mutate(alert_rate = 10000 * n / zs_population)

ggplot(data = x_alert_rate_zone_sante, aes(x = epiweek_report_label, 
                                           y = alert_rate, fill = status)) +
  geom_col() +
  facet_wrap( ~ top_zones) +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par semaine \net par 10,000 personnes",
       title = paste0("Nombre d'alertes par 10000 personnes par statut\n", 
                      "de validation, et zone de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b") 
 
```


<!-- ======================================================= -->
## Health Zone Past 3 Weeks

### Total, Past 3 Weeks

``` {r alert_rate_zone_de_sante_weekly_validation_status_recent}

x_alert_rate_zone_sante <- x %>%
  filter(
         as_population > 0,
         top_zones != "other") %>% 
  count(status, top_zones, zs_population) %>%
  mutate(alert_rate = 10000 * n / (zs_population * 3))

ggplot(data = x_alert_rate_zone_sante, aes(x = top_zones, 
                                           y = alert_rate, fill = status)) +
  geom_col() +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par semaine \net par 10,000 personnes",
       title = paste0("Nombre d'alertes par 10000 personnes par statut\n", 
                      "de validation et zone de santé - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  theme(legend.position = "bottom") 


```

### Daily, Past 3 Weeks

``` {r alert_rate_zone_de_sante_daily_status, fig.height = 5, fig.width = 12}

x_alert_rate_zone_sante <- x %>%
  filter(date > start_date,
         zs_population > 0,
         top_zones != "other") %>%
  count(date, status, top_zones, zs_population) %>%
  mutate(alert_rate = 10000 * n / zs_population)

ggplot(data = x_alert_rate_zone_sante, aes(x = date, 
                                           y = alert_rate, fill = status)) +
  geom_col() +
  facet_wrap( ~ top_zones) +
  scale_validations + 
  labs(x = "",
       y = "Nombre d'alertes par jour \net par 10,000 personnes",
       title = paste0("Nombre d'alertes par 10000 personnes par statut\n", 
                      "de validation et zone de santé - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_weeks

```


<!-- ======================================================= -->
## Health Area

### Total, Since Database Start

``` {r alert_rate_aire_de_sante_weekly_val_total, fig.width = 14}

date_span <- database_date - min(x$date)
number_weeks <- as.numeric(date_span/7)

x_alert_rate_zone_sante <- x %>%
  filter(top_zones != "other",
         as_population > 0) %>%
  count(status, top_zones, aire_de_sante, as_population) %>%
  mutate(alert_rate = 10000 * n / (as_population * number_weeks))

ggplot(data = x_alert_rate_zone_sante, aes(x = aire_de_sante, 
                                           y = alert_rate, fill = status)) +
  geom_col() +
  scale_validations + 
  facet_grid(. ~ top_zones, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes par semaine \net par 10,000 personnes",
       title = paste0("Nombre d'alertes par 10000 personnes par statut ", 
                      "de validation et aire de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") 


```

### Weekly, Since Database Start

``` {r alert_rate_aire_de_sante_over_time, fig.width = 12, fig.height = 8}

x_alert_rate_aire_sante <- x %>%
  filter(top_aires != "other",
         as_population > 0) %>%
  count(epiweek_report_label, status, top_aires, as_population) %>%
  mutate(alert_rate = 10000 * n / as_population) 

ggplot(data = x_alert_rate_aire_sante, aes(x = epiweek_report_label, 
                                           y = alert_rate, fill = status)) +
  geom_col() +
  facet_wrap( ~ top_aires) +
  scale_validations +
  labs(x = "",
       y = "Nombre d'alertes par semaine et \npar 10,000 personnes",
       title = paste0("Nombre d'alertes par 10000 personnes par statut", 
                      " de validation et aire de santé - Komanda")) +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b") 

```

<!-- ======================================================= -->
## Health Area - Past 3 Weeks 

``` {r alert_rate_as_3_weeks, fig.width = 14}

x_alert_rate_aire_sante_recent <- x_recent %>%
  filter(top_zones != "other",
         as_population > 0) %>%
  count(status, aire_de_sante, top_zones, as_population) %>%
  mutate(alert_rate = (10000 * n) / (as_population * 3)) 

ggplot(data = x_alert_rate_aire_sante_recent, aes(x = aire_de_sante, 
                                                  y = alert_rate, fill = status)) +
  geom_col() +
  scale_validations +
  facet_grid(. ~ top_zones, scales = "free_x", space = "free") +
  theme(strip.text.x = element_text(size = 12, angle = 0, color = "#5c8a8a"),
        strip.background = element_rect(fill = "#e0ebeb", color = "#5c8a8a")) +
  labs(x = "",
       y = "Nombre d'alertes par 10000 \npersonnes et par semaine",
       title = paste0("Nombre d'alertes par 10000 personnes - Komanda"), 
        subtitle = "Données des 3 dernières semaines ") + 
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") 

```


### Daily, Past 3 Weeks

``` {r alert_rate_aire_de_sante_recent, fig.width = 12, fig.height = 8}

x_alert_rate_aire_sante <- x_recent %>%
  filter(
         top_aires != "other",
         as_population > 0) %>%
  count(date, status, top_aires, as_population) %>%
  mutate(alert_rate = 10000 * n / (as_population)) 

ggplot(data = x_alert_rate_aire_sante, aes(x = date, 
                                           y = alert_rate, fill = status)) +
  geom_col() +
  facet_wrap( ~ top_aires) +
  scale_validations +
  labs(x = "",
       y = "Nombre d'alertes par jour et \npar 10,000 personnes",
       title = paste0("Nombre d'alertes par 10000 personnes par statut", 
                      " de validation et aire de santé - Komanda"),
       subtitle = "Données des 3 dernières semaines ") +
  large_txt +
  rotate_x_text(45) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "1 week", date_labels =  "%d %b") 

```


<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
  

# Final status of alerts {.tabset .tabset-fade .tabset-pills}

## Outline

This section investigates information on the proportions of false positive and false negative that became real cases.

*NOTE*: As of the latest database from Komanda on Dec, 12th 2019, the **final outcome of the alerts**, in the variable `classif_final` was given only for non-cases.

Please add the additionnal codes for the final status analysis once this variable starts being completed.


<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->
# Export data and tables  {.tabset .tabset-fade .tabset-pills}


```{r reactivate_recent, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

<!-- ======================================================= -->
## Outline

We export the clean data to the clean data folder and some of the relevant tables, which will be placed in the current
working directory.



<!-- ======================================================= -->
## Export clean data

We export some of the clean database, placed in `produced_rds/` as well as in
`data/clean/`:

```{r export_rds, eval = TRUE}

## check if a directory exists and if not then creates it
if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}

## create the text for the file name with the database date
rds_file_name <- sprintf("%sclean_%s.rds",
                     undated_file_name(current_komanda),
                     format(database_date, "%Y-%m-%d"))
rds_file_name

## save the rds file in the produced_rds folder
rio::export(x,
            file.path("produced_rds", rds_file_name))

```

We copy these files to the `data/clean` folder:

```{r copy_rds, eval = TRUE}
# copy some files into `data/clean/`

if (!dir.exists("data/clean")) {
  dir.create("data/clean")
}

# Provide the destination of where to copy the data
destination <- here("data",
                    "clean",
                    rds_file_name)
# Copy the rds data
file.copy(from = file.path("produced_rds", rds_file_name),
          to = destination,
          overwrite = TRUE)

```



<!-- ======================================================= -->
## Excel files

The following code exports all tables named in `to_report` to `xslx` files,
stored inside the folder `produced_xlsx`:

### Cleaned alerts database

```{r exports_tables, eval = TRUE}

cleaned_alerts_database_komanda <- x

to_export <- c("cleaned_alerts_database_komanda",
               "table_validation_overall_time",
               "table_validation_overall_past_3_weeks",
               "table_hz_total_validation",
               "table_hz_over_time_validation",
               "table_hz_total_recent_validation",
               "table_hz_over_time_recent_validation",
               "table_ha_total_validation",
               "table_ha_over_time_validation",
               "table_ha_total_recent_validation",
               "table_ha_over_time_recent_validation",
               "table_origins",
               "table_origins_recent",
               "table_hz_total_origins",
               "table_hz_over_time_origins",
               "table_hz_total_recent_origins",
               "table_hz_over_time_recent_origins",
               "table_ha_total_origins",
               "table_ha_over_time_origins",
               "table_ha_total_recent_origins",
               "table_ha_over_time_recent_origins",
               "table_decisions",
               "table_sens_spec",
               "table_decisions_recent",
               "table_hz_total_decisions",
               "table_hz_over_time_decisions",
               "table_hz_total_recent_decisions",
               "table_hz_over_time_recent_decisions",
               "table_ha_total_decisions",
               "table_ha_over_time_decisions",
               "table_ha_total_recent_decisions",
               "table_ha_over_time_recent_decisions",
               "table_unknown_as")
               
```

```{r xlsx_exports, eval = TRUE}

## check if a directory exists and if not then creates it
if (!dir.exists("produced_xlsx")) {
  dir.create("produced_xlsx")
}

for (e in to_export) {
  rio::export(get(e),
              file.path("produced_xlsx",
                        paste0(e, ".xlsx")))
}

```


We copy the main data file to the `data/clean` folder:

```{r export_xlsx, eval = TRUE}

## create the text for the file name with the database date
xlsx_file_name <- sprintf("%sclean_%s.xlsx",
                     undated_file_name(current_komanda),
                     format(database_date, "%Y-%m-%d"))
xlsx_file_name

## save the rds file in the produced_rds folder
rio::export(x,
            file.path("produced_xlsx", xlsx_file_name))

```


```{r copy_xlsx, eval = TRUE}
# copy some files into `data/clean/`

# Provide the destination of where to copy the data
destination <- here("data",
                    "clean",
                    xlsx_file_name)
# Copy the rds data
file.copy(from = file.path("produced_xlsx", xlsx_file_name),
          to = destination,
          overwrite = TRUE)

```



Click on the following links to open the files (only works if the files above
have been generated and are in the same folder as this document):


```{r xlsx_links, results = "asis", eval = TRUE}

for (e in to_export) {
  txt <- sprintf("- [%s.xlsx](%s.xlsx)",
                 e,
                 file.path("produced_xlsx",
                           e))
  cat(txt, sep = "\n")
}

```


<!-- ======================================================= -->
## R objects

The following code exports all tables named in `to_report` to `rds` files,
stored inside the folder `produced_rds`:

```{r rds_exports, eval = TRUE}

if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}

for (e in to_export) {
  rio::export(get(e),
              file.path("produced_rds",
                        paste0(e, ".rds")))
}

```

<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# System information {.tabset .tabset-fade .tabset-pills}

<!-- ======================================================= -->
## Outline

The following information documents the system on which the document was
compiled.


<!-- ======================================================= -->
## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

<!-- ======================================================= -->
## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


<!-- ======================================================= -->
## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```

<!-- ======================================================= -->
## Compilation parameters

This shows which parameters were passed through `params` at compilation time:

```{r params}
params
```
